---
title: "DOMC_Fig04_2024_11_17"
author: "Glade Bogar"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Make figures and supplements for DOMC figure 4 - Total C, C:N ratio by depth and comboined biomass ECM + ECM-associated 16S

#set working directory: setwd("~/Desktop/GAD_ELN/DO/Analysis/DOMC_Fig04_2024_11_17")

##Calculate actual fungal populations (ng/g) by multiplying RNA-based per-sample relative abundances to PLFA categories
#   ITS ASVs, functional groups, per-sample functional groups (in ng/g)

##Calculate ECM-associated 16S ASVs
#   Unfiltered DeSeq2
#   Stat - DeSeq2 Results
#   multiply ECM-associated 16S ASV relative abundance by total bacterial PLFA to get ECM-associated 16S ng/g

##4 figures, paneled:
#   stored C by Depth, Tree Heath
#   C:N by Depth, Tree Heath
#   stored C by ECM + ECM-associated 16S ng/g
#   C:N by ECM + ECM-associated 16S ng/g

#load required packages:
```{r}
library("phyloseq")
library("vegan")
library("tidyverse")
library("DESeq2")
```

#load required data:
```{r}
ps16S <- read_rds("../microeco/Results/ps16S_faprotax.rds")
psITS <- read_rds("../Microbial_Community_ITS/Results/psITS_remove_8_reads_fungaltraits.rds")
fungal_functions <- read_csv("../Microbial_Community_ITS/Results/ITS_Functions_By_Sample/Primary_lifestyle_by_sample.csv")
metadata <- read_csv("../../TidyData/Metadata/Results/per_sample_metadata.csv")
```

## Correct ITS read abundances by their per-sample population (PLFA); multiply RNA-based read counts by PLFA ng/g

#Use DeSeq2 to find significantly correlated 16S ASVs and genera with ECM ng/g, in living and dead trees.

#ASVs, RNA

#read depths:
```{r}
sort(sample_sums(psITS))
```

#subset exclude controls:
```{r}
psITS
psITS <- subset_samples(psITS, core_top_depth_cm >= 0)
psITS
```

#select RNA-only:
```{r}
psITS
psITS_RNA <- subset_samples(psITS, nucleic_acid == "RNA")
psITS_RNA
```

#remove taxa that aren't in the dataset anymore:
```{r}
psITS_RNA
psITS_RNA <- prune_taxa(taxa_sums(psITS_RNA) > 0, psITS_RNA)
psITS_RNA
```

#get sample data:
```{r}
psITS_RNA_sample_data <- data.frame(sample_data(psITS_RNA))
```

#select PLFA data:
```{r}
PLFA_total_fungi <- select(psITS_RNA_sample_data, PLFA_Total_Fungi_ng_g)
PLFA_total_fungi <- PLFA_total_fungi$PLFA_Total_Fungi_ng_g
```

#transform to relative abundances:
```{r}
psITS_RNA_relative <- transform_sample_counts(psITS_RNA, function(x){x / sum(x)})
```

#pre-transform sample sums:
```{r}
sample_sums(psITS_RNA_relative)
```

#transform ASV table:
```{r}
# Divide the counts in sample n by size_factors[n] and save in matrix `mat`
mat <- sweep(otu_table(psITS_RNA_relative), 1 + taxa_are_rows(psITS_RNA_relative), PLFA_total_fungi, FUN = `*`)

# Make a new phyloseq object with the transformed counts
psITS_RNA_relative_PLFAcorr <- psITS_RNA_relative
otu_table(psITS_RNA_relative_PLFAcorr) <- otu_table(mat, taxa_are_rows = taxa_are_rows(psITS_RNA_relative))
```

#post-transform sample sums:
```{r}
sample_sums(psITS_RNA_relative_PLFAcorr)
```

#should match PLFA numbers:
```{r}
PLFA_total_fungi
```

#worked.

#re-calculate per-sample fungal functional traits for append to 16S.

#List 'taxonomic' ranks"
```{r}
rank_names(psITS_RNA_relative_PLFAcorr)
```

#rename (pre-written code):
```{r}
psITS_fungaltraits_relative <- psITS_RNA_relative_PLFAcorr
```

#extract taxa table
```{r}
# Extract abundance matrix from the phyloseq object
TAXA1 <- as(tax_table(psITS_fungaltraits_relative), "matrix")
# Coerce to data.frame
psITS_fungaltraits_taxa <- as.data.frame(TAXA1)
# Save as .csv
#write_csv(psITS_fungaltraits_taxa, "Results/Fungal_Traits_Taxa_Table.csv")
```

#Extract OTU table
```{r}
# Extract abundance matrix from the phyloseq object
OTU1 <- as(otu_table(psITS_fungaltraits_relative), "matrix")
# Coerce to data.frame
psITS_fungaltraits_OTU <- as.data.frame(OTU1)
# Save as .csv
#write_csv(psITS_fungaltraits_OTU, "Results/Fungal_Traits_OTU_Table.csv")
```

#Extract sample names
```{r}
samplenames <- as.data.frame(sample_names(psITS_fungaltraits_relative))
#write_csv(samplenames, "Results/Fungal_Traits_Sample_Names.csv")
```

#Vandepol split genus Mortierella in 2020 - manually assigning genus mortierella's functions to what are now genera actinomortierella, gryganskiella, linnemannia, lunasporangiospora, necromortierella, podila.
#Note this in the methods.

#Add samplenames to first column of OTU_Table:
```{r}
psITS_fungaltraits_OTU <- rownames_to_column(psITS_fungaltraits_OTU, var = "DO_sample_names")
```

### 1:Primary_lifestyle

#replace entire first row of OTU_table with Primary_lifestyle column of Taxa_table.csv, including heading.
```{r}
Primary_lifestyle_transposed <- t(psITS_fungaltraits_taxa$Primary_lifestyle)
Primary_lifestyle_transposed <- as.data.frame(Primary_lifestyle_transposed)
column_title <- as.data.frame(c("Primary_lifestyle"))
Primary_lifestyle_transposed <- bind_cols(column_title, Primary_lifestyle_transposed)
colnames(Primary_lifestyle_transposed) <- colnames(psITS_fungaltraits_OTU)
psITS_fungaltraits_OTU <- data.frame(lapply(psITS_fungaltraits_OTU, as.character), stringsAsFactors = FALSE)
Primary_lifestyle <- bind_rows(Primary_lifestyle_transposed, psITS_fungaltraits_OTU)
```

#transpose, fix column names, remove rownames (ASVs), re-code as numeric:
```{r}
Primary_lifestyle <- t(Primary_lifestyle)

colnames(Primary_lifestyle) <- as.character(Primary_lifestyle[1, ])
Primary_lifestyle <- as.data.frame(Primary_lifestyle[-1, ])
rownames(Primary_lifestyle) <- NULL

recode_numeric <- select(Primary_lifestyle, -Primary_lifestyle)
recode_numeric <- data.frame(lapply(recode_numeric, as.numeric))

Primary_lifestyle <- bind_cols(Primary_lifestyle$Primary_lifestyle, recode_numeric)
Primary_lifestyle <- dplyr::rename(Primary_lifestyle, Primary_lifestyle = "...1")
```

#collapse all duplicate values:
```{r}
metadata_Primary_lifestyle_grouped <- Primary_lifestyle %>%
  group_by(Primary_lifestyle) %>%
  summarise(across(everything(), sum))
```

#Transpose, edit for later leftjoin:
```{r}
metadata_Primary_lifestyle_grouped_t <- as.data.frame(t(metadata_Primary_lifestyle_grouped))
metadata_Primary_lifestyle_grouped_t <- rownames_to_column(metadata_Primary_lifestyle_grouped_t, var = "Sequencing_ID")
```

#check sums:
```{r}
checksums <- metadata_Primary_lifestyle_grouped_t
checksums <- checksums[-1,-1]
checksums <- mutate(checksums, across(everything(), as.numeric))
checksums <- mutate(checksums, traitsums = rowSums(across(everything())))

checksums$traitsums
```

#export collapsed metadata:
```{r}
write_csv(metadata_Primary_lifestyle_grouped_t, "Results/Primary_lifestyle_by_sample.csv", col_names = FALSE)
```

#add ECM ng/g to 16S Metadata

#read depths:
```{r}
sort(sample_sums(ps16S))
```

#subset exclude controls:
```{r}
ps16S
ps16S <- subset_samples(ps16S, core_top_depth_cm >= 0)
ps16S
```

#select RNA-only:
```{r}
ps16S
ps16S_RNA <- subset_samples(ps16S, nucleic_acid == "RNA")
ps16S_RNA
```

#remove taxa that aren't in the dataset anymore:
```{r}
ps16S_RNA
ps16S_RNA <- prune_taxa(taxa_sums(ps16S_RNA) > 0, ps16S_RNA)
ps16S_RNA
```

#get sample data:
```{r}
ps16S_RNA_sample_data <- data.frame(sample_data(ps16S_RNA))
```

#load 'fungal functions (created above):
```{r}
fungal_functions <- read_csv("Results/Primary_lifestyle_by_sample.csv")
```

#extract, join data:
```{r}
psITS_metadata <- data.frame(sample_data(psITS_RNA_relative_PLFAcorr))
psITS_metadata <- rownames_to_column(psITS_metadata, var = "SeqID")
fungal_functions <- dplyr::rename(fungal_functions, SeqID = "Primary_lifestyle")

psITS_metadata <- left_join(psITS_metadata, fungal_functions)

ps16S_metadata <- data.frame(sample_data(ps16S_RNA))
ps16S_metadata <- rownames_to_column(ps16S_metadata, var = "SeqID")

psITS_metadata <- select(psITS_metadata, nucleic_acid, DO_ID, ectomycorrhizal)

ps16S_metadata$DO_Name <- str_c(ps16S_metadata$DO_ID, ps16S_metadata$nucleic_acid, sep = "_")
psITS_metadata$DO_Name <- str_c(psITS_metadata$DO_ID, psITS_metadata$nucleic_acid, sep = "_")
psITS_metadata <- select(psITS_metadata, DO_Name, ectomycorrhizal)

ps16S_metadata <- left_join(ps16S_metadata, psITS_metadata, by = "DO_Name")

ps16S_metadata <- column_to_rownames(ps16S_metadata, var = "SeqID")

sample_data(ps16S_RNA) <- sample_data(ps16S_metadata)

ps16S_RNA
```

#ECM ng/g-associated 16S

#ASVs

#RNA

#filter: present at at least 2 sites

#designate ASVs (pre-written code):
```{r}
ps16S_ASVs_RNA <- ps16S_RNA
```

#To start, filter for asvs seen in at least 1 time under each of 2 trees:
```{r}
ps16S_ASVs_RNA_for_filter <- ps16S_ASVs_RNA

ps16S_ASVs_RNA_for_filter <- merge_samples(ps16S_ASVs_RNA_for_filter, "Tree")
ps16S_ASVs_RNA_for_filter

ps16S_ASVs_RNA_abundant <- filter_taxa(ps16S_ASVs_RNA_for_filter, function(x) sum(x > 1) >= 2, TRUE)
ps16S_ASVs_RNA_abundant
```

#How much of the community do these represent?

#First, relativize phyloseq:
```{r}
ps16S_ASVs_RNA_relative <- transform_sample_counts(ps16S_ASVs_RNA, function(x){(x / sum(x))})
```

#Next, get names of subsetted taxa:
```{r}
ps16S_ASVs_RNA_abundant_names <- names(taxa_sums(ps16S_ASVs_RNA_abundant))
```

#Subset relativized dataset:
```{r}
ps16S_ASVs_RNA_relative
ps16S_ASVs_RNA_relative_abundant <- prune_taxa(ps16S_ASVs_RNA_abundant_names, ps16S_ASVs_RNA_relative)
ps16S_ASVs_RNA_relative_abundant
```

#Bar graph:
```{r}
plot_bar(ps16S_ASVs_RNA_relative_abundant, x = "DO_ID", y = "Abundance", fill = "Phylum")
```

#pretty good.

#Subset un-relativized dataset for analysis:
```{r}
ps16S_ASVs_RNA_abundant
ps16S_ASVs_RNA_abundant_onepercent <- prune_taxa(ps16S_ASVs_RNA_abundant_names, ps16S_ASVs_RNA)
ps16S_ASVs_RNA_abundant_onepercent

sort(sample_sums(ps16S_ASVs_RNA_abundant_onepercent))
```

#deseq2 does not run on relativized data. rarefy to common depth, post-filtering:
```{r}
sort(sample_sums(ps16S_ASVs_RNA_abundant_onepercent))

ps16S_ASVs_RNA_abundant_onepercent <- rarefy_even_depth(ps16S_ASVs_RNA_abundant_onepercent, sample.size = 24491, rngseed = 12, replace = TRUE, trimOTUs = TRUE)

sort(sample_sums(ps16S_ASVs_RNA_abundant_onepercent))
```


#Test ECM ng/g as a numeric predictor.

#Filter for only samples with %ECM:
```{r}
ps16S_ASVs_RNA_abundant_onepercent
ps16S_ASVs_RNA_abundant_onepercent <- subset_samples(ps16S_ASVs_RNA_abundant_onepercent, ectomycorrhizal != "NA")
ps16S_ASVs_RNA_abundant_onepercent
```

#scale and center ECM - to interpret as logfold change per 100 ng/g ECM increments:
```{r}
ps16S_ASVs_RNA_abundant_onepercent_sample_data <- data.frame(sample_data(ps16S_ASVs_RNA_abundant_onepercent))

ps16S_ASVs_RNA_abundant_onepercent_sample_data <- mutate(ps16S_ASVs_RNA_abundant_onepercent_sample_data, ECM_scaled = (((ectomycorrhizal) - mean(ectomycorrhizal))/100))

sample_data(ps16S_ASVs_RNA_abundant_onepercent) <- ps16S_ASVs_RNA_abundant_onepercent_sample_data

ps16S_ASVs_RNA_abundant_onepercent
```

#use deseq2 to find microbes that are differentially abundant with greater % ECM:
```{r}
DESeq_ECM <- phyloseq_to_deseq2(physeq = ps16S_ASVs_RNA_abundant_onepercent, design = ~ECM_scaled)
DESeq_ECM <- DESeq(DESeq_ECM, test = "Wald", fitType = "parametric")
```

#DESeq2 Results:
```{r}
DESeq_ECM_Results <- results(DESeq_ECM, cooksCutoff = FALSE)
alpha <- 0.05
sigtab <- DESeq_ECM_Results[which(DESeq_ECM_Results$padj < alpha), ]
sigtab <- cbind(as(sigtab, "data.frame"), as(tax_table(ps16S_ASVs_RNA_abundant_onepercent)[rownames(sigtab), ], "matrix"))
dim(sigtab)

write_csv(sigtab, "Results/DESeq2_16S_ASVs_RNA_ECM_2trees.csv")
```

#Visualize
```{r}
library("ggplot2")
theme_set(theme_bw())
#scale_fill_discrete <- function(palname = "Set1", ...) {
#    scale_fill_brewer(palette = palname, ...)
#}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
```

#many significant asvs 

#export:
```{r}
pdf(file = "Results/DESeq2_16S_ASVs_RNA_ECM_2trees.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches

library("ggplot2")
theme_set(theme_bw())
#scale_fill_discrete <- function(palname = "Set1", ...) {
#    scale_fill_brewer(palette = palname, ...)
#}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

dev.off()
```


#However, how abundant are these asvs? And in how many samples were they found?

#Select our significantly variable ASVs:
```{r}
sigtab_names <- rownames_to_column(sigtab)
Variable_ASVs <- sigtab_names$rowname
```

#get only significantly variable ASVs, relativized:
```{r}
ps16S_ASVs_RNA_variable <- prune_taxa(Variable_ASVs, ps16S_ASVs_RNA_relative)
ps16S_ASVs_RNA_variable
```

#How abundant are these?
```{r}
plot_bar(ps16S_ASVs_RNA_variable, x = "DO_ID", y = "Abundance", fill = "Phylum")
```

#They're mostly just found in T06 at depth; wonder if that won't be an issue for our interpretation?

#Where are the ectos, I guess is a question?
```{r}
ECM_jitter <- ggplot(ps16S_metadata, aes(x = Tree_Health, y = ectomycorrhizal)) +
  
  geom_boxplot(outlier.alpha = 0,
               aes(x = Tree_Health, y = ectomycorrhizal, fill = Tree_Health)) +
  
  geom_point(aes(x = Tree_Health, y = ectomycorrhizal, fill = Tree_Health),
             position = position_jitterdodge(dodge.width = .75, jitter.width = .5),
             alpha = 1) +
  
  scale_fill_manual(values=c("#009900", "#CC6600")) +
  
  ylim(0, NA) +
  
  theme_classic()

ECM_jitter
```

#That's good, by tree?
```{r}
ECM_jitter <- ggplot(ps16S_metadata, aes(x = Tree, y = ectomycorrhizal)) +
  
  geom_boxplot(outlier.alpha = 0,
               aes(x = Tree, y = ectomycorrhizal, fill = Tree_Health)) +
  
  geom_point(aes(x = Tree, y = ectomycorrhizal, fill = Tree),
             position = position_jitterdodge(dodge.width = .75, jitter.width = .5),
             alpha = 1) +
  
  ylim(0, NA) +
  
  theme_classic()

ECM_jitter
```

#hmm. Kind of weird then that we don't see really many ECM-associated bacteria in many of the T03 samples?

#Let's graph that by depth and tree health status: NOTE this isn't taking an average, just piling dintinct sample abundances on top of each other.
```{r}
pdf(file = "Results/ps16S_ASVs_RNA_ECM_2trees.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches

p1 <- plot_bar(ps16S_ASVs_RNA_variable, x = "core_bottom_depth_cm", y = "Abundance", fill = "Phylum")

#print(p1)

p2 <- p1 + facet_grid(cols = vars(Tree_Health))

print(p2)

dev.off()
```


#now subset ASVs with only positive correlations with %ECM:
```{r}
sigtab <- filter(sigtab, log2FoldChange > 0)
dim(sigtab)

write_csv(sigtab, "Results/DESeq2_16S_ASVs_RNA_ECM_positive_2trees.csv")
```

#Visualize
```{r}
library("ggplot2")
theme_set(theme_bw())
#scale_fill_discrete <- function(palname = "Set1", ...) {
#    scale_fill_brewer(palette = palname, ...)
#}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
```

#many significant asvs 

#export:
```{r}
pdf(file = "Results/DESeq2_16S_ASVs_RNA_ECM_positive_2trees.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches

library("ggplot2")
theme_set(theme_bw())
#scale_fill_discrete <- function(palname = "Set1", ...) {
#    scale_fill_brewer(palette = palname, ...)
#}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

dev.off()
```


#However, how abundant are these asvs? And in how many samples were they found?

#Select our significantly variable ASVs:
```{r}
sigtab_names <- rownames_to_column(sigtab)
Variable_ASVs <- sigtab_names$rowname
```

#get only significantly variable ASVs, relativized:
```{r}
ps16S_ASVs_RNA_variable <- prune_taxa(Variable_ASVs, ps16S_ASVs_RNA_relative)
ps16S_ASVs_RNA_variable
```

#How abundant are these?
```{r}
plot_bar(ps16S_ASVs_RNA_variable, x = "DO_ID", y = "Abundance", fill = "Phylum")
```

#Almost all in T06 at depth.

#Let's graph that by depth and tree health status:
```{r}
pdf(file = "Results/ps16S_ASVs_RNA_ECM_positive_2trees.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches

p1 <- plot_bar(ps16S_ASVs_RNA_variable, x = "DO_ID", y = "Abundance", fill = "Phylum")

#print(p1)

p2 <- p1 + facet_grid(cols = vars(Tree_Health))

print(p2)

dev.off()
```


#get the relative abundances of all ECM-associated 16S ASVs:
```{r}
sample_sums(ps16S_ASVs_RNA_variable)
```

#add to metadata:
```{r}
ps16S_metadata$ECM_associated_16S_percent <- sample_sums(ps16S_ASVs_RNA_variable)
```

#multiply by total bacterial PLFA:
```{r}
ps16S_metadata <- mutate(ps16S_metadata, ECM_associated_16S_ng_g = ECM_associated_16S_percent * PLFA_Total_Bacteria_ng_g)
```

#combine ECM ng/g and ECM-associated 16S ASV ng/g:
```{r}
ps16S_metadata <- mutate(ps16S_metadata, ECM_associated_biomass = ECM_associated_16S_ng_g + ectomycorrhizal)
```

#add soil_depth_cm variable (would calculate but it's defined to be 7.5 cm):
```{r}
ps16S_metadata$soil_depth_cm <- 7.5
```

#calculate g carbon per cm^3 of soil:
```{r}
ps16S_metadata <- mutate(ps16S_metadata, mass_C_g_per_cm3_soil = bulk_density_g_cm3 * carbon_percent)
```

#calculate kg carbon per m^3 of soil:
```{r}
ps16S_metadata <- mutate(ps16S_metadata, mass_C_kg_per_m2_soil = (mass_C_g_per_cm3_soil * soil_depth_cm * 100 * 100)/1000)
```

#calculate metric tons (Mg) carbon in each depth fraction per hectare:
```{r}
ps16S_metadata <- mutate(ps16S_metadata, mass_C_Mg_per_ha = (mass_C_kg_per_m2_soil * 10000 / 1000))
```

#repeat that for all-sample metadata:

#add soil_depth_cm variable (would calculate but it's defined to be 7.5 cm):
```{r}
metadata$soil_depth_cm <- 7.5
```

#calculate g carbon per cm^3 of soil:
```{r}
metadata <- mutate(metadata, mass_C_g_per_cm3_soil = bulk_density_g_cm3 * carbon_percent)
```

#calculate kg carbon per m^3 of soil:
```{r}
metadata <- mutate(metadata, mass_C_kg_per_m2_soil = (mass_C_g_per_cm3_soil * soil_depth_cm * 100 * 100)/1000)
```

#calculate metric tons (Mg) carbon in each depth fraction per hectare:
```{r}
metadata <- mutate(metadata, mass_C_Mg_per_ha = (mass_C_kg_per_m2_soil * 10000 / 1000))
```

#   graph 1: stored C by Depth, Tree Health:
```{r}
C_tons_ha <- ggplot(metadata, aes(x = core_bottom_depth_cm,
                                                            y = mass_C_Mg_per_ha,
                                                            color = Tree_Health)) +
  geom_point() +
  
  geom_smooth(method = 'lm', formula = 'y ~ log(x)' ) +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  theme_classic()

C_tons_ha
```

#try a different smoothing method?
```{r}
#C_tons_ha <- ggplot(metadata, aes(x = core_bottom_depth_cm,
#                                                            y = mass_C_Mg_per_ha,
#                                                            color = Tree_Health)) +
 # geom_point() +
  
#  geom_smooth(method = "nls", formula = y ~ a * x^b, 
#                se = FALSE, method.args = list(start = c(a = 1, b = 1))) +
  
#  ylim(0, NA) +
  
#  xlim(0, NA) +
  
#  theme_classic()

#C_tons_ha
```

#failed, worth a shot, cou;d debug if we need to, see the GDM analysis fro what this ought to look like.

#export:
```{r}
pdf(file = "Results/C_Mg_by_depth_tree_health.pdf",   # The directory you want to save the file in
    width = 7, # The width of the plot in inches
    height = 5) # The height of the plot in inches

C_tons_ha

dev.off()
```

#   graph 2: C:N by Depth, Tree Health:
```{r}
C_N <- ggplot(metadata, aes(x = core_bottom_depth_cm,
                                                            y = C_per_N_Ratio,
                                                            color = Tree_Health)) +
  geom_point() +
  
  geom_smooth(method = 'lm') +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  theme_classic()

C_N
```


#failed, worth a shot

#export:
```{r}
pdf(file = "Results/C_N_Ratio_by_depth_tree_health.pdf",   # The directory you want to save the file in
    width = 7, # The width of the plot in inches
    height = 5) # The height of the plot in inches

C_N

dev.off()
```


#   graph 3: stored C by ECM + ECM-associated 16S ng/g:
#   note that this is a reduced dataset: RNA-based fungal communities only available for 21/36 samples.
```{r}
C_tons_ha <- ggplot(ps16S_metadata, aes(x = ECM_associated_biomass,
                                                            y = mass_C_Mg_per_ha)) +
  geom_point() +
  
  geom_smooth(method = 'lm') +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  theme_classic()

C_tons_ha
```


#export:
```{r}
pdf(file = "Results/C_Mg_by_ECM.pdf",   # The directory you want to save the file in
    width = 7, # The width of the plot in inches
    height = 5) # The height of the plot in inches

C_tons_ha

dev.off()
```

#   graph 4: C:N by ECM + ECM-associated 16S ng/g:
#   note that this is a reduced dataset: RNA-based fungal communities only available for 21/36 samples.
```{r}
C_N <- ggplot(ps16S_metadata, aes(x = ECM_associated_biomass,
                                                            y = C_per_N_Ratio)) +
  geom_point() +
  
  geom_smooth(method = 'lm') +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  theme_classic()

C_N
```


#export:
```{r}
pdf(file = "Results/C_N_Ratio_by_ECM.pdf",   # The directory you want to save the file in
    width = 7, # The width of the plot in inches
    height = 5) # The height of the plot in inches

C_N

dev.off()
```

#export complete metadata for Anna:
```{r}
write_csv(ps16S_metadata, "Results/metadata_with_ECMasssociated.csv")
```

