---
title: "DOMC_Fig04_2024_11_17"
author: "Glade Bogar"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load required packages:
```{r}
library("tidyverse")
library("cowplot")
```

#load required data:
```{r}
C_metadata <- read.csv("../../Analysis/Carbon/Density_Corrected_C/Results/per_sample_metadata_Ccorrected_mip.csv")

ecm_metadata = read.csv("../../Analysis/Microbial Communities/Fungal-Bacterial Interactions/DeSeq2 Indicator Species Analysis/Results/metadata_with_ECMasssociated3trees_mip.csv")
```


```{r}
# set common features of graphs

CN_lab = "C:N"
C_lab = "Mg C / hectare"
depth_lab = "Depth (cm)"
ecm_lab = "ECM-assoicated biomass (ng lipid/g soil)"
fontsize = 7
```


#   graph 1: stored C by Depth, Tree Health:
```{r}
C_by_depth <- ggplot(C_metadata, aes(x = core_bottom_depth_cm,
                                                            y = mass_C_Mg_per_ha,
                                                            color = Tree_Health)) +
  geom_point() +
  
  geom_smooth(method = 'lm', formula = 'y ~ log(x)' ) +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  scale_color_manual(name = "Tree mortality", values = c("darkblue", "deeppink")) +
  
  theme_classic(base_size = fontsize)

C_by_depth
leg = get_legend(C_by_depth)
C_by_depth <- ggplot(C_metadata, aes(x = core_bottom_depth_cm,
                                                            y = mass_C_Mg_per_ha,
                                                            color = Tree_Health)) +
  geom_point() +
  
  geom_smooth(method = 'lm', formula = 'y ~ log(x)' ) +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  scale_color_manual(name = "Tree mortality", values = c("darkblue", "deeppink")) +
  
  theme_classic(base_size = fontsize) + 
  
  theme(legend.position = "none") +
  
  xlab(depth_lab) +
  
  ylab(C_lab)

C_by_depth
```



#   graph 2: C:N by Depth, Tree Health:
```{r}
C_N_by_depth <- ggplot(C_metadata, aes(x = core_bottom_depth_cm,
                                                            y = C_per_N_Ratio,
                                                            color = Tree_Health)) +
  geom_point() +
  
  geom_smooth(method = 'lm') +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  scale_color_manual(name = "Tree mortality", values = c("darkblue", "deeppink")) +
  
  theme_classic(base_size = fontsize) + 
  
  theme(legend.position = "none") +
  
  xlab(depth_lab) +
  ylab(CN_lab)

C_N_by_depth
```



#   graph 3: stored C by ECM + ECM-associated 16S ng/g:
#   note that this is a reduced dataset: RNA-based fungal communities only available for 21/36 samples.
```{r}
C_by_ecm <- ggplot(ecm_metadata, aes(x = ECM_associated_biomass,
                                                            y = mass_C_Mg_per_ha)) +
  geom_point() +
  
  geom_smooth(method = 'lm') +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  theme_classic(base_size = fontsize) + 
  
  xlab(ecm_lab) +
  ylab(C_lab)

C_by_ecm
```


#   graph 4: C:N by ECM + ECM-associated 16S ng/g:
#   note that this is a reduced dataset: RNA-based fungal communities only available for 21/36 samples.
```{r}
C_N_by_ecm <- ggplot(ecm_metadata, aes(x = ECM_associated_biomass,
                                                            y = C_per_N_Ratio)) +
  geom_point() +
  
  geom_smooth(method = 'lm') +
  
  ylim(0, NA) +
  
  xlim(0, NA) +
  
  theme_classic(base_size = fontsize) + 
  
  xlab(ecm_lab) +
  ylab(CN_lab)

C_N_by_ecm
```
# Compile figure with Cowplot
```{r}

# First row
row1 <- plot_grid(C_N_by_depth, C_by_depth, labels = c("A", "B"), label_y = 0.98)

# Second row
row2 <- plot_grid(C_N_by_ecm, C_by_ecm, labels = c("C", "D"), label_y = 0.98)

# Stack rows vertically
combined_plot <- plot_grid(row1, row2, ncol = 1)

fig5 = ggdraw(combined_plot) + draw_grob(leg, hjust = -0.4, vjust = -0.405)

fig5
```



```{r}
# create the jpg figure file in this results folder and the reports folder

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 140
hw_ratio = 5/7 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

jpeg("Results/Figure5.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig5
dev.off()

jpeg("../../Reports/Figure5.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig5
dev.off()
```

```{r}
# create the png figure file in this results folder and the reports folder

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 140
hw_ratio = 5/7 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

png("Results/Figure5.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig5
dev.off()

png("../../Reports/Figure5.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig5
dev.off()


```

