---
title: "Calculate NMDSs"
author: "Anna Goodman"
date: "2025-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Perform ordination

#   16S ASVs, DNA - only, with fitted Factor Analysis sample metadata vectors; color by trees (custom; living warm tone, dead cold tone), opacity by depth (outline original color)
#   ITS ASVs, DNA - only, with fitted Factor Analysis sample metadata vectors
#   ITS Niche, DNA - only, with fitted Factor Analysis sample metadata vectors
#   16S Functions, DNA - only, with fitted Factor Analysis sample metadata vectors

#   16S ASVs, DNA - RNA distances (same aesthetics) - vector % RNA-based ECM
#   ITS ASVs, DNA - RNA distances (same aesthetics) - vector % RNA-based ECM
#   ITS Niche, DNA - RNA distances (same aesthetics) - vector % RNA-based ECM
#   16S Functions, DNA - RNA distances (same aesthetics) - vector % RNA-based ECM

### load required packages:
```{r}
library("phyloseq")
library("vegan")
library("tidyverse")
library("phyloseq")
```

#load pre-filtered, relativized phyloseq objects:
```{r}
# 16S ASVs
## with data from both nucleic acids
ps16S_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_mip.rds")
## DNA only
ps16S_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_DNA_mip.rds")
## RNA only
ps16S_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_RNA_mip.rds")


# ITS ASVs
## with data from both nucleic acids
psITS_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_mip.rds")
## DNA only
psITS_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_DNA_mip.rds")
## RNA only
psITS_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_RNA_mip.rds")

# 16S traits
## with data from both nucleic acids
ps16S_traits_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_mip.rds")
## DNA only
ps16S_traits_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_DNA_mip.rds")
## RNA only
ps16S_traits_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_RNA_mip.rds")


# ITS traits
## with data from both nucleic acids
psITS_traits_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_mip.rds")
## DNA only
psITS_traits_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_DNA_mip.rds")
## RNA only
psITS_traits_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_RNA_mip.rds")
```

# Write function to preform NMDS ordination with common settings
```{r}

NMDS_from_ps_relative = function (ps_relative) {
  set.seed(4721)
  NMDS <- ordinate(ps_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
return(NMDS)
}
```

# Make 16S ASV NMDSs
```{r}
NMDS_ps16S_relative = NMDS_from_ps_relative(ps16S_relative)
NMDS_ps16S_relative_DNA = NMDS_from_ps_relative(ps16S_relative_DNA)
NMDS_ps16S_relative_RNA = NMDS_from_ps_relative(ps16S_relative_RNA)

# Save NMDSs as rds files
write_rds(NMDS_ps16S_relative, "Results/NMDS_ps16S_relative_mip.rds")
write_rds(NMDS_ps16S_relative_DNA, "Results/NMDS_ps16S_relative_DNA_mip.rds")
write_rds(NMDS_ps16S_relative_RNA, "Results/NMDS_ps16S_relative_RNA_mip.rds")
```

# Make ITS ASV NMDSs
```{r}
NMDS_psITS_relative = NMDS_from_ps_relative(psITS_relative)
NMDS_psITS_relative_DNA = NMDS_from_ps_relative(psITS_relative_DNA)
NMDS_psITS_relative_RNA = NMDS_from_ps_relative(psITS_relative_RNA)

# Save NMDSs as rds files
write_rds(NMDS_psITS_relative, "Results/NMDS_psITS_relative_mip.rds")
write_rds(NMDS_psITS_relative_DNA, "Results/NMDS_psITS_relative_DNA_mip.rds")
write_rds(NMDS_psITS_relative_RNA, "Results/NMDS_psITS_relative_RNA_mip.rds")
```

# Make 16S traits NMDSs
```{r}
NMDS_ps16S_traits_relative = NMDS_from_ps_relative(ps16S_traits_relative)
NMDS_ps16S_traits_relative_DNA = NMDS_from_ps_relative(ps16S_traits_relative_DNA)
NMDS_ps16S_traits_relative_RNA = NMDS_from_ps_relative(ps16S_traits_relative_RNA)

# Save NMDSs as rds files
write_rds(NMDS_ps16S_traits_relative, "Results/NMDS_ps16S_traits_relative_mip.rds")
write_rds(NMDS_ps16S_traits_relative_DNA, "Results/NMDS_ps16S_traits_relative_DNA_mip.rds")
write_rds(NMDS_ps16S_traits_relative_RNA, "Results/NMDS_ps16S_traits_relative_RNA_mip.rds")
```

# Make ITS trait NMDSs
```{r}
NMDS_psITS_traits_relative = NMDS_from_ps_relative(psITS_traits_relative)
NMDS_psITS_traits_relative_DNA = NMDS_from_ps_relative(psITS_traits_relative_DNA)
NMDS_psITS_traits_relative_RNA = NMDS_from_ps_relative(psITS_traits_relative_RNA)

# Save NMDSs as rds files
write_rds(NMDS_psITS_traits_relative, "Results/NMDS_psITS_traits_relative_mip.rds")
write_rds(NMDS_psITS_traits_relative_DNA, "Results/NMDS_psITS_traits_relative_DNA_mip.rds")
write_rds(NMDS_psITS_traits_relative_RNA, "Results/NMDS_psITS_traits_relative_RNA_mip.rds")
```


