---
title: "Figure_02_AG_11_April_2025"
author: "Anna Goodman"
date: "2025-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```


------------------------------------------------------------------------------
#FIGURE 2
------------------------------------------------------------------------------
# set plot characteristics
```{r}
#define colors
factor_levels <- unique(ordination_scores$Tree)
# colors <- c("cyan", "blue", "azure4", "red", "hotpink", "darkorange")
# colors from hcl.colors("Batlow", n = 8), omitting two colors in the center of the scale

colors <- c("#201158", "#005260", "#007C46", "#D99C4C", "#FFAAB7", "#FFCEF9")

names(colors) <- factor_levels

#define point size ranges
commonrange = c(4, 1.5)
outlinerange = c(4.5, 2)

#define line weights
commonweight = 0.5

#define font size
fontsize = 7
```


#DNA to RNA NMDSs

#16S ASVs

#Check which samples are present in DNA dataset but not RNA:
```{r}
ps16S_DNA <- subset_samples(ps16S_relative, nucleic_acid == "DNA")
ps16S_RNA <- subset_samples(ps16S_relative, nucleic_acid == "RNA")

sample_names(ps16S_DNA) <- sample_data(ps16S_DNA)$DO_ID

sample_names(ps16S_RNA) <- sample_data(ps16S_RNA)$DO_ID

setdiff(sample_names(ps16S_DNA), sample_names(ps16S_RNA))
```

#Are any samples present in RNA but not DNA?
```{r}
setdiff(sample_names(ps16S_RNA), sample_names(ps16S_DNA))
```

#Remove these samples, leaving paired RNA/DNA only:
```{r}
ps16S_DNA
ps16S_DNA <- subset_samples(ps16S_DNA, !(DO_ID  %in% c("DO_T05_15-22.5", "DO_T05_30-37.5", "DO_T06_7.5-15", "DO_T06_37.5-45")))
ps16S_DNA

ps16S_RNA
ps16S_RNA <- subset_samples(ps16S_RNA, DO_ID  != "DO_T02_37.5-45")
ps16S_RNA
```

#re-name, re-combine:
```{r}
sample_data(ps16S_DNA)$DO_Name <- str_c(sample_data(ps16S_DNA)$DO_ID, sample_data(ps16S_DNA)$nucleic_acid, sep = "_")
sample_names(ps16S_DNA) <- sample_data(ps16S_DNA)$DO_Name

sample_data(ps16S_RNA)$DO_Name <- str_c(sample_data(ps16S_RNA)$DO_ID, sample_data(ps16S_RNA)$nucleic_acid, sep = "_")
sample_names(ps16S_RNA) <- sample_data(ps16S_RNA)$DO_Name

ps16S_relative <- merge_phyloseq(ps16S_DNA, ps16S_RNA)
ps16S_relative
ps16S_relative <- prune_taxa(taxa_sums(ps16S_relative) > 0, ps16S_relative)
ps16S_relative
```

#re-calculate distance matrix:
```{r}
ps16S_relative_df <- as(sample_data(ps16S_relative),"data.frame")
ps16S_relative_dist_matrix <- phyloseq::distance(ps16S_relative, "bray")
```


#calculate NMDS only with paired samples
```{r}
NMDS_ps16S_relative <- ordinate(ps16S_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
```

```{r}
# organize data for plotting
# Create a data frame for plotting
NMDS_ps16S_relative_points <- data.frame(NMDS_ps16S_relative$points)
NMDS_ps16S_relative_points <- rownames_to_column(NMDS_ps16S_relative_points, var = "DO_Name")
NMDS_ps16S_relative_points <- left_join(NMDS_ps16S_relative_points, ps16S_relative_df, by = "DO_Name")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_ps16S_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

#define colors
#factor_levels <- unique(ordination_scores$Tree)
# colors from hcl.colors("Batlow", n = 8), omitting two colors in the center of the scale

#names(colors) <- factor_levels

for (x in (1:length(NMDS_ps16S_relative_points$Tree))) {
  newString = paste(NMDS_ps16S_relative_points$Tree[x], "-", NMDS_ps16S_relative_points$Tree_Health[x])
  NMDS_ps16S_relative_points$Tree[x] = newString
}

NMDS_ps16S_relative_points$Tree <- factor(NMDS_ps16S_relative_points$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))

#define colors
factor_levels <- unique(NMDS_ps16S_relative_points$Tree)
names(colors) <- factor(factor_levels, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))
```

#plot the correct way:

```{r}

factor_levels <- unique(ordination_scores$Tree)
# colors <- c("cyan", "blue", "azure4", "red", "hotpink", "darkorange")
# colors from hcl.colors("Batlow", n = 8), omitting two colors in the center of the scale

colors <- c("#201158", "#005260", "#007C46", "#D99C4C", "#FFAAB7", "#FFCEF9")

names(colors) <- factor_levels

#plot

ps16S_ASV_DNAvsRNA_plot <- ggplot(NMDS_ps16S_relative_points, aes(x = MDS1,
                 y = MDS2,
                 size = core_bottom_depth_cm,
                 shape = nucleic_acid)) +
  
   # geom_point(color = "gray", size = 4.5) +

  
  geom_point(color = "black", stroke = 0.5, fill = NMDS_ps16S_relative_points$Tree) +
  
  scale_size_continuous(range = commonrange) +
  scale_shape_manual(values = c(21, 24)) +

 scale_fill_manual(values = colors) +  # Custom colors for each factor level
  # scale_color_manual(values = colors) +
  
  geom_segment(inherit.aes = FALSE, data = nucleic_acid_distance,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
  
  labs(title = "16S ASVs",
       x = "",
       y = "") +
  theme_classic(base_size = fontsize)

ps16S_ASV_DNAvsRNA_plot
```


#ITS ASVs

#Check which samples are present in DNA dataset but not RNA:
```{r}
psITS_DNA <- subset_samples(psITS_relative, nucleic_acid == "DNA")
psITS_RNA <- subset_samples(psITS_relative, nucleic_acid == "RNA")

sample_names(psITS_DNA) <- sample_data(psITS_DNA)$DO_ID

sample_names(psITS_RNA) <- sample_data(psITS_RNA)$DO_ID

setdiff(sample_names(psITS_DNA), sample_names(psITS_RNA))
```

#Are any samples present in RNA but not DNA?
```{r}
setdiff(sample_names(psITS_RNA), sample_names(psITS_DNA))
```

#Remove these samples, leaving paired RNA/DNA only:
```{r}
psITS_DNA
psITS_DNA <- subset_samples(psITS_DNA, !(DO_ID  %in% c("DO_T03_37.5-45",
                                                       "DO_T04_30-37.5",
                                                       "DO_T02_7.5-15",
                                                       "DO_T04_15-22.5",
                                                       "DO_T03_30-37.5",
                                                       "DO_T05_22.5-30",
                                                       "DO_T05_30-37.5",
                                                       "DO_T06_7.5-15",
                                                       "DO_T06_37.5-45")))
psITS_DNA

#psITS_RNA
#psITS_RNA <- subset_samples(psITS_RNA, DO_ID  != "")
psITS_RNA
```

#re-name, re-combine:
```{r}
sample_data(psITS_DNA)$DO_Name <- str_c(sample_data(psITS_DNA)$DO_ID, sample_data(psITS_DNA)$nucleic_acid, sep = "_")
sample_names(psITS_DNA) <- sample_data(psITS_DNA)$DO_Name

sample_data(psITS_RNA)$DO_Name <- str_c(sample_data(psITS_RNA)$DO_ID, sample_data(psITS_RNA)$nucleic_acid, sep = "_")
sample_names(psITS_RNA) <- sample_data(psITS_RNA)$DO_Name

psITS_relative <- merge_phyloseq(psITS_DNA, psITS_RNA)
psITS_relative
psITS_relative <- prune_taxa(taxa_sums(psITS_relative) > 0, psITS_relative)
psITS_relative
```

#re-calculate distance matrix:
```{r}
psITS_relative_df <- as(sample_data(psITS_relative),"data.frame")
psITS_relative_dist_matrix <- phyloseq::distance(psITS_relative, "bray")
```


#calculate NMDS only with paired samples
```{r}
NMDS_psITS_relative <- ordinate(psITS_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
```

#plot the correct way:
## pre- plot data prep
```{r} 
# Create a data frame for plotting
NMDS_psITS_relative_points <- data.frame(NMDS_psITS_relative$points)
NMDS_psITS_relative_points <- rownames_to_column(NMDS_psITS_relative_points, var = "DO_Name")
NMDS_psITS_relative_points <- left_join(NMDS_psITS_relative_points, psITS_relative_df, by = "DO_Name")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_psITS_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

```

```{r}
# Create a data frame for plotting
NMDS_psITS_relative_points <- data.frame(NMDS_psITS_relative$points)
NMDS_psITS_relative_points <- rownames_to_column(NMDS_psITS_relative_points, var = "DO_Name")
NMDS_psITS_relative_points <- left_join(NMDS_psITS_relative_points, psITS_relative_df, by = "DO_Name")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_psITS_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

#define colors
factor_levels <- unique(ordination_scores$Tree)
#colors <- c("cyan", "blue", "azure4", "red", "hotpink", "darkorange")
names(colors) <- factor_levels

#plot

psITS_ASV_DNAvsRNA_plot <- ggplot(NMDS_psITS_relative_points) +
  
      geom_point(data = NMDS_psITS_relative_points,
             aes(x = MDS1,
                 y = MDS2,
                 size = core_bottom_depth_cm,
                 shape = nucleic_acid,
                 color = "black")) +
    scale_size_continuous(range = outlinerange) +
  
  geom_point(data = NMDS_psITS_relative_points,
             aes(x = MDS1,
                 y = MDS2,
                 size = core_bottom_depth_cm,
                 shape = nucleic_acid,
                 color = Tree)) +
  scale_size_continuous(range = commonrange) +
  


  
  scale_color_manual(values = colors) +  # Custom colors for each tree
  
  geom_segment(data = nucleic_acid_distance,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
  
  labs(title = "ITS ASVs",
       x = "",
       y = "") +
  theme_classic(base_size = fontsize)

psITS_ASV_DNAvsRNA_plot
```



#ITS traits

#Check which samples are present in DNA dataset but not RNA:
```{r}
psITS_traits_DNA <- subset_samples(psITS_traits_relative, nucleic_acid == "DNA")
psITS_traits_RNA <- subset_samples(psITS_traits_relative, nucleic_acid == "RNA")

sample_names(psITS_traits_DNA) <- sample_data(psITS_traits_DNA)$DO_ID

sample_names(psITS_traits_RNA) <- sample_data(psITS_traits_RNA)$DO_ID

setdiff(sample_names(psITS_traits_DNA), sample_names(psITS_traits_RNA))
```

#Are any samples present in RNA but not DNA?
```{r}
setdiff(sample_names(psITS_traits_RNA), sample_names(psITS_traits_DNA))
```

#Remove these samples, leaving paired RNA/DNA only:
```{r}
psITS_traits_DNA
psITS_traits_DNA <- subset_samples(psITS_traits_DNA, !(DO_ID  %in% c("DO_T03_37.5-45",
                                                       "DO_T04_30-37.5",
                                                       "DO_T02_7.5-15",
                                                       "DO_T04_15-22.5",
                                                       "DO_T03_30-37.5",
                                                       "DO_T05_22.5-30",
                                                       "DO_T05_30-37.5",
                                                       "DO_T06_7.5-15",
                                                       "DO_T06_37.5-45")))
psITS_traits_DNA

#psITS_traits_RNA
#psITS_traits_RNA <- subset_samples(psITS_traits_RNA, DO_ID  != "")
psITS_traits_RNA
```

#re-name, re-combine:
```{r}
sample_data(psITS_traits_DNA)$DO_Name <- str_c(sample_data(psITS_traits_DNA)$DO_ID, sample_data(psITS_traits_DNA)$nucleic_acid, sep = "_")
sample_names(psITS_traits_DNA) <- sample_data(psITS_traits_DNA)$DO_Name

sample_data(psITS_traits_RNA)$DO_Name <- str_c(sample_data(psITS_traits_RNA)$DO_ID, sample_data(psITS_traits_RNA)$nucleic_acid, sep = "_")
sample_names(psITS_traits_RNA) <- sample_data(psITS_traits_RNA)$DO_Name

psITS_traits_relative <- merge_phyloseq(psITS_traits_DNA, psITS_traits_RNA)
psITS_traits_relative
psITS_traits_relative <- prune_taxa(taxa_sums(psITS_traits_relative) > 0, psITS_traits_relative)
psITS_traits_relative
```

#re-calculate distance matrix:
```{r}
psITS_traits_relative_df <- as(sample_data(psITS_traits_relative),"data.frame")
psITS_traits_relative_dist_matrix <- phyloseq::distance(psITS_traits_relative, "bray")
```


#calculate NMDS only with paired samples
```{r}
NMDS_psITS_traits_relative <- ordinate(psITS_traits_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
```

```{r}
# Create a data frame for plotting
NMDS_psITS_traits_relative_points <- data.frame(NMDS_psITS_traits_relative$points)
NMDS_psITS_traits_relative_points <- rownames_to_column(NMDS_psITS_traits_relative_points, var = "DO_Name")
NMDS_psITS_traits_relative_points <- left_join(NMDS_psITS_traits_relative_points, psITS_traits_relative_df, by = "DO_Name")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_psITS_traits_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

# add tree health to tree identity strings

for (x in (1:length(NMDS_psITS_traits_relative_points$Tree))) {
  newString = paste(NMDS_psITS_traits_relative_points$Tree[x], "-", NMDS_psITS_traits_relative_points$Tree_Health[x])
  NMDS_psITS_traits_relative_points$Tree[x] = newString
}

print(NMDS_psITS_traits_relative_points$Tree)

NMDS_psITS_traits_relative_points$Tree <- factor(NMDS_psITS_traits_relative_points$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))

#define colors
factor_levels <- unique(NMDS_psITS_traits_relative_points$Tree)
#colors <- c("cyan3", "blue", "azure4", "red", "hotpink", "darkorange")
names(colors) <- factor_levels
```


#plot the correct way:
```{r}


#plot

psITS_traits_ASV_DNAvsRNA_plot <- ggplot(NMDS_psITS_traits_relative_points) +
  
  geom_point(data = NMDS_psITS_traits_relative_points,
             aes(x = MDS1,
                 y = MDS2,
                 size = core_bottom_depth_cm,
                 shape = nucleic_acid,
                 color = Tree)) +
  
  scale_color_manual(values = colors) +  # Custom colors for each factor level
  
  geom_segment(data = nucleic_acid_distance,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
  
  scale_size_continuous(range = commonrange) +
  
   labs(title = "ITS traits",
       x = "",
       y = "",
       size = "Core bottom depth (cm)",
                 shape = "Nucleic Acid",
                 color = "Tree identity") +
  theme_classic(base_size = fontsize)

psITS_traits_ASV_DNAvsRNA_plot
```


#16S traits

#Check which samples are present in DNA dataset but not RNA:
```{r}
ps16S_traits_DNA <- subset_samples(ps16S_traits_relative, nucleic_acid == "DNA")
ps16S_traits_RNA <- subset_samples(ps16S_traits_relative, nucleic_acid == "RNA")

sample_names(ps16S_traits_DNA) <- sample_data(ps16S_traits_DNA)$DO_ID

sample_names(ps16S_traits_RNA) <- sample_data(ps16S_traits_RNA)$DO_ID

setdiff(sample_names(ps16S_traits_DNA), sample_names(ps16S_traits_RNA))
```

#Are any samples present in RNA but not DNA?
```{r}
setdiff(sample_names(ps16S_traits_RNA), sample_names(ps16S_traits_DNA))
```

#Remove these samples, leaving paired RNA/DNA only:
```{r}
ps16S_traits_DNA
ps16S_traits_DNA <- subset_samples(ps16S_traits_DNA, !(DO_ID  %in% c("DO_T05_15-22.5",
                                                                     "DO_T05_30-37.5",
                                                                     "DO_T06_7.5-15",
                                                                     "DO_T06_37.5-45")))
ps16S_traits_DNA

ps16S_traits_RNA
ps16S_traits_RNA <- subset_samples(ps16S_traits_RNA, DO_ID  != "DO_T02_37.5-45")
ps16S_traits_RNA
```

#re-name, re-combine:
```{r}
sample_data(ps16S_traits_DNA)$DO_Name <- str_c(sample_data(ps16S_traits_DNA)$DO_ID, sample_data(ps16S_traits_DNA)$nucleic_acid, sep = "_")
sample_names(ps16S_traits_DNA) <- sample_data(ps16S_traits_DNA)$DO_Name

sample_data(ps16S_traits_RNA)$DO_Name <- str_c(sample_data(ps16S_traits_RNA)$DO_ID, sample_data(ps16S_traits_RNA)$nucleic_acid, sep = "_")
sample_names(ps16S_traits_RNA) <- sample_data(ps16S_traits_RNA)$DO_Name

ps16S_traits_relative <- merge_phyloseq(ps16S_traits_DNA, ps16S_traits_RNA)
ps16S_traits_relative
ps16S_traits_relative <- prune_taxa(taxa_sums(ps16S_traits_relative) > 0, ps16S_traits_relative)
ps16S_traits_relative
```

#re-calculate distance matrix:
```{r}
ps16S_traits_relative_df <- as(sample_data(ps16S_traits_relative),"data.frame")
ps16S_traits_relative_dist_matrix <- phyloseq::distance(ps16S_traits_relative, "bray")
```


#calculate NMDS only with paired samples
```{r}
NMDS_ps16S_traits_relative <- ordinate(ps16S_traits_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
```


#plot the correct way:
```{r}
# Create a data frame for plotting
NMDS_ps16S_traits_relative_points <- data.frame(NMDS_ps16S_traits_relative$points)
NMDS_ps16S_traits_relative_points <- rownames_to_column(NMDS_ps16S_traits_relative_points, var = "DO_Name")
NMDS_ps16S_traits_relative_points <- left_join(NMDS_ps16S_traits_relative_points, ps16S_traits_relative_df, by = "DO_Name")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_ps16S_traits_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

#define colors
factor_levels <- unique(ordination_scores$Tree)
# colors <- c("cyan", "blue", "azure4", "red", "hotpink", "darkorange")
names(colors) <- factor_levels

#plot

ps16S_traits_ASV_DNAvsRNA_plot <- ggplot(NMDS_ps16S_traits_relative_points) +
  
  geom_point(data = NMDS_ps16S_traits_relative_points,
             aes(x = MDS1,
                 y = MDS2,
                 size = core_bottom_depth_cm,
                 shape = nucleic_acid,
                 color = Tree)) +
  
  scale_fill_manual(values = colors) +  # Custom colors for each factor level
  
  geom_segment(data = nucleic_acid_distance,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
  
  scale_size_continuous(range = commonrange) +
  
  labs(title = "16S traits",
       x = "",
       y = "") +
  theme_classic(base_size = fontsize)

ps16S_traits_ASV_DNAvsRNA_plot
```



```{r}

legend = get_legend(psITS_traits_ASV_DNAvsRNA_plot)
ASV_16S = ps16S_ASV_DNAvsRNA_plot + theme(legend.position = "none") + xlab("")
ASV_ITS = psITS_ASV_DNAvsRNA_plot + theme(legend.position = "none")
trait_16S = ps16S_traits_ASV_DNAvsRNA_plot + theme(legend.position = "none") 
trait_ITS = psITS_traits_ASV_DNAvsRNA_plot + theme(legend.position = "none") 

all_no_legend = plot_grid(ASV_16S, ASV_ITS, trait_16S, trait_ITS,
                 nrow = 2, 
                 ncol = 2, align = 'vh',
                 rel_heights = c())
                 #labels = c("16S ASVs", "ITS ASVs", "16S traits", "ITS traits"))

labeled = all_no_legend +
  draw_text("NMDS2", x = 0.02, y = 0.45, hjust = "left", angle=90, size = fontsize) +
  draw_text("NMDS1", x = 0.55, y = 0.02, size = fontsize) # x value is 1/2 of the 2/3 of the horizontal space which is the panels

fig2 = plot_grid(labeled, legend, align = "h", rel_widths = c(3, 1))

#create common x and y labels

#fig2 = fig2_leg + draw_text("NMDS2", x = 0.025, y = 0., hjust = "left", angle=90) + draw_text("NMDS1", x = 0.4, y = 0.025, vjust = "bottom") # x value is 1/2 of the 2/3 of the horizontal space which is the panels

```

```{r}
fig2
```

```{r}
# print the figure

hw_ratio = 6/8
w_mm = 140
h_mm = w_mm*hw_ratio


ppi = 1000

tiff("C:/Users/goodm/OneDrive/Documents/Bogar Lab/Deep Oaks (DO)/SBB Submission/Figure2.tiff", width = w_mm, height = h_mm, units = "mm", res = ppi)
fig2
dev.off()
```

