---
title: 'Figure 2: DNA vs. RNA NMDS'
author: "Anna Goodman"
date: "2025-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load required packages
```{r}
library(tidyverse)
```


------------------------------------------------------------------------------
#FIGURE 2
------------------------------------------------------------------------------
#load pre-filtered, relativized, DNA-only phyloseq objects:
```{r}
# 16S ASVs
ps16S_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_mip.rds")

# ITS ASVs
psITS_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_mip.rds")

# 16S traits
ps16S_traits_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_mip.rds")

# ITS traits
psITS_traits_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_mip.rds")
```

# load in NMDSs - filtered for samples with community data from BOTH dna and rna
```{r}
# 16S ASVs
NMDS_ps16S_relative <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/ps16S_relative_NMDS_mip.rds")

# ITS ASVs
NMDS_psITS_relative <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/psITS_relative_NMDS_mip.rds")

# 16S traits
NMDS_ps16S_traits_relative <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/ps16S_traits_relative_NMDS_mip.rds")

# ITS traits
NMDS_psITS_traits_relative <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/psITS_traits_relative_NMDS_mip.rds")
```


```{r}
metadata = read.csv("../../TidyData/Metadata/Results/per_sample_metadata_mip.csv")
```


# set plot characteristics
```{r}
#define colors
factor_levels <- unique(metadata$Tree)
# colors <- c("cyan", "blue", "azure4", "red", "hotpink", "darkorange")
# colors from hcl.colors("Batlow", n = 8), omitting two colors in the center of the scale

colors <- c("#201158", "#005260", "#007C46", "#D99C4C", "#FFAAB7", "#FFCEF9")

names(colors) <- factor_levels

#define point size ranges
commonrange = c(4, 1.5)

#define line weights
commonweight = 0.5

#define font size
fontsize = 7
```

```{r}
filter_ps_for_na_comparison = function (ps_relative) {
  
      is_dna <- sample_data(ps_relative)$nucleic_acid == "DNA"
      is_rna <- sample_data(ps_relative)$nucleic_acid == "RNA"
      
      # Remove NA values just in case
      is_dna[is.na(is_dna)] <- FALSE
      is_rna[is.na(is_rna)] <- FALSE
      
      keep_DNA <- !(sample_data(ps_DNA)$DO_ID %in% dna_only)
      ps_DNA <- prune_samples(keep_DNA, ps_DNA)
      
      keep_RNA <- !(sample_data(ps_RNA)$DO_ID %in% rna_only)
      ps_RNA <- prune_samples(keep_RNA, ps_RNA)

      
      message("Samples in ps_DNA: ", nsamples(ps_DNA))
      message("Samples in ps_RNA: ", nsamples(ps_RNA))


    # Select samples are present in DNA dataset but not RNA, and vice versa:
      dna_only = setdiff(sample_names(ps_DNA), sample_names(ps_RNA))
      rna_only = setdiff(sample_names(ps_RNA), sample_names(ps_DNA))
      
    # Remove these samples, leaving paired RNA/DNA only:
      ps_DNA <- subset_samples(ps_DNA, !(DO_ID  %in% dna_only))
      ps_RNA <- subset_samples(ps_RNA, !(DO_ID  %in% rna_only))
  
    #re-name, re-combine:
      sample_data(ps_DNA)$DO_Name <- str_c(sample_data(ps_DNA)$DO_ID, sample_data(ps_DNA)$nucleic_acid, sep = "_")
      sample_names(ps_DNA) <- sample_data(ps_DNA)$DO_Name

      sample_data(ps_RNA)$DO_Name <- str_c(sample_data(ps_RNA)$DO_ID, sample_data(ps_RNA)$nucleic_acid, sep = "_")
      sample_names(ps_RNA) <- sample_data(ps_RNA)$DO_Name

      ps_relative <- merge_phyloseq(ps_DNA, ps_RNA)
      ps_relative <- prune_taxa(taxa_sums(ps_relative) > 0, ps_relative) #remove taxa that aren't in the dataset anymore (only appeared in samples which were not represented as both DNA and RNA)

return(ps_relative)
}
```

```{r}

make_nmds_for_na_comparison = function (ps_for_comparison) {

  NMDS <- ordinate(ps_for_comparison,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
return(NMDS)
}
```


```{r}
makepanel_fig2 = function (ps_for_comparison, nmds_for_na_comparison){
  ps_sampledata <- as(sample_data(ps_relative),"data.frame")
  
  df_points <- data.frame(nmds$points)
  df_points <- rownames_to_column(df_points, var = "DO_Name")
  df_points <- left_join(df_points, ps_sampledata, by = "DO_Name")
  
  # Prepare data for pairwise lines
  nucleic_acid_distance <- df_points %>%
    group_by(DO_ID) %>%
    summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
              MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
    ungroup()
  
  dist_frame <- left_join(nucleic_acid_distance, ps_sampledata, by = "DO_ID")
    
  panel = ggplot(df_points) +
          geom_point(data = df_points,
                aes(x = MDS1,
                y = MDS2,
                size = core_bottom_depth_cm,
                shape = nucleic_acid,
                color = Tree)) +
    
          #geom_point(shape = 21, color = "black", stroke = 0.5) + # pch=21: filled circle with black border
           # scale_shape_manual(values = c(21, 24)) +
          scale_color_manual(values = colors) +  # Custom colors for each factor level
          scale_fill_manual(values = colors) +   # Fill colors to match
          scale_color_manual(values = colors) +  # Custom colors for each factor level
    
          geom_segment(data = df_lines,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
           scale_size_continuous(range = commonrange) +
           labs(title = panel_name,
           x = "",
           y = "") +
           theme_classic(base_size = fontsize) +
          theme(legend.position = "none") 
                
return(panel)
}
```

```{r}
ps16S_comparison = filter_ps_for_na_comparison(ps16S_relative)
```




# prepare dataframes to plot
```{r}

ps_sampledata <- as(sample_data(ps16S_relative),"data.frame")

# Create a data frame for plotting
NMDS_ps16S_relative_points <- data.frame(NMDS_ps16S_relative$points)
NMDS_ps16S_relative_points <- rownames_to_column(NMDS_ps16S_relative_points, var = "DO_ID")
NMDS_ps16S_relative_points <- left_join(NMDS_ps16S_relative_points, ps_sampledata, by = "DO_ID")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_ps16S_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance_ps16S_ASVs <- left_join(nucleic_acid_distance, metadata, by = "DO_ID")
```





#ITS traits

##Check which samples are present in DNA dataset but not RNA:
```{r}
psITS_traits_DNA <- subset_samples(psITS_traits_relative, nucleic_acid == "DNA")
psITS_traits_RNA <- subset_samples(psITS_traits_relative, nucleic_acid == "RNA")

sample_names(psITS_traits_DNA) <- sample_data(psITS_traits_DNA)$DO_ID

sample_names(psITS_traits_RNA) <- sample_data(psITS_traits_RNA)$DO_ID

setdiff(sample_names(psITS_traits_DNA), sample_names(psITS_traits_RNA))
```

##Are any samples present in RNA but not DNA?
```{r}
setdiff(sample_names(psITS_traits_RNA), sample_names(psITS_traits_DNA))
```

##Remove these samples, leaving paired RNA/DNA only:
```{r}
psITS_traits_DNA
psITS_traits_DNA <- subset_samples(psITS_traits_DNA, !(DO_ID  %in% c("DO_T03_37.5-45",
                                                       "DO_T04_30-37.5",
                                                       "DO_T02_7.5-15",
                                                       "DO_T04_15-22.5",
                                                       "DO_T03_30-37.5",
                                                       "DO_T05_22.5-30",
                                                       "DO_T05_30-37.5",
                                                       "DO_T06_7.5-15",
                                                       "DO_T06_37.5-45")))
```

##re-name, re-combine:
```{r}
sample_data(psITS_traits_DNA)$DO_Name <- str_c(sample_data(psITS_traits_DNA)$DO_ID, sample_data(psITS_traits_DNA)$nucleic_acid, sep = "_")
sample_names(psITS_traits_DNA) <- sample_data(psITS_traits_DNA)$DO_Name

sample_data(psITS_traits_RNA)$DO_Name <- str_c(sample_data(psITS_traits_RNA)$DO_ID, sample_data(psITS_traits_RNA)$nucleic_acid, sep = "_")
sample_names(psITS_traits_RNA) <- sample_data(psITS_traits_RNA)$DO_Name

psITS_traits_relative <- merge_phyloseq(psITS_traits_DNA, psITS_traits_RNA)

psITS_traits_relative <- prune_taxa(taxa_sums(psITS_traits_relative) > 0, psITS_traits_relative)

```

##re-calculate distance matrix:
```{r}
psITS_traits_relative_df <- as(sample_data(psITS_traits_relative),"data.frame")
psITS_traits_relative_dist_matrix <- phyloseq::distance(psITS_traits_relative, "bray")
```


##calculate NMDS only with paired samples
```{r}
NMDS_psITS_traits_relative <- ordinate(psITS_traits_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
```
## prepare dataframes for plotting
```{r}
# Create a data frame for points
NMDS_psITS_traits_relative_points <- data.frame(NMDS_psITS_traits_relative$points)
NMDS_psITS_traits_relative_points <- rownames_to_column(NMDS_psITS_traits_relative_points, var = "DO_Name")
NMDS_psITS_traits_relative_points <- left_join(NMDS_psITS_traits_relative_points, psITS_traits_relative_df, by = "DO_Name")

# Prepare data frame for lines
nucleic_acid_distance <- NMDS_psITS_traits_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance_psITS_traits <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

```

### make the plot to mine the legend from
```{r}

# add tree health to tree identity strings

copyof_ITS_traits_points = NMDS_psITS_traits_relative_points

for (x in (1:length(copyof_ITS_traits_points$Tree))) {
  newString = paste(copyof_ITS_traits_points$Tree[x], "-", copyof_ITS_traits_points$Tree_Health[x])
  copyof_ITS_traits_points$Tree[x] = newString
}



copyof_ITS_traits_points$Tree <- factor(copyof_ITS_traits_points$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))

#define colors
factor_levels <- unique(copyof_ITS_traits_points$Tree)
#colors <- c("cyan3", "blue", "azure4", "red", "hotpink", "darkorange")
names(colors) <- factor_levels


#plot

legend_plot <- ggplot(copyof_ITS_traits_points) +
  
  geom_point(data = copyof_ITS_traits_points,
             aes(x = MDS1,
                 y = MDS2,
                 size = core_bottom_depth_cm,
                 shape = nucleic_acid,
                 color = Tree)) +
  
  scale_color_manual(values = colors) +  # Custom colors for each factor level
  
  geom_segment(data = nucleic_acid_distance_psITS_traits,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
  
  scale_size_continuous(range = commonrange) +
  
   labs(title = "ITS traits",
       x = "",
       y = "",
       size = "Core bottom depth (cm)",
                 shape = "Nucleic Acid",
                 color = "Tree identity") +
  theme_classic(base_size = fontsize)

legend_plot
```




#16S traits

##Check which samples are present in DNA dataset but not RNA:
```{r}
ps16S_traits_DNA <- subset_samples(ps16S_traits_relative, nucleic_acid == "DNA")
ps16S_traits_RNA <- subset_samples(ps16S_traits_relative, nucleic_acid == "RNA")

sample_names(ps16S_traits_DNA) <- sample_data(ps16S_traits_DNA)$DO_ID

sample_names(ps16S_traits_RNA) <- sample_data(ps16S_traits_RNA)$DO_ID

setdiff(sample_names(ps16S_traits_DNA), sample_names(ps16S_traits_RNA))
```

##Are any samples present in RNA but not DNA?
```{r}
setdiff(sample_names(ps16S_traits_RNA), sample_names(ps16S_traits_DNA))
```

##Remove these samples, leaving paired RNA/DNA only:
```{r}
ps16S_traits_DNA
ps16S_traits_DNA <- subset_samples(ps16S_traits_DNA, !(DO_ID  %in% c("DO_T05_15-22.5",
                                                                     "DO_T05_30-37.5",
                                                                     "DO_T06_7.5-15",
                                                                     "DO_T06_37.5-45")))

ps16S_traits_RNA <- subset_samples(ps16S_traits_RNA, DO_ID  != "DO_T02_37.5-45")

```

##re-name, re-combine:
```{r}
sample_data(ps16S_traits_DNA)$DO_Name <- str_c(sample_data(ps16S_traits_DNA)$DO_ID, sample_data(ps16S_traits_DNA)$nucleic_acid, sep = "_")
sample_names(ps16S_traits_DNA) <- sample_data(ps16S_traits_DNA)$DO_Name

sample_data(ps16S_traits_RNA)$DO_Name <- str_c(sample_data(ps16S_traits_RNA)$DO_ID, sample_data(ps16S_traits_RNA)$nucleic_acid, sep = "_")
sample_names(ps16S_traits_RNA) <- sample_data(ps16S_traits_RNA)$DO_Name

ps16S_traits_relative <- merge_phyloseq(ps16S_traits_DNA, ps16S_traits_RNA)
ps16S_traits_relative
ps16S_traits_relative <- prune_taxa(taxa_sums(ps16S_traits_relative) > 0, ps16S_traits_relative)
ps16S_traits_relative
```

##re-calculate distance matrix:
```{r}
ps16S_traits_relative_df <- as(sample_data(ps16S_traits_relative),"data.frame")
ps16S_traits_relative_dist_matrix <- phyloseq::distance(ps16S_traits_relative, "bray")
```


##calculate NMDS only with paired samples
```{r}
NMDS_ps16S_traits_relative <- ordinate(ps16S_traits_relative,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
```


## Prepare dataframes for plotting:
```{r}
# Create a data frame for plotting
NMDS_ps16S_traits_relative_points <- data.frame(NMDS_ps16S_traits_relative$points)
NMDS_ps16S_traits_relative_points <- rownames_to_column(NMDS_ps16S_traits_relative_points, var = "DO_Name")
NMDS_ps16S_traits_relative_points <- left_join(NMDS_ps16S_traits_relative_points, ps16S_traits_relative_df, by = "DO_Name")

# Prepare data for pairwise lines
nucleic_acid_distance <- NMDS_ps16S_traits_relative_points %>%
  group_by(DO_ID) %>%
  summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
            MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
  ungroup()

nucleic_acid_distance_ps16S_traits <- left_join(nucleic_acid_distance, sample_metadata, by = "DO_ID")

```


# write function to generate plots
```{r}
makePanel = function (panel_name, df_points, df_lines) {

  #define colors
factor_levels <- unique(ordination_scores$Tree)
# colors <- c("cyan", "blue", "azure4", "red", "hotpink", "darkorange")
# colors from hcl.colors("Batlow", n = 8), omitting two colors in the center of the scale

colors <- c("#201158", "#005260", "#007C46", "#D99C4C", "#FFAAB7", "#FFCEF9")

names(colors) <- factor_levels

#define point size ranges
commonrange = c(4, 1.5)

#define line weights
commonweight = 0.5

#define font size
fontsize = 7

  panel = ggplot(df_points) +
          geom_point(data = df_points,
                aes(x = MDS1,
                y = MDS2,
                size = core_bottom_depth_cm,
                shape = nucleic_acid,
                color = Tree)) +
    
      #geom_point(shape = 21, color = "black", stroke = 0.5) + # pch=21: filled circle with black border
   # scale_shape_manual(values = c(21, 24)) +
  scale_color_manual(values = colors) +  # Custom colors for each factor level
  scale_fill_manual(values = colors) +   # Fill colors to match
          scale_color_manual(values = colors) +  # Custom colors for each factor level
    
          geom_segment(data = df_lines,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               size = commonweight) +
           scale_size_continuous(range = commonrange) +
           labs(title = panel_name,
           x = "",
           y = "") +
           theme_classic(base_size = fontsize) +
          theme(legend.position = "none") 
                
return(panel)

}
```

## test function
```{r}

ASV_16S = makepanel_fig2("16S ASVs", NMDS_ps16S_relative_points, nucleic_acid_distance_ps16S_ASVs)

ASV_ITS = makepanel_fig2("ITS ASVs", NMDS_psITS_relative_points, nucleic_acid_distance_psITS_ASVs)


trait_16S = makepanel_fig2("16S traits", NMDS_ps16S_traits_relative_points, nucleic_acid_distance_ps16S_traits)

trait_ITS = makepanel_fig2("ITS traits", NMDS_psITS_traits_relative_points, nucleic_acid_distance_psITS_traits)
```

```{r}
legend = get_legend(legend_plot)
 

all_no_legend = plot_grid(ASV_16S, ASV_ITS, trait_16S, trait_ITS,
                 nrow = 2, 
                 ncol = 2, align = 'vh')
               #  rel_heights = c()
                 #labels = c("16S ASVs", "ITS ASVs", "16S traits", "ITS traits"))

labeled = all_no_legend +
  draw_text("NMDS2", x = 0.02, y = 0.45, hjust = "left", angle=90, size = fontsize) +
  draw_text("NMDS1", x = 0.55, y = 0.02, size = fontsize) # x value is 1/2 of the 2/3 of the horizontal space which is the panels

fig2 = plot_grid(labeled, legend, align = "h", rel_widths = c(3, 1))

#create common x and y labels

#fig2 = fig2_leg + draw_text("NMDS2", x = 0.025, y = 0., hjust = "left", angle=90) + draw_text("NMDS1", x = 0.4, y = 0.025, vjust = "bottom") # x value is 1/2 of the 2/3 of the horizontal space which is the panels

```

```{r}
fig2
```

```{r}
# create the figure file in this results folder

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 140
hw_ratio = 5/7 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

jpeg("Results/Figure2.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
jpeg("../../Reports/Figure2.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)


fig2
dev.off()
```

