---
title: "DO_Make_Metadata_for_phyloseq"
author: "Glade Bogar"
date: "2024-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Load required packages:
```{r}
library(tidyverse)
packageVersion('tidyverse')

library(phyloseq)
packageVersion('phyloseq')
```

#Load required data:
```{r}
ps16S <- read_rds("../../Analysis/Microbial Communities/Phyloseq_16S/Results/ps16S_remove_3_reads.rds")
psITS <- read_rds("../../Analysis/Microbial Communities/Phyloseq_ITS/Results/psITS_remove_8_reads_fungaltraits.rds")
sample_metadata <- read_csv("Results/per_sample_metadata_mip.csv")
amplification <- read_csv("../RNA-DNA Extractions/Results/barcodes_tidy_mip.csv")
extraction <- read_csv("../RNA-DNA Extractions/Results/nucleic_acid_extractions_for_library_prep_tidy_mip.csv")
```

#tidy the extraction dataset, which we're using only to link the sample IDs to the extraction IDs: 
```{r}
#filter out unused samples
extraction <- filter(extraction, Extraction_ID != "NA")
#select only relevant ID columns
extraction <- select(extraction, c(Extraction_ID, DO_ID, date_collected))
```


```{r}
"DO_ID" %in% colnames(extraction)
"DO_ID" %in% colnames(amplification)

head(phyloseq_metadata)
```


#link extraction ID to barcode pairings 
```{r}
# join by extraction ID
amplification_metadata <- full_join(amplification, extraction, by = "Extraction_ID")
```

#Join and separate 16S, ITS samples:
```{r}
phyloseq_metadata <- left_join(amplification_metadata, sample_metadata, by = "DO_ID")

phyloseq_metadata_16S <- filter(phyloseq_metadata, amplification == "PCR_16S")
phyloseq_metadata_ITS <- filter(phyloseq_metadata, amplification == "PCR_ITS") 
```

#Get the current phyloseq object sample names:
```{r}
ps16S_sample_names <- as.data.frame(sample_names(ps16S))
ps16S_sample_names <- rename(ps16S_sample_names, phyloseq_name = "sample_names(ps16S)")

psITS_sample_names <- as.data.frame(sample_names(psITS))
psITS_sample_names <- rename(psITS_sample_names, phyloseq_name = "sample_names(psITS)")
```

#Recreate these sample names as a column in metadata files:
```{r}
phyloseq_metadata_16S$phyloseq_name <- str_c("DO_",phyloseq_metadata_16S$REV_barcode,"_",phyloseq_metadata_16S$FWD_barcode,"_R1.fastq.gz")
phyloseq_metadata_ITS$phyloseq_name <- str_c("DO_",phyloseq_metadata_ITS$REV_barcode,"_",phyloseq_metadata_ITS$FWD_barcode,"_R1.fastq.gz")
```

#Left join to phyloseq names:
```{r}
phyloseq_metadata_16S <- left_join(ps16S_sample_names, phyloseq_metadata_16S)
phyloseq_metadata_ITS <- left_join(psITS_sample_names, phyloseq_metadata_ITS)
```

#Write phyloseq metadata:
```{r}
write_csv(phyloseq_metadata_16S, "Results/phyloseq_metadata_16S_mip.csv")
write_csv(phyloseq_metadata_ITS, "Results/phyloseq_metadata_ITS_mip.csv")
```


























