---
title: "Supplementary Figure 3"
author: "Anna Goodman"
date: "2024-12-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# load packages
```{r}
library(tidyverse)
library(ggpubr)
library(cowplot)
library(ggplot2)
```

# load data
```{r}
# load in the metadata with PLFA data for all 36 samples
metadata36 <- read.csv("../../TidyData/Metadata/Results/per_sample_metadata_mip.csv")
```

# refactor depth variable
```{r}
discrete.depth = metadata36 # make a copy dataframe of the original metadata

discrete.depth$sample_depth_cm <- factor(discrete.depth$sample_depth_cm, levels=c("0-7.5", "7.5-15", "15-22.5", "22.5-30", "30-37.5", "37.5-45")) # values in this column are characters so plots won't automatically arrange data in numerical order; set them in order of increasing depth (descending from surface)
```

# define universal plotting layers/modifiers so I can edit all panels at once
```{r}
boxplot = geom_boxplot(alpha = 0.075) # add boxplot layer, make fill mostly transparent

points = geom_jitter(position = position_jitterdodge(jitter.width = 0, dodge.width = 0.75)) # add raw datapoint layer. Don't actually jitter, but dodge so living and dead points are separated and overlay their boxplots

fillcolor = scale_fill_manual(name = "Tree mortality", values = c("darkblue", "deeppink")) # color the semi-transparent box fill, name legend

dotlinecolor = scale_color_manual(name = "Tree mortality", values = c("darkblue", "deeppink")) # color the points and lines, name legend consistently

background = theme_minimal() # visually simplify background

xaxislabel = xlab("Depth (cm)") #labs(x = "Depth (cm)", y = "Total microbial biomass (ng PLFA/g soil)") # formalize text on axis labels

font = theme(text = element_text(size = 7)) # set font size

margins = theme(plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 15)) # set margin widths: widen left margin of each plot so the panel labels can fit in top left. 5.5 is default pt size for margins in ggplot.

ymax = ylim(0, 2050) # for bac and fun panels only: set common y-axis limit to enable side-by-side comparison 

add_leg_box = theme( # optional: add to whichever panel the legend grob gets grabbed from so the final legend has a box around it 
    legend.background = element_rect(
      color = "black",
      linewidth = 0.5,
      linetype = "solid"
    )
  )
```

# make microbial biomass over depth panel
```{r}
mic = mic = ggplot(discrete.depth, aes(sample_depth_cm, PLFA_Total_Living_Microbial_Biomass_ng_g, fill = Tree_Health, color = Tree_Health)) +
  boxplot +
  points +
  fillcolor +
  dotlinecolor +
  background +
  xaxislabel +
  ylab("Total microbial biomass (ng PLFA/g soil)") +
  font +
  margins

mic
```

# make bacterial biomass over depth panel
```{r}
bac = ggplot(discrete.depth, aes(x = sample_depth_cm, y = PLFA_Total_Bacteria_ng_g, fill = Tree_Health, color = Tree_Health)) +  
  boxplot +
  points +
  fillcolor +
  dotlinecolor +
  background +
  xaxislabel +
  ylab("Total bacterial biomass (ng PLFA/g soil)") +
  font +
  margins +
  ymax

bac
```

# make fungal biomass over depth panel
```{r}
fun = ggplot(discrete.depth, aes(x = sample_depth_cm, y = PLFA_Total_Fungi_ng_g, fill = Tree_Health, color = Tree_Health)) +  
  boxplot +
  points +
  fillcolor +
  dotlinecolor +
  background +
  xaxislabel +
  ylab("Total fungal biomass (ng PLFA/g soil)") +
  font +
  margins + 
  ymax

fun
```

# make fungi:bacteria biomass ratio over depth panel
```{r}
discrete.depth = discrete.depth %>% 
  mutate(PLFA_fungi_to_bacteria_ratio = (PLFA_Total_Fungi_ng_g/PLFA_Total_Bacteria_ng_g)) # add column to discrete.depth metadata dataframe with computed fungi:bacteria ratio, based on PLFA biomass of fungi and bacteria

fb_ratio = ggplot(discrete.depth, aes(x = sample_depth_cm, y = PLFA_fungi_to_bacteria_ratio, fill = Tree_Health, color = Tree_Health)) +
  boxplot +
  points +
  fillcolor +
  dotlinecolor +
  background +
  xaxislabel +
  ylab("Fungi:bacteria biomass ratio") +
  font +
  margins

fb_ratio
```

#make null/blank panel
```{r}
blank = ggplot(discrete.depth, aes(x = , y = )) + theme_minimal()
blank
```

# combine boxplots as pannels of a single figure - fun:bac in own row
```{r}
supFig3_2rows_noleg = ggpubr::ggarrange(
  mic, bac, fun, blank, fb_ratio, blank, # list of boxplots to be panels
  labels = c("A", "B", "C", "", "D", ""), # labels for each panel
  font.label = list(size = 10, color = "black", face = "bold", family = NULL),
  widths = c(4,4,4),
  heights = c(3,3),
  common.legend = TRUE, # common legend for all panels
  legend = "none", # legend displayed above all panels
  align = "h", # align panels horizontally and vertically
  nrow = 2, ncol = 3 # arrange panels: 1 row for plfa data, 1 row for ratio
)

supFig3_2rows_noleg


```

```{r}
leg = get_legend(fun) # legend same for all panels so could grab it from any
supFig3_2rows = ggdraw(supFig3_2rows_noleg) + draw_grob(leg, hjust = -0.34, vjust = 0.24)
supFig3_2rows
```

### export:
```{r}
# set size and resolution parameters

# requirements: 1000 ppi (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 190
hw_ratio = 1/2 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4
```

```{r}
# create the jpg figure file in this results folder and in the reports folder

jpeg("Results/SupFig3_2rows_outline.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)

supFig3_2rows
dev.off()
```
