---
title: "Supplementary Figure 3"
author: "Anna Goodman"
date: "2024-12-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# load packages
```{r}
library(tidyverse)
library(ggpubr)
library(cowplot)
```

# load data
```{r}
# load in the metadata with PLFA data for all 36 samples
metadata36 <- read.csv("../../TidyData/Metadata/Results/per_sample_metadata_mip.csv")
```

# make depth discrete, not continuous
```{r}
discrete.depth = mutate(metadata36, core_bottom_depth_cm = as.character(core_bottom_depth_cm)) # make a dataframe in which depth is a discrete variable
discrete.depth$core_bottom_depth_cm <- factor(discrete.depth$core_bottom_depth_cm, levels=c("7.5", "15", "22.5", "30", "37.5", "45")) # put the levels of depth in order
```

# define terms for standardization across more than one plot panel
```{r}
fontSizeControl = theme(text = element_text(size = 7)) # code to add to each ggplot to set font size
ymax = 1650 # max ng lipid/g soil to show in bac and fun plot panels

makeMargins = theme(plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 15)) # widen left margin of each plot so the panel labels can fit in top left. 5.5 is default pt size for margins in ggplot.

boxplot_thinlines = geom_boxplot(linewidth = 0.2)

add_leg_box = theme(
    legend.background = element_rect(
      color = "black",
      linewidth = 0.5,
      linetype = "solid"
    )
  )
```


# make boxplot of total living microbial biomass
```{r}
mic = ggplot(discrete.depth, aes(x = core_bottom_depth_cm, y = PLFA_Total_Living_Microbial_Biomass_ng_g, fill = Tree_Health)) +
  boxplot_thinlines +
  labs(x = "Core bottom depth (cm)", y = "Total microbial biomass (ng PLFA/g soil)") +
  theme_minimal()  + scale_fill_manual(name = "Tree mortality", values = c("darkblue", "deeppink")) + # colorblind-friendly colors 
  fontSizeControl +
  makeMargins

```

# make boxplot of total living bacterial biomass
```{r}
bac = ggplot(discrete.depth, aes(x = core_bottom_depth_cm, y = PLFA_Total_Bacteria_ng_g, fill = Tree_Health)) +
  boxplot_thinlines +
  labs(x = "Core bottom depth (cm)", y = "Total bacterial biomass (ng PLFA/g soil)") +
  theme_minimal() + scale_fill_manual(name = "Tree mortality", values = c("darkblue", "deeppink"))+ # colorblind-friendly colors
  fontSizeControl +
  ylim(0, ymax) + # set y-axis equal to other PLFA y axes
  makeMargins
```

# make boxplot of total living fungal biomass
```{r}
fun = ggplot(discrete.depth, aes(x = core_bottom_depth_cm, y = PLFA_Total_Fungi_ng_g, fill = Tree_Health)) +
  boxplot_thinlines +
  labs(x = "Core bottom depth (cm)", y = "Total fungal biomass (ng PLFA/g soil)") +
  theme_minimal()+
  scale_fill_manual(name = "Tree mortality", values = c("darkblue", "deeppink"))+ # colorblind-friendly colors
  fontSizeControl +
  ylim(0, ymax) + # set y-axis equal to other PLFA y axes
  makeMargins +
  add_leg_box
```

# make boxplot of fungi:bacteria ratio
```{r}
discrete.depth = discrete.depth %>% 
  mutate(PLFA_fungi_to_bacteria_ratio = (PLFA_Total_Fungi_ng_g/PLFA_Total_Bacteria_ng_g)) # add column to discrete.depth metadata dataframe with computed fungi:bacteria ratio, based on PLFA biomass of fungi and bacteria

fb_ratio = ggplot(discrete.depth, aes(x = core_bottom_depth_cm, y = PLFA_fungi_to_bacteria_ratio, fill = Tree_Health)) +
  boxplot_thinlines +
  labs(x = "Core bottom depth (cm)", y = "Fungi:bacteria ratio") +
  theme_minimal() + scale_fill_manual(name = "Tree mortality", values = c("darkblue", "deeppink"))+ # colorblind-friendly colors
  fontSizeControl +
  makeMargins
```

#make null/blank panel
```{r}
blank = ggplot(discrete.depth, aes(x = , y = )) + theme_minimal()
blank
```

# combine boxplots as pannels of a single figure - fun:bac in own row
```{r}
supFig3_2rows_noleg = ggpubr::ggarrange(
  mic, bac, fun, blank, fb_ratio, blank, # list of boxplots to be panels
  labels = c("A", "B", "C", "", "D", ""), # labels for each panel
  font.label = list(size = 10, color = "black", face = "bold", family = NULL),
  widths = c(4,4,4),
  heights = c(3,3),
  common.legend = TRUE, # common legend for all panels
  legend = "none", # legend displayed above all panels
  align = "h", # align panels horizontally and vertically
  nrow = 2, ncol = 3 # arrange panels: 1 row for plfa data, 1 row for ratio
)

supFig3_2rows_noleg


```

```{r}
leg = get_legend(fun) # legend same for all panels so could grab it from any
supFig3_2rows = ggdraw(supFig3_2rows_noleg) + draw_grob(leg, hjust = -0.34, vjust = 0.24)
supFig3_2rows
```




### export:
```{r}
# set size and resolution parameters

# requirements: 1000 ppi (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 190
hw_ratio = 1/2 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4
```

```{r}
# create the jpg figure file in this results folder and in the reports folder

jpeg("Results/SupFig3_2rows.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)

supFig3_2rows
dev.off()
```

```{r}
# create A FUN ONLY jpg figure file in this results folder and in the reports folder

jpeg("Results/SupFig3_fun_only.jpg", width = (0.65*w_i) * ppi, height = h_i * ppi, res = ppi)


fun
dev.off()
```

```{r}
# create the png figure file in this results folder and in the reports folder

png("Results/SupFig3.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
png("../../Reports/SupFig3.png", width = w_i * ppi, height = h_i * ppi, res = ppi)


supFig3
dev.off()
```
