---
title: "DO_Remove_Low_Abundance_Reads_ITS"
author: "Glade Bogar"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries
```{r}
library("phyloseq")
packageVersion("phyloseq")

library("ggplot2")
packageVersion("ggplot2")

library("tidyverse")
packageVersion("tidyverse")

#install.packages('devtools')
#library("devtools")
#devtools::install_github("gauravsk/ranacapa")
library("ranacapa")
packageVersion("ranacapa")

library("vegan")
packageVersion("vegan")
```

#load ITS phyloseq object:
```{r}
psITS <- read_rds("Results/psITS.rds")
```

#histogram
```{r}
hist(sample_sums(psITS))
```

#How deep are the samples?
```{r}
sort(sample_sums(psITS))
```

#Let's try something very permissive - 0.01% of the shallowest real sample.

#How deep are the samples?
```{r}
sort(sample_sums(psITS))
```

#Remove 0.01% reads (rounded up) of the shallowest 'real' sample (= 8).

#load DADA2 output data:
```{r}
asv_table_ITS <- readRDS("Results/ASV_Table_ITS.rds")
taxa_table_ITS <- readRDS("Results/Taxa_Table_ITS.rds")
```

#load metadata:
```{r}
metadata_ITS <- read.csv("../../../TidyData/Metadata/Results/phyloseq_metadata_ITS_mip.csv") #CRITICAL THAT YOU USE read.csv, NOT read_csv - BREAKS PHYLOSEQ SAMPLENAMES SOMEHOW
metadata_ITS_2 <- metadata_ITS[,-1]
rownames(metadata_ITS_2) <- metadata_ITS[,1]
metadata_ITS <- metadata_ITS_2
```

#load DADA2 output data:

#Now manually remove 8 reads from every observation in the ASV table:
```{r}
asv_table_ITS_sub8 <- asv_table_ITS - 8
```

#change every negative value to 0:
```{r}
asv_table_ITS_sub8[asv_table_ITS_sub8 < 0] <- 0
```

#re-create ITS phyloseq object:
```{r}
psITS_sub8 <- phyloseq(otu_table(asv_table_ITS_sub8, taxa_are_rows=FALSE), 
               sample_data(metadata_ITS), 
               tax_table(taxa_table_ITS))
psITS_sub8
```

#check the sample sums:
```{r}
sort(sample_sums(psITS_sub8))
```

#remove low read samples:
```{r}
#psITS_sub8
#psITS_sub8 <- subset_samples(psITS_sub8, sample_sums(psITS_sub8) > 3000)
#psITS_sub8
```

#remove taxa that have been removed entirely:
```{r}
psITS_sub8
psITS_sub8 <- filter_taxa(psITS_sub8, function(x) max(x) > 0, TRUE)
psITS_sub8
```

#create rarefction curve:
```{r}
rarecurve_ITS <- ggrare(psITS_sub8, step = 1000, label = "Extraction_ID", se = FALSE)

pdf(file = "Results/rarefactionITS_remove_8_reads_mip.pdf",
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches


rarecurve_ITS

dev.off()
```

# save rarefaction curve for use in supplementary figure 2
```{r}
saveRDS(rarecurve_ITS, file = "Results/rarefaction_ITS_mip.rds")
```


#Calculate NMDS:
```{r}
NMDS_ITS <- ordinate(psITS_sub8, 
                    method = "NMDS", #Ordinate with NMDS
                    distance = "bray", #Use Bray-Curtis distances
                    trymax = 50) #Iterate 50 times
```

#Plot NMDS:
```{r}
plot_ordination(psITS_sub8, #Designate phyloseq
                NMDS_ITS, #Designate NMDS
                label = "Extraction_ID",
                color = "core_top_depth_cm", #Color samples by variable
                title = "Fungi by Depth") +
                theme_classic() #Set ggplot theme = classic
```

#Plot NMDS:
```{r}
plot_ordination(psITS_sub8, #Designate phyloseq
                NMDS_ITS, #Designate NMDS
                label = "Extraction_ID",
                color = "Tree", #Color samples by variable
                title = "Fungi by Tree") +
                theme_classic() #Set ggplot theme = classic
```

#Plot NMDS:
```{r}
plot_ordination(psITS_sub8, #Designate phyloseq
                NMDS_ITS, #Designate NMDS
                label = "Extraction_ID",
                color = "nucleic_acid", #Color samples by variable
                title = "Fungi by Nucleic Acid") +
                theme_classic() #Set ggplot theme = classic
```


#Plot NMDS:
```{r}
plot_ordination(psITS_sub8, #Designate phyloseq
                NMDS_ITS, #Designate NMDS
                label = "Extraction_ID",
                color = "pH_FM_CaCl2", #Color samples by variable
                title = "Fungi by pH") +
                theme_classic() #Set ggplot theme = classic
```

#Plot NMDS:
```{r}
pdf(file = "Results/nmds_ITS_trees_remove_8_reads_mip.pdf",
    width = 7, # The width of the plot in inches
    height = 7) # The height of the plot in inches

plot_ordination(psITS_sub8, #Designate phyloseq
                NMDS_ITS, #Designate NMDS
                label = "Extraction_ID",
                color = "Tree",
                title = "Fungi by Tree") +
                theme_classic() #Set ggplot theme = classic

dev.off()
```


#relativize
```{r}
psITS_sub8_rel <- transform_sample_counts(psITS_sub8, function(x){x / sum(x)})
```

#top 100
```{r}
pdf(file = "Results/top100_bar_remove_8_reads_mip.pdf",
    width = 10, # The width of the plot in inches
    height = 10) # The height of the plot in inches

TopNOTUs <- names(sort(taxa_sums(psITS_sub8_rel), TRUE)[1:100])
psITS_100 = prune_taxa(TopNOTUs, psITS_sub8_rel)
plot_bar(psITS_100, fill = "Phylum")

dev.off()
```

#Save phyloseq object:
```{r}
write_rds(psITS_sub8, "Results/psITS_remove_8_reads.rds")
```




