---
title: "Soil_general_tidy"
output: html_document
date: "2024-09-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#if needed, install tidyverse:
```{r}
#install.packages("tidyverse")
```
#load and check tidyverse package version:
```{r}
library(tidyverse)
packageVersion('tidyverse')
```
#update tidyverse (if not just installed):
```{r}
tidyverse_update()
```
#import data, saved as .csv files: skipping the first 7 rows since it's not necessarry
```{r}
Soil_general <- read_csv("../../Data/Brookside/SOILd_90403_083024.csv", skip = 7)

```

#extract, rename :
```{r}
Soil_general_tidy <- Soil_general


```
#Unite your sample titles:
```{r}
Soil_general_tidy_merged <- Soil_general_tidy %>%
  unite('Sample_id', 'Sample Location', `Sample Description #1`, `Sample Description #2`, sep="_")

```

#rename example%, to an example_percent format
```{r}
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Total_Exchange_Capacity_meq_per_100_g" = "Total Exchange Capacity (meq/100 g)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "organic_matter_percent" = "Organic Matter (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Estimated_Nitrogen_Release_number_N_per_acre" = "Estimated Nitrogen Release (# N/acre)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "S_ppm" = "S* (ppm)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "P_mg_per_kg" = "P* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Ca_mg_per_kg" = "Ca* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Mg_mg_per_kg" = "Mg* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "K_mg_per_kg" = "K* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Na_mg_per_kg" = "Na* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Ca_percent" = "Ca** (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Mg_percent" = "Mg** (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "K_percent" = "K** (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Na_percent" = "Na** (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "other_bases_percent" = "Other Bases** (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "H_percent" = "H** (%)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "B_mg_per_kg" = "B* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Fe_mg_per_kg" = "Fe* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Mn_mg_per_kg" = "Mn* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Cu_mg_per_kg" = "Cu* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Zn_mg_per_kg" = "Zn* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Al_mg_per_kg" = "Al* (mg/kg)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "NO3-N_ppm" = "NO3-N (ppm)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "NH4-N_ppm" = "NH4-N (ppm)")
Soil_general_tidy_merged <- rename(Soil_general_tidy_merged, "Bray_I_P_ppm" = "Bray I P (ppm)")
```
#change variables that are not supposed to be characters into doubles
```{r}
Soil_general_tidy_merged$B_mg_per_kg <- as.numeric(Soil_general_tidy_merged$B_mg_per_kg)
Soil_general_tidy_merged$`NO3-N_ppm` <- as.numeric(Soil_general_tidy_merged$`NO3-N_ppm`)
Soil_general_tidy_merged$`NH4-N_ppm` <- as.numeric(Soil_general_tidy_merged$`NH4-N_ppm`)
```
#replace NAs with 0
```{r}

   Soil_general_tidy_0 <- mutate_all(Soil_general_tidy_merged, ~ifelse(is.na(.), 0, .))
```

#remove Lab Number or other unwanted columns

```{r}
Soil_general_tidy_0 <- select(Soil_general_tidy_0, -"Lab Number")
```

#remove empty rows
```{r}
Soil_general_tidy_0 <- filter(Soil_general_tidy_0, Soil_general_tidy_0$Sample_id != "NA_NA_NA")
```


#export cleaned data:
```{r}
write_csv(Soil_general_tidy_0, "Results/Soil_general_Tidy_mip.csv")
```

