---
title: "Figure_01_AG_11_April_2025"
author: "Anna Goodman"
date: "2025-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("phyloseq")
library("vegan")
library("tidyverse")
library("phyloseq")
library("cowplot")
```


#load pre-filtered, relativized, DNA-only phyloseq objects:
```{r}
# 16S ASVs
ps16S_relative_DNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_DNA_mip.rds")

# ITS ASVs
psITS_relative_DNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_DNA_mip.rds")

# 16S traits
ps16S_traits_relative_DNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_DNA_mip.rds")

# ITS traits
psITS_traits_relative_DNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_DNA_mip.rds")
```

# load in NMDS ordinations
```{r}
# 16S ASVs
NMDS_ps16S_relative_DNA <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/NMDS_ps16S_relative_DNA_mip.rds")

# ITS ASVs
NMDS_psITS_relative_DNA <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/NMDS_psITS_relative_DNA_mip.rds")

# 16S traits
NMDS_ps16S_traits_relative_DNA <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/NMDS_ps16S_traits_relative_DNA_mip.rds")

# ITS traits
NMDS_psITS_traits_relative_DNA <- read_rds("../../Analysis/Microbial Communities/NMDS ordinations/Results/NMDS_psITS_traits_relative_DNA_mip.rds")
```

#load in factor analysis results:
```{r}
factors = read_csv("../../Analysis/Microbial Communities/Factor Analysis/Results/factors_2_mip.csv")

# rename factors
factors <- dplyr::rename(factors, "Proximity to Soil Surface" = "F1")
factors <- dplyr::rename(factors, Weathering = "F2")
```



# Prepare ordination scores and vectors for plotting
```{r}
prep_ordscores_and_vectors = function(ps, nmds, factors_df) {
  ps_vector_data <- sample_data(ps) # take data(?) out of the phyloseq object. This is the same as ps16S_relative_DNA_data at this point, except I'm not 100% sure it's a dataframe? but it's not a tibble
  ps_vector_data_rownames <- row.names(ps_vector_data) # store the rownames from the sample data dataframe; (AVITI sample names)
  ps_vector_data <- as_tibble(ps_vector_data) # make the dataframe(?) a tibble
  ps_vector_data <- left_join(ps_vector_data, factors_df) # adding results of the factor analysis to this tibble -- ok so we are indeed replicating the last section of code, but with a tibble
  ps_vector_data <- select(ps_vector_data,
                                         "Proximity to Soil Surface",
                                         Weathering,
                                         Tree) # simplifying the tibble, selecting only the columns we need here

  ps_vector_data <- as.data.frame(ps_vector_data) # convert tibble to back to a dataframe so it will tolerate rowname reassignment


  rownames(ps_vector_data) <- ps_vector_data_rownames # putting the rownames back onto it? 

#Fit vectors with all numeric data, except chemistry:
  fit <- envfit(nmds, ps_vector_data, na.rm = TRUE) # compute envfit vectors using NMDS results and factor analysis data

#select subset: p < 0.05:
  fit_pval <- data.frame(fit$vectors$pvals) # p-values from envfit vector models into a dataframe
  fit_pval <- rownames_to_column(fit_pval, var = "variable") # re-orient dataframe
  colnames(fit_pval) = c("variable", "pval")
  fit_significant <- filter(fit_pval, pval < 0.05) # keep only vectors with p < 0.05
  fit_significant <- fit_significant$variable

#re-fit with subset:
  ps_vector_data_rownames <- row.names(ps_vector_data)

  ps_vector_data_subset <- 
    select(ps_vector_data, all_of(fit_significant))
  rownames(ps_vector_data_subset) <- ps_vector_data_rownames

  fit <- envfit(nmds, ps_vector_data_subset, na.rm = TRUE)

# Extract ordination scores (coordinates)
ordination_scores <- as.data.frame(scores(nmds, display = "sites"))
colnames(ordination_scores) <- c("NMDS1", "NMDS2")

# Extract metadata (sample data) from the phyloseq object
metadata <- sample_data(ps)

# Add the factor column to ordination scores for coloring
ordination_scores$Tree <- metadata$Tree
ordination_scores$core_bottom_depth_cm <- metadata$core_bottom_depth_cm
ordination_scores$Tree_Health <- metadata$Tree_Health

# Define custom colors for the six levels of the factor
factor_levels <- unique(ordination_scores$Tree)
colors <- c("#FFCEF9", "#FFAAB7", "#D99C4C", "#007C46", "#005260", "#201158")

names(colors) <- factor_levels

# Extract the environmental vectors (arrows)
vectors <- as.data.frame(scores(fit, display = "vectors"))
vectors$R2 <- fit$vectors$r

# Scale the vectors based on their R² value
# Here we multiply the vectors' components by their R² value to scale them accordingly
vectors$NMDS1_scaled <- vectors$NMDS1 * sqrt(vectors$R2) # envfit() automatically scales the vector by R; this line is half of the scalar-multiplication by R (again) required to scale by R2
vectors$NMDS2_scaled <- vectors$NMDS2 * sqrt(vectors$R2) # envfit() automatically scales the vector by R; this line is the other half of the scalar-multiplication by R (again) required to scale by R2

result = list("ordination_scores" = ordination_scores, "vectors" = vectors)


}
```

#16S ASVs:
```{r}
ordvec_16S_ASV = prep_ordscores_and_vectors(ps16S_relative_DNA, NMDS_ps16S_relative_DNA, factors)

ordscores_16S_ASV = ordvec_16S_ASV$ordination_scores
vectors_16S_ASV = ordvec_16S_ASV$vectors 
```
#ITS ASVs:
```{r}
ordvec_ITS_ASV = prep_ordscores_and_vectors(psITS_relative_DNA, NMDS_psITS_relative_DNA, factors)

ordscores_ITS_ASV = ordvec_ITS_ASV$ordination_scores
vectors_ITS_ASV = ordvec_ITS_ASV$vectors 
```
#16S traits:
```{r}
ordvec_16S_traits = prep_ordscores_and_vectors(ps16S_traits_relative_DNA, NMDS_ps16S_traits_relative_DNA, factors)

ordscores_16S_traits = ordvec_16S_traits$ordination_scores
vectors_16S_traits = ordvec_16S_traits$vectors 
```

#ITS ASVs:
```{r}
ordvec_ITS_traits = prep_ordscores_and_vectors(psITS_traits_relative_DNA, NMDS_psITS_traits_relative_DNA, factors)

ordscores_ITS_traits = ordvec_ITS_traits$ordination_scores
vectors_ITS_traits = ordvec_ITS_traits$vectors 
```

# Generate individual elements (four NMDSs + a single legend) to be included in the final figure
## Define a function to create a given NMDS, given its: name (string), ordination scores (data frame), vector coordinates (data frame), and x and y coordinates for placing each vectors' R2 labels (numeric)

```{r}

makePanel_fig1 = function(panel_name, ordscores, vecs, r2x1, r2y1, r2x2, r2y2) {
  
# add tree health information to the tree identity string, so tree health is visible in the legend
ordscores$Tree <- paste(ordscores$Tree, "-", ordscores$Tree_Health)

# make it so the tree identities display in a logical order in the legend (by changing these strings to factor levels)
ordscores$Tree <- factor(ordscores$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))

# define colors
colors <- c("#FFCEF9", "#FFAAB7", "#D99C4C", "#007C46", "#005260", "#201158")
# assign each color to an individual tree identiy
names(colors) <- levels(ordscores$Tree)
# check to see what names(colors) is right now
print(paste(panel_name, " colornames: ", names(colors)))

#define range for the point sizes, which will be scaled with scale_fill_continuous later
commonrange = c(3.2, 1.2)

#define weight for any lines
commonweight = 0.5

#define font size. remember to specify "pt" whenever size_unit is an option
fontsize = 7

# reformat the vector data; this was originally to make it compatible with ggrepel, which I decided not to use
## put the names of each vector in a column
vecs$vector_name <- rownames(vecs)
## make a new column called "type" and put the vector names in there
vecs$type <- ifelse(
  vecs$vector_name == "Proximity to Soil Surface",
  "Proximity to Soil Surface",
  "Weathering"
)

# make a dataframe for the R2 values corresponding to each vector, keeping the arrow's terminating x and y coordinates, and writing a full label for the R2 values
vec_r2_labels <- data.frame(
  x = vecs$NMDS1_scaled,
  y = vecs$NMDS2_scaled,
  label = paste0("R² = ", round(vecs$R2, 2))
)

# make the x and y coordinates of the R2 labels manually adjustable; r2x1, r2y1, etc., are parameters fed into the panel-making function 
vec_r2_labels$x[1] = vec_r2_labels$x[1] + r2x1 # adjusts the proximity to soil surface r2 label horizontally 
vec_r2_labels$y[1] = vec_r2_labels$y[1] + r2y1 # adjusts the proximity to soil surface r2 label vertically 
if (is.na(vec_r2_labels$x[2]) == FALSE) {
    vec_r2_labels$x[2] = vec_r2_labels$x[2] + r2x2 # adjusts the weathering r2 label horizontally, if one exists (no weathering factor for ITS traits)
}
if (is.na(vec_r2_labels$y[2]) == FALSE) {
    vec_r2_labels$y[2] = vec_r2_labels$y[2] + r2y2 # adjusts the weathering r2 label vertically, if one exists (no weathering factor for ITS traits)
}

# actually build the panel
panel <- ggplot(ordscores, aes(x = NMDS1, y = NMDS2, color = Tree, fill = Tree, size = core_bottom_depth_cm)) +

  # plot points
  geom_point(shape = 21, color = "black", stroke = 0.5) + # shape = 21 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  scale_size_continuous(range = commonrange) +

  theme_classic(base_size = fontsize) +
  labs(title = panel_name, x = "", y = "") +
  theme(legend.position = "none")+

  # plot factor analysis vectors
  geom_segment(
    data = vecs,
    aes(x = 0, y = 0, xend = NMDS1_scaled, yend = NMDS2_scaled, linetype = type),
    inherit.aes = FALSE,
    arrow = arrow(type = "closed", length = unit(2, "mm")),
    color = "black",
    linewidth = 0.4
  ) +

  # plot R² labels near vector tips
  geom_text(
    data = vec_r2_labels,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    size = 7, size.unit = "pt",
    color = "black") +

  # Legend for vector types
  scale_linetype_manual(
  name = "Soil factors",
  values = c(
    "Proximity to Soil Surface" = "solid",
    "Weathering" = "dotted"
  ),
  guide = guide_legend(override.aes = list(
    color = "black",
    size = fontsize, size.unit = "pt"  
  ))
)


return(panel)
}

```

## Generate each NMDS panel
```{r}
# generate all panels, specifying their titles and the x,y coordinates of each R2 label
ASV_16S = makePanel_fig1("16S ASVs", ordscores_16S_ASV, vectors_16S_ASV, -0.3, 0.08, 0.55, 0)
ASV_ITS = makePanel_fig1("ITS ASVs", ordscores_ITS_ASV, vectors_ITS_ASV, 0, -0.1, 0.19, -0.12)
trait_16S = makePanel_fig1("16S traits", ordscores_16S_traits, vectors_16S_traits, -0.15, 0.02, -0.05, 0.03)
trait_ITS = makePanel_fig1("ITS traits", ordscores_ITS_traits, vectors_ITS_traits, -0.2, 0.17, 0.1, 0.1)
```

## Make the legend 
### (by making another NMDS plot, which has the fully-informative legend we want but is imperfect in other ways, and extracting the legend as a grob)
```{r}

# Add tree health (living/dead) to the strings indicating tree identity, such that each trees' health is described in the legend
copyof_ordscores = ordscores_16S_ASV

copyof_ordscores$Tree <- paste(copyof_ordscores$Tree, "-", copyof_ordscores$Tree_Health)

copyof_ordscores$Tree <- factor(copyof_ordscores$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))

#define colors

colors <- c("#FFCEF9", "#FFAAB7", "#D99C4C", "#007C46", "#005260", "#201158")
names(colors) <- levels(copyof_ordscores$Tree)
print("colors:")
print(colors)
print("ordscores$Tree:")
print(copyof_ordscores$Tree)

#define point size ranges
commonrange = c(3.2, 1.2)

#define line weights
commonweight = 0.5

#define font size
fontsize = 7


vectors_16S_ASV$vector_name <- rownames(vectors_16S_ASV)  # Pull rownames into a column

vectors_16S_ASV$type <- ifelse(
  vectors_16S_ASV$vector_name == "Proximity to Soil Surface",
  "Proximity to Soil Surface",
  "Weathering"
)

leg_panel <- ggplot(copyof_ordscores) +

  # Sample points
 geom_point(data = copyof_ordscores,
             aes(x = NMDS1,
                 y = NMDS2,
                 color = Tree,
                 size = core_bottom_depth_cm)) +
  scale_color_manual(values = colors) +
  scale_size_continuous(range = commonrange) +

  theme_classic(base_size = fontsize) +
  labs(title = "all legend components?",
       x = "",
       y = "",
                 size = "Core bottom depth (cm)",
                color = "Tree Identity") +
  theme(legend.position = "right")+

  # Environmental vector arrows
  geom_segment(
    data = vectors_16S_ASV,
    aes(x = 0, y = 0, xend = NMDS1_scaled, yend = NMDS2_scaled, linetype = type),
    inherit.aes = FALSE,
    arrow = arrow(type = "closed", length = unit(2, "mm")),
    color = "black",
    size = 0.4
  ) +

  # Legend for vector types
  scale_linetype_manual(
  name = "Soil factors",
  values = c(
    "Proximity to Soil Surface" = "solid",
    "Weathering" = "dotted"
  ),
  guide = guide_legend(override.aes = list(
    color = "black",
    size = 7, size.unit = "pt"  # Keep or tweak as needed
  ))
)

leg_panel
```


# Compile the final figure
```{r} 
# combine individual plots

all_no_legend = plot_grid(ASV_16S, ASV_ITS, trait_16S, trait_ITS,
                 nrow = 2, 
                 ncol = 2, align = 'vh')
                 #labels = c("16S ASVs", "ITS ASVs", "16S traits", "ITS traits"))

labeled = all_no_legend +
  draw_text("NMDS2", x = 0.02, y = 0.45, hjust = "left", angle=90, size = fontsize) +
  draw_text("NMDS1", x = 0.55, y = 0.02, size = fontsize) # x value is 1/2 of the 2/3 of the horizontal space which is the panels

legend = get_legend(leg_panel)

fig1 = ggdraw(plot_grid(labeled, legend, align = "h", rel_widths = c(3, 1)))

fig1
```

# Save the final figure
```{r}
# create the figure file in this results folder

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 140
hw_ratio = 5/7 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

jpeg("Results/Figure1.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig1
dev.off()

jpeg("../../Reports/Figure1.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig1
dev.off()

png("Results/Figure1.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig1
dev.off()

png("../../Reports/Figure1.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig1
dev.off()
```

```

