---
title: 'Figure 3: DNA vs. RNA NMDS'
author: "Anna Goodman"
date: "2025-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load required packages
```{r}
library(tidyverse)
```


------------------------------------------------------------------------------
#Figure 3
------------------------------------------------------------------------------
#load pre-filtered, relativized phyloseq objects:
```{r}
# 16S ASVs
ps16S_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_mip.rds")

# ITS ASVs
psITS_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_mip.rds")

# 16S traits
ps16S_traits_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_mip.rds")

# ITS traits
psITS_traits_relative <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_mip.rds")
```

# load metadata
```{r}
metadata = read.csv("../../TidyData/Metadata/Results/per_sample_metadata_mip.csv")
```


# set universal plot characteristics
```{r}
#define colors

tree_labels <- c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living')

colors <- c("#FFCEF9", "#FFAAB7", "#D99C4C", "#007C46", "#005260", "#201158") # colors from hcl.colors("Batlow", n = 8), omitting two colors in the center of the scale

names(colors) <- tree_labels

shapes <- c(21, 24)
names(shapes) <- c("DNA", "RNA") 

#define point size ranges
commonrange = c(4, 1.5)

#define line weights
commonweight = 0.5

#define font size
fontsize = 7
```

# define functions:

## filter phyloseq objects -- this figure shows pairwise comaparisons between
                              # RNA-derived microbial community composition data
                              # and DNA-derived microbial community composition
                              # data gathered from the same soil cores.
                              # Because some extractions or library preps failed,
                              # we need to filter out data from soil cores for 
                              # which we don't have BOTH DNA- and RNA-based data.
```{r}
filter_ps_for_nucleic_acid_comparison = function (ps_relative) {

# Extract sample_data and convert to data.frame
sd <- as(sample_data(ps_relative), "data.frame")

# Filter out NA values
sd_clean <- sd %>%
  filter(!is.na(DO_ID), !is.na(nucleic_acid))

# Filter out all samples for which we don't have BOTH RNA- and DNA-based data
#   group the rows of sample data (currently referring to a 
#         single extraction->library prep sample) by the soil core sample they
#         originate from (aka DO_ID)
sd_grouped_by_soil_core <- sd_clean %>%
  # Group rows by DO_ID — each group contains all samples sharing the same DO_ID
  group_by(DO_ID) %>%
  
  # Summarize the group to calculate:
  summarize(
    has_DNA = any(nucleic_acid == "DNA"),   # TRUE if there's at least one DNA sample in the group
    has_RNA = any(nucleic_acid == "RNA"),   # TRUE if there's at least one RNA sample in the group
    total = n()                             # Total number of samples for this DO_ID
  ) %>%
  
  # Filter to keep only groups that:
  # - Have at least one DNA sample
  # - Have at least one RNA sample
  filter(has_DNA, has_RNA)


# Keep only samples from DO_IDs that passed filter
paired_samples <- sd_clean %>%
  filter(DO_ID %in% sd_grouped_by_soil_core$DO_ID)


# Subset phyloseq object
ps_relative_comparison <- prune_samples(rownames(paired_samples), ps_relative)
ps_relative_comparison <- prune_taxa(taxa_sums(ps_relative_comparison) > 0, ps_relative_comparison)

return(ps_relative_comparison)

}

```


## make NMDS ordinations
```{r}

nmds_for_nucleic_acid_comparison = function (ps_for_comparison) {

  NMDS <- ordinate(ps_for_comparison,
                          method = "NMDS", #Ordinate with NMDS
                          distance = "bray", #Use Bray-Curtis distances
                          trymax = 100, #Iterate 100 times
                          k = 2) #ordinate in 2 axes
return(NMDS)
  
}
```

## make a single panel 
```{r}
makepanel_fig3 = function (panel_name, ps_for_comparison, nmds){
  
  # convert sample_data from the phyloseq object into a dataframe
  ps_sampledata <- as(sample_data(ps_for_comparison),"data.frame")
  ps_sampledata <- rownames_to_column(ps_sampledata, var = "barcode_pair") # make the rownames into a column so we can join with the nmds ordination points later

  # add tree health information to the tree identity string, so tree health is visible in the legend
  ps_sampledata$Tree <- paste(ps_sampledata$Tree, "-", ps_sampledata$Tree_Health)
  ps_sampledata$Tree <- factor(ps_sampledata$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))
  
  df_points <- data.frame(nmds$points)
  df_points <- rownames_to_column(df_points, var = "barcode_pair")
  df_points <- left_join(df_points, ps_sampledata, by = "barcode_pair")
  
  df_points$Tree <- factor(df_points$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))
  df_points$nucleic_acid <- factor(df_points$nucleic_acid, levels=c('DNA', 'RNA'))

  # Prepare data for pairwise lines
  nucleic_acid_distance <- df_points %>%
    group_by(DO_ID) %>%
    summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
              MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
    ungroup()
  
  dist_frame <- left_join(nucleic_acid_distance, ps_sampledata, by = "DO_ID")
  
  panel = ggplot(df_points) +
          geom_point(data = df_points, color = "black",
                aes(x = MDS1,
                y = MDS2,
                size = core_bottom_depth_cm,
                shape = nucleic_acid,
                fill = Tree,
                stroke = 0.5)) +

          scale_fill_manual(values = colors) +   
          scale_shape_manual(values = shapes) +
    
          geom_segment(data = nucleic_acid_distance,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               linewidth = commonweight) +
           scale_size_continuous(range = commonrange) +
           labs(title = panel_name,
           x = "",
           y = "") +
           theme_classic(base_size = fontsize) +
    theme(legend.position = "none") 
  
panel <- ggplotGrob(panel)
panel$layout$clip[panel$layout$name == "panel"] <- "off"

panel = ggdraw(panel)  

return(panel)
}
```


# execute functions!
## filter phyloseq objects for RNA AND DNA samples
```{r}
ps16S_comparison = filter_ps_for_nucleic_acid_comparison(ps16S_relative)
psITS_comparison = filter_ps_for_nucleic_acid_comparison(psITS_relative)
ps16S_traits_comparison = filter_ps_for_nucleic_acid_comparison(ps16S_traits_relative)
psITS_traits_comparison = filter_ps_for_nucleic_acid_comparison(psITS_traits_relative)

```
## ordinate NMDSs
```{r}
nmds_16S = nmds_for_nucleic_acid_comparison(ps16S_comparison)
nmds_ITS = nmds_for_nucleic_acid_comparison(psITS_comparison)
nmds_16S_traits = nmds_for_nucleic_acid_comparison(ps16S_traits_comparison)
nmds_ITS_traits = nmds_for_nucleic_acid_comparison(psITS_traits_comparison)
```
## make each panel of the plot
```{r}
ASV_16S = makepanel_fig3("16S ASVs", ps16S_comparison, nmds_16S)

ASV_ITS = makepanel_fig3("ITS ASVs", psITS_comparison, nmds_ITS)

trait_16S = makepanel_fig3("16S traits", ps16S_traits_comparison, nmds_16S_traits)

trait_ITS = makepanel_fig3("ITS traits", psITS_traits_comparison, nmds_ITS_traits)
```

# make a panel outside the function just so we can extract the legend from it as a grob
```{r}

  # convert sample_data from the phyloseq object into a dataframe
  ps_sampledata <- as(sample_data(ps16S_comparison),"data.frame")
  # make the rownames into a column so we can join with the nmds ordination points later
  ps_sampledata <- rownames_to_column(ps_sampledata, var = "barcode_pair")

      # add tree health information to the tree identity string, so tree health is visible in the legend
  ps_sampledata$Tree <- paste(ps_sampledata$Tree, "-", ps_sampledata$Tree_Health)
  ps_sampledata$Tree <- factor(ps_sampledata$Tree, levels=c('T02 - dead', 'T05 - dead', 'T07 - dead', 'T03 - living', 'T04 - living', 'T06 - living'))
  
  df_points <- data.frame(nmds_16S$points) # pull points from the NMDS into a dataframe for plotting
  df_points <- rownames_to_column(df_points, var = "barcode_pair") # make the rownames into a column so we can join with the sample_data
  df_points <- left_join(df_points, ps_sampledata, by = "barcode_pair") # df_points is NMDS ordination points + sample_data
  
  # Prepare data for pairwise comparison lines
  nucleic_acid_distance <- df_points %>% 
    group_by(DO_ID) %>%
    summarise(MDS1_DNA = MDS1[1], MDS2_DNA = MDS2[1], 
              MDS1_RNA = MDS1[2], MDS2_RNA = MDS2[2]) %>%
    ungroup() 
  
  dist_frame <- left_join(nucleic_acid_distance, ps_sampledata, by = "DO_ID") # make another dataframe with the line comparison data
  
  legend_plot = ggplot(df_points) +
          geom_point(data = df_points,
                aes(x = MDS1,
                y = MDS2,
                size = core_bottom_depth_cm,
                shape = nucleic_acid,
                color = Tree)) +
          scale_color_manual(values = colors) +  # Custom colors for each factor level

          geom_segment(data = nucleic_acid_distance,
               aes(x = MDS1_DNA,
                   y = MDS2_DNA,
                   xend = MDS1_RNA,
                   yend = MDS2_RNA), 
               color = "black",
               linewidth = commonweight) +
           scale_size_continuous(range = commonrange) +
    
           labs(title = "16S ASVs -- legend plot",
                    size = "Core bottom depth (cm)",
                    shape = "Nucleic acid",
                    color = "Tree Identity") +
            theme(legend.position = "right")+
           theme_classic(base_size = fontsize)
  

  legend_plot
```

# assemble the figure
```{r}

# get legend as grob
legend = get_legend(legend_plot)



all_no_legend = plot_grid(ASV_16S, ASV_ITS, trait_16S, trait_ITS,
                 nrow = 2, 
                 ncol = 2, align = 'vh')

labeled = all_no_legend +
  draw_text("NMDS2", x = 0.02, y = 0.45, hjust = "left", angle=90, size = fontsize) +
  draw_text("NMDS1", x = 0.55, y = 0.02, size = fontsize) # x value is 1/2 of the 2/3 of the horizontal space which is the panels

fig3 = plot_grid(labeled, legend, align = "h", rel_widths = c(3, 1))

#create common x and y labels

#fig3 = fig3_leg + draw_text("NMDS2", x = 0.025, y = 0., hjust = "left", angle=90) + draw_text("NMDS1", x = 0.4, y = 0.025, vjust = "bottom") # x value is 1/2 of the 2/3 of the horizontal space which is the panels

```

```{r}
fig3
```

```{r}
# create the figure file in this results and reports folders

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 140
hw_ratio = 5/7 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

jpeg("Results/Figure3_NMDS_DNA-RNA_mip.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig3
dev.off()

jpeg("../../Reports/Figure3_NMDS_DNA-RNA_mip.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig3
dev.off()


png("Results/Figure3_NMDS_DNA-RNA_mip.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig3
dev.off()

png("../../Reports/Figure3_NMDS_DNA-RNA_mip.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
fig3
dev.off()
```