---
title: "Microbiome Network Analysis: DNA vs RNA (16S)"
author: "Laura Bogar"
output: html_document
---

Note: this is NOT part of the Goodman et al. manuscript; rather, Laura is trying to explore network analyses for additional insight into this data and possible use in teaching.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Install NetCoMi and dependencies if not already present:

# Huge is required for SpiecEasi, but is old and archived.
# package_url <- "https://cran.r-project.org/src/contrib/Archive/huge/huge_1.3.5.tar.gz"
# install.packages(package_url, repos = NULL, type = "source")

# Note: If you don't have FORTRAN (i.e. you're on a Mac), you may need to install it.
# Download the correct package here: https://mac.r-project.org/tools/

# devtools::install_github("zdk123/SpiecEasi")
# devtools::install_github("GraceYoon/SPRING")
# # Then install NetCoMi
# devtools::install_github("stefpeschel/NetCoMi", 
#                          repos = c("https://cloud.r-project.org/",
#                                    BiocManager::repositories()))

# library(here) # Not working 13 Feb 2026
library(tidyverse)
library(phyloseq)
library(NetCoMi)

# here::i_am("Analysis/Networks/")
setwd("./Analysis/Networks/")
source("save_network_results.R")
```

## 1. Load Data
We are pulling in the 16S relative abundance objects. For this exploratory analysis, we will compare the DNA-only and RNA-only datasets.

```{r load-data}
# # 16S DNA
# ps16S_DNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_DNA_mip.rds")

#16S RNA
ps16S_RNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_RNA_mip.rds")

# # ITS DNA
# psITS_relative_DNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_DNA_mip.rds")

# ITS RNA
psITS_RNA <- read_rds("../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_RNA_mip.rds")
```

## 2. Pre-processing & Filtering
Networks are very sensitive to "noise." If an ASV only appears in one sample, it can create a spurious correlation. We will filter to keep only the most prevalent taxa (e.g., present in >10% of samples).

```{r filtering}
# Function to filter rare taxa to keep the network manageable and meaningful
filter_ps <- function(ps) {
  # Keep ASVs present in at least 10% of samples
  ps_filtered <- filter_taxa(ps, function(x) sum(x > 0) > (0.10 * length(x)), TRUE)
  return(ps_filtered)
}

# ps_dna_filt <- filter_ps(ps16S_DNA)
ps16_rna_pref <- filter_ps(ps16S_RNA)
psITS_rna_pref <- filter_ps(psITS_RNA)

# 2. Abundance Filter: Keep only the top 300 or 400 taxa
# For 27 samples, 300 taxa is plenty for a robust network.
top300_rna_names <- names(sort(taxa_sums(ps16_rna_pref), decreasing = TRUE)[1:300])
ps16_rna_filt <- prune_taxa(top300_rna_names, ps16_rna_pref)

top300_rna_names_ITS <- names(sort(taxa_sums(psITS_rna_pref), decreasing = TRUE)[1:300])

psITS_rna_filt <- prune_taxa(top300_rna_names_ITS, psITS_rna_pref)

### "handshake" for cross-kingdom analysis

# Replace barcode names with DO_ID for 16S
sample_names(ps16_rna_filt) <- sample_data(ps16_rna_filt)$`DO_ID`

# Replace barcode names with `DO_ID` for ITS
sample_names(psITS_rna_filt) <- sample_data(psITS_rna_filt)$`DO_ID`

# NOW try the handshake again
common_samples <- intersect(sample_names(ps16_rna_filt), 
                            sample_names(psITS_rna_filt))

message("Success! Found ", length(common_samples), " overlapping samples.")

# 2. Subset both objects to ONLY those common samples
# This ensures they are perfectly paired (e.g., both have 26 samples)
ps16_joint <- prune_samples(common_samples, ps16_rna_filt)
psITS_joint <- prune_samples(common_samples, psITS_rna_filt)

# 3. Optional but safe: Sort them to be 100% sure the order matches
ps16_joint <- ps16_joint %>% phyloseq::prune_samples(sample_names(ps16_joint)[order(sample_names(ps16_joint))], .)
psITS_joint <- psITS_joint %>% phyloseq::prune_samples(sample_names(psITS_joint)[order(sample_names(psITS_joint))], .)
# MAKING A TEST SET

# Define the names of the top 50 taxa by total abundance
# top50_names <- names(sort(taxa_sums(ps16S_relative_DNA), decreasing = TRUE)[1:50])
# 
# # Prune the phyloseq object to keep only those names
# ps_test <- prune_taxa(top50_names, ps16S_relative_DNA)

# Aggregating to Genus level (optional, but makes plots cleaner for students)
# ps_dna_gen <- tax_glom(ps_dna_filt, taxrank = "Genus")
# ps_rna_gen <- tax_glom(ps_rna_filt, taxrank = "Genus")
```

## 3. Constructing the Network
We use the `netConstruct` function. We'll use the **SparCC** method, which is robust for compositional microbiome data.

```{r construct}
# We create a combined network object to compare the two "groups" (DNA vs RNA)

# 1. Setup Parallel Cores
# Using 7 out of 8 to leave 1 for the OS
cores_to_use <- 7

# 2. Run the RNA 16S Network
message("Starting 16S Network...")
net_16s <- netConstruct(ps16_rna_filt, 
                        measure = "sparcc", 
                        cores = cores_to_use, 
                        filtTax = "none", # We already filtered manually
                        normMethod = "none", # SparCC handles compositionality
                         sparsMethod = "none",
                        # thresh = 0.2 # Only show correlations > 0.2
                        verbose = 3) 
                        

analyzed_16s <- netAnalyze(net_16s)

# 3. Run the RNA ITS Network
message("Starting ITS Network...")
net_its <- netConstruct(psITS_rna_filt, 
                        measure = "sparcc", 
                        cores = cores_to_use, 
                        filtTax = "none", # We already filtered manually
                        normMethod = "none", # SparCC handles compositionality
                        sparsMethod = "none",
                        # thresh = 0.2, # Only show correlations > 0.2
                        verbose = 3)

analyzed_its <- netAnalyze(net_its)

### Cross-kingdom network analysis
# 1. Samples are ROWS and Taxa are COLUMNS
# (Check: dim should be ~27 x 300)
bac <- as(otu_table(ps16_joint), "matrix")
fun <- as(otu_table(psITS_joint), "matrix")

# 2. Add prefixes to COLUMNS
colnames(bac) <- paste0("Bac_", colnames(bac))
colnames(fun) <- paste0("Fun_", colnames(fun))

# 3. Combine them side-by-side
# (Check: dim should be 27 x 600)
combined_mat <- cbind(bac, fun)

# 4. Final Verification
message("Final Matrix has ", nrow(combined_mat), " samples and ", ncol(combined_mat), " total taxa.")



# 1. Clear out the tax_table for the construction (SparCC only needs the OTU matrix)
# 1. Extract and TRANSPOSE (t) so samples are rows
data1_mat <- t(as(otu_table(ps16_joint), "matrix"))
data2_mat <- t(as(otu_table(psITS_joint), "matrix"))

# 2. Add prefixes to column names (taxa)
colnames(data1_mat) <- paste0("Bac_", colnames(data1_mat))
colnames(data2_mat) <- paste0("Fun_", colnames(data2_mat))

# 3. Run as a SINGLE dataset
# We don't use data2 here; we put everything in 'data'
net_cross <- netConstruct(data = combined_mat, 
                          measure = "sparcc",
                          cores = 7,
                          sparsMethod = "none",
                          verbose = 3)

# 4. Analyze as usual
analyzed_cross <- netAnalyze(net_cross, hubPar = "eigenvector")


# 6. COLLECT EVERYTHING INTO ONE LIST
# This is the "Master List" the helper function will use
final_networks_list <- list(
  "16S_RNA"      = analyzed_16s,
  "ITS_RNA"      = analyzed_its,
  "Cross_Domain_RNA" = analyzed_cross
)

# 7. EXPORT ALL AT ONCE
# This creates the folder and saves all 3 files using the names above
save_network_results(final_networks_list, path = "Finished_RNA_Networks_Feb13")

# 1. Clear the plotting device just in case
dev.off() 

# 2. Run with absolute minimal arguments
analyzed_cross <- readRDS("Finished_RNA_Networks_Feb13/Cross_Domain_RNA.rds")

# 1. Create a manual color vector (300 Bac, 300 Fun)
# This bypasses the name-matching error
my_colors <- c(rep("skyblue", 300), rep("orange", 300))
names(my_colors) <- colnames(analyzed_cross$fit$assoMat1)

plot(analyzed_cross, 
     edgeFilter = "highestWeight",    # Tell it to give you the TOP connections
     edgeFilterPar = 200,      # Show the 200 strongest lines in the whole dataset
     rmSingles = "all", 
     colorVec = my_colors, 
     # --- LABEL SETTINGS ---
     labels = TRUE,
     shortenLabels = "simple", 
     labelLength = 10,       # Shows first 10 characters of the ASV name
     cexLabels = 0.6,        # Makes the text small so it doesn't overlap as much
     # ----------------------
     nodeSize = "degree",
     title1 = "Oak Cross-Domain: Top 200 Strongest Connections")

# Get the taxa names from the association matrix inside 'input'
net_taxa_names <- colnames(analyzed_cross$input$assoMat1)

# Check the last few
tail(net_taxa_names)

# 1. Create the color vector based on the prefix in the input names
dynamic_colors <- ifelse(grepl("Bac_", net_taxa_names), "skyblue", "orange")

# 2. Assign the names to the vector so NetCoMi can map them to the nodes
names(dynamic_colors) <- net_taxa_names

# 3. Plot with the corrected colors
plot(analyzed_cross, 
     edgeFilter = "highestWeight", 
     edgeFilterPar = 500,        # Increased to 500 to try and "find" the fungi
     rmSingles = "all", 
     colorVec = dynamic_colors, 
     nodeSize = "degree",
     labels = FALSE,
     title1 = "Oak Rhizosphere: Blue=16S, Orange=ITS")

```



<!-- ## 4. Network Analysis (Centrality) -->
<!-- Who are the "hubs" or "influencers" in these networks? We calculate properties like Degree and Eigenvector Centrality. -->

<!-- ```{r analyze} -->
<!-- net_analyzed <- netAnalyze(net_comp,  -->
<!--                            centrLCC = TRUE,  -->
<!--                            avDist = TRUE, -->
<!--                            hubPar = "eigenvector") -->
<!-- ``` -->

<!-- ## 5. Visualization -->
<!-- Now we plot the two networks side-by-side. We will color the nodes by Phylum to see if certain groups (like Proteobacteria) cluster together. -->

<!-- ```{r plot, fig.width=12, fig.height=8} -->
<!-- plot(net_analyzed,  -->
<!--      shortenLabels = "none", -->
<!--      labelScale = FALSE, -->
<!--      rmSmallConnect = TRUE, -->
<!--      # Color by Phylum (Tax rank 2 usually) -->
<!--      nodeColor = "rank",  -->
<!--      featVecCol = 2,  -->
<!--      # Highlight Hubs -->
<!--      hubBorderCol = "black", -->
<!--      title1 = "16S DNA Network (Potential)",  -->
<!--      title2 = "16S RNA Network (Active)", -->
<!--      showLegend = TRUE) -->
<!-- ``` -->

<!-- ## 6. Comparison Statistics -->
<!-- Is the DNA network significantly different from the RNA network? `netCompare` uses permutation tests to find out. -->

<!-- ```{r compare} -->
<!-- comp_stats <- netCompare(net_analyzed, permTest = FALSE) # Set permTest = TRUE for final paper -->
<!-- summary(comp_stats) -->
<!-- ``` -->