---
title: "DeSeq2 Indicator Species Analysis"
author: "Anna Goodman"
date: "2025-06-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#load required packages:
```{r}
library("phyloseq")
library("vegan")
library("tidyverse")
library("DESeq2")
```

#load required data:
```{r}
ps16S <- read_rds("../Assign_traits_16S/Results/ps16S_faprotax_mip.rds")
psITS <- read_rds("../Assign_traits_ITS/Results/psITS_remove_8_reads_fungaltraits_mip.rds")
metadata <- read_csv("../../../TidyData/Metadata/Results/per_sample_metadata_mip.csv")
```

-------------------------------------------------------------------------------
# Transform ITS phyloseq object to hold imputed fungal biomass
## tidy phyloseq object

###read depths:
```{r}
sort(sample_sums(psITS))
```

###subset exclude controls:
```{r}
psITS
psITS <- subset_samples(psITS, core_top_depth_cm >= 0)
psITS
```

###select RNA-only:
```{r}
psITS
psITS_RNA <- subset_samples(psITS, nucleic_acid == "RNA")
psITS_RNA
```

###remove taxa that aren't in the dataset anymore:
```{r}
psITS_RNA
psITS_RNA <- prune_taxa(taxa_sums(psITS_RNA) > 0, psITS_RNA)
psITS_RNA
```

###get sample data:
```{r}
psITS_RNA_sample_data <- data.frame(sample_data(psITS_RNA))
```

##Impute fungal biomass as per-sample fungal PLFA x per-sample RNA based relative abundance of a given group of fungi

###select PLFA data:
```{r}
PLFA_total_fungi <- select(psITS_RNA_sample_data, PLFA_Total_Fungi_ng_g)
PLFA_total_fungi <- PLFA_total_fungi$PLFA_Total_Fungi_ng_g
```

###transform to relative abundances:
```{r}
psITS_RNA_relative <- transform_sample_counts(psITS_RNA, function(x){x / sum(x)})
```

###pre-transform sample sums:
```{r}
sample_sums(psITS_RNA_relative)
```

###transform [otu?] table:
```{r}
# Divide the counts in sample n by size_factors[n] and save in matrix `mat`
mat <- sweep(otu_table(psITS_RNA_relative), 1 + taxa_are_rows(psITS_RNA_relative), PLFA_total_fungi, FUN = `*`)

# Make a new phyloseq object with the transformed counts
psITS_RNA_relative_PLFAcorr <- psITS_RNA_relative
otu_table(psITS_RNA_relative_PLFAcorr) <- otu_table(mat, taxa_are_rows = taxa_are_rows(psITS_RNA_relative))
```

###post-transform sample sums:
```{r}
sample_sums(psITS_RNA_relative_PLFAcorr)
```

### Sums should match total fungal PLFA numbers:
```{r}
PLFA_total_fungi
```
#### worked.


## Re-calculate per-sample fungal functional traits for append to 16S.

### List 'taxonomic' ranks:
```{r}
rank_names(psITS_RNA_relative_PLFAcorr)
```

### extract taxa table
```{r}
# Extract abundance matrix from the phyloseq object
TAXA1 <- as(tax_table(psITS_RNA_relative_PLFAcorr), "matrix")
# Coerce to data.frame
psITS_fungaltraits_taxa <- as.data.frame(TAXA1)
```

### Extract OTU table
```{r}
# Extract abundance matrix from the phyloseq object
OTU1 <- as(otu_table(psITS_RNA_relative_PLFAcorr), "matrix")
# Coerce to data.frame
psITS_fungaltraits_OTU <- as.data.frame(OTU1)
```

### Extract sample names
```{r}
samplenames <- as.data.frame(sample_names(psITS_RNA_relative_PLFAcorr))
```

#### Vandepol split genus Mortierella in 2020 - manually assigning genus mortierella's functions to what are now genera actinomortierella, gryganskiella, linnemannia, lunasporangiospora, necromortierella, podila.
#### Note this in the methods.

### Add samplenames to first column of OTU_Table:
```{r}
psITS_fungaltraits_OTU <- rownames_to_column(psITS_fungaltraits_OTU, var = "DO_sample_names")
```

### 1:Primary_lifestyle

#replace entire first row of OTU_table with Primary_lifestyle column of Taxa_table.csv, including heading.
```{r}
Primary_lifestyle_transposed <- t(psITS_fungaltraits_taxa$Primary_lifestyle)
Primary_lifestyle_transposed <- as.data.frame(Primary_lifestyle_transposed)
column_title <- as.data.frame(c("Primary_lifestyle"))
Primary_lifestyle_transposed <- bind_cols(column_title, Primary_lifestyle_transposed)
colnames(Primary_lifestyle_transposed) <- colnames(psITS_fungaltraits_OTU)
psITS_fungaltraits_OTU <- data.frame(lapply(psITS_fungaltraits_OTU, as.character), stringsAsFactors = FALSE)
Primary_lifestyle <- bind_rows(Primary_lifestyle_transposed, psITS_fungaltraits_OTU)
```

#transpose, fix column names, remove rownames (ASVs), re-code as numeric:
```{r}
Primary_lifestyle <- t(Primary_lifestyle)

colnames(Primary_lifestyle) <- as.character(Primary_lifestyle[1, ])
Primary_lifestyle <- as.data.frame(Primary_lifestyle[-1, ])
rownames(Primary_lifestyle) <- NULL

recode_numeric <- select(Primary_lifestyle, -Primary_lifestyle)
recode_numeric <- data.frame(lapply(recode_numeric, as.numeric))

Primary_lifestyle <- bind_cols(Primary_lifestyle$Primary_lifestyle, recode_numeric)
Primary_lifestyle <- dplyr::rename(Primary_lifestyle, Primary_lifestyle = "...1")
```

#collapse all duplicate values:
```{r}
metadata_Primary_lifestyle_grouped <- Primary_lifestyle %>%
  group_by(Primary_lifestyle) %>%
  summarise(across(everything(), sum))
```

#Transpose, edit for later leftjoin:
```{r}
metadata_Primary_lifestyle_grouped_t <- as.data.frame(t(metadata_Primary_lifestyle_grouped))
metadata_Primary_lifestyle_grouped_t <- rownames_to_column(metadata_Primary_lifestyle_grouped_t, var = "Sequencing_ID")
```

#check sums:
```{r}
checksums <- metadata_Primary_lifestyle_grouped_t
checksums <- checksums[-1,-1]
checksums <- mutate(checksums, across(everything(), as.numeric))
checksums <- mutate(checksums, traitsums = rowSums(across(everything())))

checksums$traitsums
```

#export collapsed metadata:
```{r}
write_csv(metadata_Primary_lifestyle_grouped_t, "Results/imputed_biomass_of_fungal_guilds_by_sample_mip.csv", col_names = FALSE)
```
-------------------------------------------------------------------------------

#add per-sample imputed biomass of ECM (ng/g) to ps16S Metadata

###read depths:
```{r}
sort(sample_sums(ps16S))
```

###subset exclude controls:
```{r}
ps16S
ps16S <- subset_samples(ps16S, core_top_depth_cm >= 0)
ps16S
```

###select RNA-only:
```{r}
ps16S
ps16S_RNA <- subset_samples(ps16S, nucleic_acid == "RNA")
ps16S_RNA
```

###remove taxa that aren't in the dataset anymore:
```{r}
ps16S_RNA
ps16S_RNA <- prune_taxa(taxa_sums(ps16S_RNA) > 0, ps16S_RNA)
ps16S_RNA
```

###get sample data:
```{r}
ps16S_RNA_sample_data <- data.frame(sample_data(ps16S_RNA))
```

###load 'fungal functions (created above):
```{r}
fungal_functions <- read_csv("Results/imputed_biomass_of_fungal_guilds_by_sample.csv")
```

###extract, join data:
```{r}
psITS_metadata <- data.frame(sample_data(psITS_RNA_relative_PLFAcorr))
psITS_metadata <- rownames_to_column(psITS_metadata, var = "SeqID")
fungal_functions <- dplyr::rename(fungal_functions, SeqID = "Primary_lifestyle")

psITS_metadata <- left_join(psITS_metadata, fungal_functions)

ps16S_metadata <- data.frame(sample_data(ps16S_RNA))
ps16S_metadata <- rownames_to_column(ps16S_metadata, var = "SeqID")

psITS_metadata <- select(psITS_metadata, nucleic_acid, DO_ID, ectomycorrhizal)

ps16S_metadata$DO_Name <- str_c(ps16S_metadata$DO_ID, ps16S_metadata$nucleic_acid, sep = "_")
psITS_metadata$DO_Name <- str_c(psITS_metadata$DO_ID, psITS_metadata$nucleic_acid, sep = "_")
psITS_metadata <- select(psITS_metadata, DO_Name, ectomycorrhizal)

ps16S_metadata <- left_join(ps16S_metadata, psITS_metadata, by = "DO_Name")

ps16S_metadata <- column_to_rownames(ps16S_metadata, var = "SeqID")

sample_data(ps16S_RNA) <- sample_data(ps16S_metadata)

ps16S_RNA
```

```{r}
#anna debug
ps16S_RNA_mip_biomass = ps16S_RNA

```
------------------------------------------------------------------------------
#ECM ng/g-associated 16S ASVs, RNA-based, with a 3-tree filter

###designate ASVs:
```{r}
ps16S_ASVs_RNA <- ps16S_RNA
```

###To start, filter for asvs seen in at least 1 time under each of 3 trees:
```{r}
ps16S_ASVs_RNA_for_filter <- ps16S_ASVs_RNA

ps16S_ASVs_RNA_for_filter <- merge_samples(ps16S_ASVs_RNA_for_filter, "Tree")
ps16S_ASVs_RNA_for_filter

ps16S_ASVs_RNA_abundant <- filter_taxa(ps16S_ASVs_RNA_for_filter, function(x) sum(x > 1) >= 3, TRUE)
ps16S_ASVs_RNA_abundant
```

##How much of the community do ASVs which appear beneath 3 trees represent?

###First, relativize phyloseq:
```{r}
ps16S_ASVs_RNA_relative <- transform_sample_counts(ps16S_ASVs_RNA, function(x){(x / sum(x))})
```

###Next, get names of subsetted taxa:
```{r}
ps16S_ASVs_RNA_abundant_names <- names(taxa_sums(ps16S_ASVs_RNA_abundant))
```

###Subset relativized dataset:
```{r}
ps16S_ASVs_RNA_relative
ps16S_ASVs_RNA_relative_abundant <- prune_taxa(ps16S_ASVs_RNA_abundant_names, ps16S_ASVs_RNA_relative)
ps16S_ASVs_RNA_relative_abundant
```

###Bar graph:
```{r}
plot_bar(ps16S_ASVs_RNA_relative_abundant, x = "DO_ID", y = "Abundance", fill = "Phylum")
```

### 60+ %. fewer than with a 2 tree filter, but still majority.

###Subset un-relativized dataset for analysis:
```{r}
ps16S_ASVs_RNA_abundant
ps16S_ASVs_RNA_abundant_onepercent <- prune_taxa(ps16S_ASVs_RNA_abundant_names, ps16S_ASVs_RNA)
ps16S_ASVs_RNA_abundant_onepercent

sort(sample_sums(ps16S_ASVs_RNA_abundant_onepercent))
```

###deseq2 does not run on relativized data. rarefy to common depth, post-filtering:
```{r}
sort(sample_sums(ps16S_ASVs_RNA_abundant_onepercent))

ps16S_ASVs_RNA_abundant_onepercent <- rarefy_even_depth(ps16S_ASVs_RNA_abundant_onepercent, sample.size = 22907, rngseed = 12, replace = TRUE, trimOTUs = TRUE)

sort(sample_sums(ps16S_ASVs_RNA_abundant_onepercent))
```


##Test imputed biomass of ECM (ng/g) as a numeric predictor for bacterial ASV read abundance.

###Filter for only samples with %ECM:
```{r}
ps16S_ASVs_RNA_abundant_onepercent
ps16S_ASVs_RNA_abundant_onepercent <- subset_samples(ps16S_ASVs_RNA_abundant_onepercent, ectomycorrhizal != "NA")
ps16S_ASVs_RNA_abundant_onepercent
```

###scale and center ECM - to interpret as logfold change per 100 ng/g ECM increments:
```{r}
ps16S_ASVs_RNA_abundant_onepercent_sample_data <- data.frame(sample_data(ps16S_ASVs_RNA_abundant_onepercent))

ps16S_ASVs_RNA_abundant_onepercent_sample_data <- mutate(ps16S_ASVs_RNA_abundant_onepercent_sample_data, ECM_scaled = (((ectomycorrhizal) - mean(ectomycorrhizal))/100))

sample_data(ps16S_ASVs_RNA_abundant_onepercent) <- ps16S_ASVs_RNA_abundant_onepercent_sample_data

ps16S_ASVs_RNA_abundant_onepercent
```

###use deseq2 to find microbes that are differentially abundant with greater % ECM:
```{r}
DESeq_ECM <- phyloseq_to_deseq2(physeq = ps16S_ASVs_RNA_abundant_onepercent, design = ~ECM_scaled)
DESeq_ECM <- DESeq(DESeq_ECM, test = "Wald", fitType = "parametric")
```

###DESeq2 Results:
```{r}
DESeq_ECM_Results <- results(DESeq_ECM, cooksCutoff = FALSE)
alpha <- 0.05
sigtab <- DESeq_ECM_Results[which(DESeq_ECM_Results$padj < alpha), ]
sigtab <- cbind(as(sigtab, "data.frame"), as(tax_table(ps16S_ASVs_RNA_abundant_onepercent)[rownames(sigtab), ], "matrix"))
dim(sigtab)

write_csv(sigtab, "Results/DESeq2_16S_ASVs_RNA_ECM_3trees_mip.csv")
```

###Visualize
```{r}
library("ggplot2")
theme_set(theme_bw())
#scale_fill_discrete <- function(palname = "Set1", ...) {
#    scale_fill_brewer(palette = palname, ...)
#}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
```

###many significant asvs 

###export:
```{r}
pdf(file = "Results/DESeq2_16S_ASVs_RNA_ECM_3trees.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches

library("ggplot2")
theme_set(theme_bw())
#scale_fill_discrete <- function(palname = "Set1", ...) {
#    scale_fill_brewer(palette = palname, ...)
#}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

dev.off()
```


##However, how abundant are these asvs? And in how many samples were they found?

###Select our significantly variable ASVs:
```{r}
sigtab_names <- rownames_to_column(sigtab)
Variable_ASVs <- sigtab_names$rowname
```

###get only significantly variable ASVs, relativized:
```{r}
ps16S_ASVs_RNA_variable <- prune_taxa(Variable_ASVs, ps16S_ASVs_RNA_relative)
ps16S_ASVs_RNA_variable
```

###How abundant are these?
```{r}
plot_bar(ps16S_ASVs_RNA_variable, x = "DO_ID", y = "Abundance", fill = "Phylum")
```

###Same as previous; mostly just found in T06 at depth. That's as good as we can do, I think.

###Where are the ectos, I guess is a question?
```{r}
ECM_jitter <- ggplot(ps16S_metadata, aes(x = Tree_Health, y = ectomycorrhizal)) +
  
  geom_boxplot(outlier.alpha = 0,
               aes(x = Tree_Health, y = ectomycorrhizal, fill = Tree_Health)) +
  
  geom_point(aes(x = Tree_Health, y = ectomycorrhizal, fill = Tree_Health),
             position = position_jitterdodge(dodge.width = .75, jitter.width = .5),
             alpha = 1) +
  
  scale_fill_manual(values=c("#009900", "#CC6600")) +
  
  ylim(0, NA) +
  
  theme_classic()

ECM_jitter
```

###That's good, by tree?
```{r}
ECM_jitter <- ggplot(ps16S_metadata, aes(x = Tree, y = ectomycorrhizal)) +
  
  geom_boxplot(outlier.alpha = 0,
               aes(x = Tree, y = ectomycorrhizal, fill = Tree_Health)) +
  
  geom_point(aes(x = Tree, y = ectomycorrhizal, fill = Tree),
             position = position_jitterdodge(dodge.width = .75, jitter.width = .5),
             alpha = 1) +
  
  ylim(0, NA) +
  
  theme_classic()

ECM_jitter
```

###hmm. Kind of weird then that we don't see really many ECM-associated bacteria in many of the T03 samples?



# now subset ASVs with only positive correlations with %ECM:
```{r}
sigtab <- filter(sigtab, log2FoldChange > 0)
dim(sigtab)

write_csv(sigtab, "Results/DESeq2_16S_ASVs_RNA_ECM_positive_3trees_mip.csv")
```

###Visualize
```{r}

theme_set(theme_bw())

# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))


fig4 <- ggplot(sigtab,
       aes(x = Genus,
           y = log2FoldChange,
           color = Phylum,
           size = baseMean)) +
  geom_point() +
  scale_size_continuous(
    name = "Average per-sample read abudance",
    breaks = c(10, 150, 300),
    labels = c("10", "150", "300"),
    range = c(1, 10)
    ) +
  theme(text = element_text(size = 7), axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
  scale_x_discrete(labels = function(x) gsub("g__", "", x)) +
  scale_color_discrete(labels = function(x) gsub("p__", "", x))

fig4
```


#many significant asvs 


##However, how abundant are these asvs? And in how many samples were they found?

###Select our significantly variable ASVs:
```{r}
sigtab_names <- rownames_to_column(sigtab)
Variable_ASVs <- sigtab_names$rowname
```

###get only significantly variable ASVs, relativized:
```{r}
ps16S_ASVs_RNA_variable <- prune_taxa(Variable_ASVs, ps16S_ASVs_RNA_relative)
ps16S_ASVs_RNA_variable
write_rds(ps16S_ASVs_RNA_variable, "Results/ps16S_ASVs_RNA_positive_mip.rds")
```

###How abundant are these?
```{r}
plot_bar(ps16S_ASVs_RNA_variable, x = "DO_ID", y = "Abundance", fill = "Phylum")
```

###Looks very similar.


# export other objects for use in other figures/analyses
```{r}
write_rds(ps16S_ASVs_RNA_variable, "Results/ps16S_ASVs_RNA_variable_fromDeSeq2.rds")
write_csv(ps16S_metadata, "Results/ps16S_metadata_fromDeSeq2.csv")
```









