---
title: "Deep Oaks PERMANOVAs and NMDSs"
author: "Anna Goodman"
date: "2025-04-25"
output: html_document
---
# TO DO IN EDITING THIS:
## 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### load required packages:
```{r}
library("phyloseq")
library("vegan")
library("tidyverse")
library("phyloseq")
```

##Analyze 12 PERMANOVAs:
#   16S ASVs, NMDS ~ depth * tree health * nucleic acid
#   same, RNA-only, NMDS ~ depth * tree health
#   same, DNA-only, NMDS ~ depth * tree health
#   ITS ASVs, NMDS ~ depth * tree health * nucleic acid
#   same, RNA-only, NMDS ~ depth * tree health
#   same, DNA-only, NMDS ~ depth * tree health
#   ITS Niche, NMDS ~ depth * tree health * nucleic acid
#   same, RNA-only, NMDS ~ depth * tree health
#   same, DNA-only, NMDS ~ depth * tree health
#   16S Functions, NMDS ~ depth * tree health * nucleic acid
#   same, RNA-only, NMDS ~ depth * tree health
#   same, DNA-only, NMDS ~ depth * tree health



#load pre-filtered, relativized phyloseq objects:
```{r}
# 16S ASVs
## with data from both nucleic acids
ps16S_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_mip.rds")
## DNA only
ps16S_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_DNA_mip.rds")
## RNA only
ps16S_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_relative_RNA_mip.rds")


# ITS ASVs
## with data from both nucleic acids
psITS_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_mip.rds")
## DNA only
psITS_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_DNA_mip.rds")
## RNA only
psITS_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_relative_RNA_mip.rds")

# 16S traits
## with data from both nucleic acids
ps16S_traits_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_mip.rds")
## DNA only
ps16S_traits_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_DNA_mip.rds")
## RNA only
ps16S_traits_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/ps16S_traits_relative_RNA_mip.rds")


# ITS traits
## with data from both nucleic acids
psITS_traits_relative <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_mip.rds")
## DNA only
psITS_traits_relative_DNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_DNA_mip.rds")
## RNA only
psITS_traits_relative_RNA <- read_rds("../../../TidyData/Phyloseq/Tidy_relativize_phyloseq_objects/Results/psITS_traits_relative_RNA_mip.rds")
```


# Write function to preform PERMANOVAs with common settings
```{r}
PERMANOVA_from_ps_relative = function (ps_relative, nucleic_acid_type) {
  set.seed(8934)
  if (nucleic_acid_type == "single") { 
        ps_relative_df <- as(sample_data(ps_relative),"data.frame")
        ps_relative_dist_matrix <- phyloseq::distance(ps_relative, "bray")
        ps_relative_PERMANOVA <- adonis2(ps_relative_dist_matrix ~
                                          core_bottom_depth_cm * Tree_Health,
                                    strata = ps_relative_df$Tree,
                                         ps_relative_df,
                                     by = "terms")
     } else if (nucleic_acid_type == "both"){ # INCLUDES NUCLEIC ACID AS A FACTOR
         ps_relative_df <- as(sample_data(ps_relative),"data.frame")
         ps_relative_dist_matrix <- phyloseq::distance(ps_relative, "bray")
         ps_relative_PERMANOVA <- adonis2(ps_relative_dist_matrix ~
                                          core_bottom_depth_cm * Tree_Health * nucleic_acid,
                                    strata = ps_relative_df$Tree,
                                        ps_relative_df,
                                     by = "terms") 
     }
  
return(ps_relative_PERMANOVA)
  # this PERMANOVA:
      # evaluates the effects of depth, tree health, and nucleic acid (when applicable), in that order
      # blocks by tree identiy
}
```

#16S ASVs
```{r}
ps16S_relative_PERMANOVA = PERMANOVA_from_ps_relative(ps16S_relative, "both")
ps16S_relative_DNA_PERMANOVA = PERMANOVA_from_ps_relative(ps16S_relative_DNA, "single")
ps16S_relative_RNA_PERMANOVA = PERMANOVA_from_ps_relative(ps16S_relative_RNA, "single")

write_rds(ps16S_relative_PERMANOVA, "Results/ps16S_relative_PERMANOVA_mip.rds")
write_rds(ps16S_relative_DNA_PERMANOVA, "Results/ps16S_relative_DNA_PERMANOVA_mip.rds")
write_rds(ps16S_relative_RNA_PERMANOVA, "Results/ps16S_relative_RNA_PERMANOVA_mip.rds")
```

#ITS ASVs
```{r}
psITS_relative_PERMANOVA = PERMANOVA_from_ps_relative(psITS_relative, "both")
psITS_relative_DNA_PERMANOVA = PERMANOVA_from_ps_relative(psITS_relative_DNA, "single")
psITS_relative_RNA_PERMANOVA = PERMANOVA_from_ps_relative(psITS_relative_RNA, "single")

write_rds(psITS_relative_PERMANOVA, "Results/psITS_relative_PERMANOVA_mip.rds")
write_rds(psITS_relative_DNA_PERMANOVA, "Results/psITS_relative_DNA_PERMANOVA_mip.rds")
write_rds(psITS_relative_RNA_PERMANOVA, "Results/psITS_relative_RNA_PERMANOVA_mip.rds")
```

#16S traits
```{r}
ps16S_traits_relative_PERMANOVA = PERMANOVA_from_ps_relative(ps16S_traits_relative, "both")
ps16S_traits_relative_DNA_PERMANOVA = PERMANOVA_from_ps_relative(ps16S_traits_relative_DNA, "single")
ps16S_traits_relative_RNA_PERMANOVA = PERMANOVA_from_ps_relative(ps16S_traits_relative_RNA, "single")

write_rds(ps16S_traits_relative_PERMANOVA, "Results/ps16S_traits_relative_PERMANOVA_mip.rds")
write_rds(ps16S_traits_relative_DNA_PERMANOVA, "Results/ps16S_traits_relative_DNA_PERMANOVA_mip.rds")
write_rds(ps16S_traits_relative_RNA_PERMANOVA, "Results/ps16S_traits_relative_RNA_PERMANOVA_mip.rds")
```

#ITS traits
```{r}
psITS_traits_relative_PERMANOVA = PERMANOVA_from_ps_relative(psITS_traits_relative, "both")
psITS_traits_relative_DNA_PERMANOVA = PERMANOVA_from_ps_relative(psITS_traits_relative_DNA, "single")
psITS_traits_relative_RNA_PERMANOVA = PERMANOVA_from_ps_relative(psITS_traits_relative_RNA, "single")

write_rds(psITS_traits_relative_PERMANOVA, "Results/psITS_traits_relative_PERMANOVA_mip.rds")
write_rds(psITS_traits_relative_DNA_PERMANOVA, "Results/psITS_traits_relative_DNA_PERMANOVA_mip.rds")
write_rds(psITS_traits_relative_RNA_PERMANOVA, "Results/psITS_traits_relative_RNA_PERMANOVA_mip.rds")
```

# Make summary table for publication
## compile summary data from selected permanova models
```{r}

# The models we will report:

results = list("16S ASVs" = data.frame(ps16S_relative_PERMANOVA),
               "ITS ASVs" = data.frame(psITS_relative_PERMANOVA),
               "16S Traits" = data.frame(ps16S_traits_relative_PERMANOVA),
               "ITS Traits" = data.frame(psITS_traits_relative_PERMANOVA),
               "ITS Traits - DNA only" = data.frame(psITS_traits_relative_DNA_PERMANOVA),
               "ITS Traits - RNA only" = data.frame(psITS_traits_relative_RNA_PERMANOVA))

# We are reporting PERMANOVA models with data from both nucleic acids. In the Case of ITS Traits, we are also reporting single nucleic acid models only because there was a significant Tree Mortality x Nucleic Acid interaction

```

## Tidy text 
```{r}

results <- lapply(results, function(df) { # modify each summary dataframe in the following ways:
  
  # the factors are currently row names; make them their own column so we can manipulate their formatting
  df <- cbind(Factor = rownames(df), df) # copy rownames into a column, put it on the left side of the dataframe
  rownames(df) <- NULL # remove the rownames now that factors are represented in a column
  
  # make edits to all the strings in each dataframe so they're more readable in the final table
  df[] <- lapply(df, function(col) { # apply the following changes to each column of each dataframe
    if (is.character(col)) { # but only if the column has character data in it
      col <- gsub("core_bottom_depth_cm", "Depth", col)
      col <- gsub("Tree_Health", "Tree Mortality", col)
      col <- gsub("nucleic_acid", "Nucleic Acid", col)
      col <- gsub(":", " x ", col)
    }
    col
  })
  df
})


results <- Map(function(df, name) { # Add the name of the model as a column, and put it on the far left side of the dataframe
  df$Model <- name
  df = select(df, c(Model, Factor, R2, F, Pr..F.))
}, results, names(results))

results
```

## bind tidy results into a single dataframe
```{r}
table <- do.call(rbind, results)

# Fix the column names
table <- table %>%
  rename_with(~ ifelse(.x == "Pr..F.", "Pr(>F)", .x))

rownames(table) <- NULL # remove the rownames now that factors are represented in a column

colnames(table)

table
```

## export
``` {r}
write_csv(table, "Results/Permanova_table_mip.csv")
write_csv(table, "../../../Reports/Permanova_table_mip.csv")
```

