---
title: "DO_Remove_Low_Abundance_Reads_16S"
author: "Glade Bogar"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries
```{r}
library("phyloseq")
packageVersion("phyloseq")

library("ggplot2")
packageVersion("ggplot2")

library("tidyverse")
packageVersion("tidyverse")

#install.packages('devtools')
#library("devtools")
#devtools::install_github("gauravsk/ranacapa")
library("ranacapa")
packageVersion("ranacapa")

library("vegan")
packageVersion("vegan")
```

#load 16S phyloseq object:
```{r}
ps16S <- read_rds("Results/ps16S.rds")
```

#histogram
```{r}
hist(sample_sums(ps16S))
```

#How deep are the samples?
```{r}
sort(sample_sums(ps16S))
```

#subset:
```{r}
psNegatives <- ps16S %>% subset_samples(Extraction_ID %in% c("D63-D",
                                                         "Negative"))
```

#investigate:
```{r}
sort(sample_sums(psNegatives))
```

#Remove 0.01% reads (rounded up) of the shallowest 'real' sample (= 3).

#load the original 16S phyloseq components.

#load DADA2 output data:
```{r}
asv_table_16S <- readRDS("Results/ASV_Table_16S.rds")
taxa_table_16S <- readRDS("Results/Taxa_Table_16S_species.rds")
```

#load metadata:
```{r}
metadata_16S <- read.csv("../../../TidyData/Metadata/Results/phyloseq_metadata_16S_mip.csv") #CRITICAL THAT YOU USE read.csv, NOT read_csv - BREAKS PHYLOSEQ SAMPLENAMES SOMEHOW
metadata_16S_2 <- metadata_16S[,-1]
rownames(metadata_16S_2) <- metadata_16S[,1]
metadata_16S <- metadata_16S_2


```


#Now manually remove 3 reads from every observation in the ASV table:
```{r}
asv_table_16S_sub3 <- asv_table_16S - 3
```

#change every negative value to 0:
```{r}
asv_table_16S_sub3[asv_table_16S_sub3 < 0] <- 0
```

#re-create 16S phyloseq object:
```{r}
ps16S_sub3 <- phyloseq(otu_table(asv_table_16S_sub3, taxa_are_rows=FALSE), 
               sample_data(metadata_16S), 
               tax_table(taxa_table_16S))
ps16S_sub3
```

#check the sample sums:
```{r}
sort(sample_sums(ps16S_sub3))
```

#remove taxa that have been removed entirely:
```{r}
ps16S_sub3
ps16S_sub3 <- filter_taxa(ps16S_sub3, function(x) max(x) > 0, TRUE)
ps16S_sub3
```

#create rarefction curve. save to ... /Phyloseq_16S/Results
```{r}
rarecurve_16S <- ggrare(ps16S_sub3, step = 1000, label = "Extraction_ID", se = FALSE)


pdf(file = "Results/rarefaction16S_remove_3_reads_mip.pdf",
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches

rarecurve_16S

dev.off()
```

# save rarefaction curve for use in supplementary figure 2
```{r}
saveRDS(rarecurve_16S, file = "Results/rarefaction_16S_mip.rds")
```

#Calculate NMDS:
```{r}
NMDS_16S <- ordinate(ps16S_sub3, 
                    method = "NMDS", #Ordinate with NMDS
                    distance = "bray", #Use Bray-Curtis distances
                    trymax = 50) #Iterate 50 times
```

#Plot NMDS:
```{r}
plot_ordination(ps16S_sub3, #Designate phyloseq
                NMDS_16S, #Designate NMDS
                label = "Extraction_ID",
                color = "core_top_depth_cm", #Color samples by variable
                title = "Bacteria by Depth") +
                theme_classic() #Set ggplot theme = classic
```

#negatives and positive control at least filter out, but should check on the read depth of 'D63-D'.

#Plot NMDS:
```{r}
plot_ordination(ps16S_sub3, #Designate phyloseq
                NMDS_16S, #Designate NMDS
                label = "Extraction_ID",
                color = "Tree", #Color samples by variable
                title = "Bacteria by Tree") +
                theme_classic() #Set ggplot theme = classic
```

#Plot NMDS:
```{r}
plot_ordination(ps16S_sub3, #Designate phyloseq
                NMDS_16S, #Designate NMDS
                label = "Extraction_ID",
                color = "nucleic_acid", #Color samples by variable
                title = "Bacteria by Nucleic Acid") +
                theme_classic() #Set ggplot theme = classic
```

#Plot NMDS:
```{r}
plot_ordination(ps16S_sub3, #Designate phyloseq
                NMDS_16S, #Designate NMDS
                label = "Extraction_ID",
                color = "pH_FM_CaCl2", #Color samples by variable
                title = "Bacteria by pH") +
                theme_classic() #Set ggplot theme = classic
```


#Plot NMDS:
```{r}
pdf(file = "Results/nmds_16S_trees_remove_3_reads.pdf",
    width = 7, # The width of the plot in inches
    height = 7) # The height of the plot in inches

plot_ordination(ps16S_sub3, #Designate phyloseq
                NMDS_16S, #Designate NMDS
                label = "Extraction_ID",
                color = "Tree",
                title = "Bacteria by Tree") +
                theme_classic() #Set ggplot theme = classic

dev.off()
```

#Make a barchart

#relativize
```{r}
ps16S_sub3_rel <- transform_sample_counts(ps16S_sub3, function(x){x / sum(x)})
```

#print, top 100 ASVs
```{r}
pdf(file = "Results/top100_bar_16S_remove_3_reads.pdf",
    width = 10, # The width of the plot in inches
    height = 10) # The height of the plot in inches

TopNOTUs <- names(sort(taxa_sums(ps16S_sub3_rel), TRUE)[1:100])
ps16S_100 = prune_taxa(TopNOTUs, ps16S_sub3_rel)
plot_bar(ps16S_100, fill = "Phylum")

dev.off()
```

#Save phyloseq object:
```{r}
write_rds(ps16S_sub3, "Results/ps16S_remove_3_reads.rds")
```




