---
title: "Assign_Fungal_Traits"
author: "Glade Bogar"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages:
```{r}
library("phyloseq")
packageVersion("phyloseq")

library("ggplot2")
packageVersion("ggplot2")

library("tidyverse")
packageVersion("tidyverse")
```

#Load phyloseq object:
```{r}
psITS <- readRDS("../Phyloseq_ITS/Results/psITS_remove_8_reads.rds")
```

#import genus-level fungal functional traits from "FungalTraits: a user‐friendly traits database of fungi and fungus‐like stramenopiles Polme et al., 2020", genera - "13225_2020_466_MOESM4_ESM" saved as a .csv:
```{r}
fungaltraits_genera <- read.csv("../../../Data/FungalTraits_Genera.csv")
```

#tidy fungaltraits_genera for left join to taxonomy table:
```{r}
fungaltraits_genera <- as_tibble(fungaltraits_genera)
fungaltraits_genera <- fungaltraits_genera %>% mutate(across(where(is.factor), as.character))
```

```{r}
fungaltraits_genera <- dplyr::select(fungaltraits_genera,
                              -jrk_template,
                              -Phylum,
                              -Class,
                              -Order,
                              -Family,
                              -COMMENT.on.genus,
                              -X)
```

```{r}
fungaltraits_genera <- rename(fungaltraits_genera, Genus = GENUS)
fungaltraits_genera <- rename(fungaltraits_genera, Primary_lifestyle = primary_lifestyle)
```

```{r}
fungaltraits_genera$Genus_name <- c("g_")
```

```{r}
fungaltraits_genera <- unite(fungaltraits_genera, Genus, Genus_name, Genus, remove = TRUE)
```

#save formatted fungaltraits_genera for later use:
```{r}
saveRDS(fungaltraits_genera, "Results/fungaltraits_genera_mip.rds")
```

#read taxa table:
```{r}
taxa_ITS <- tax_table(psITS)
taxa_ITS <- taxa_ITS@.Data
taxa_ITS <- as.data.frame(taxa_ITS)
```

#Left join fungaltraits onto taxa_ITS, using "Genus":
```{r}
taxa_ITS_traits <- left_join(taxa_ITS, fungaltraits_genera, by = "Genus")
```

#re-format taxa table for phyloseq:
```{r}
taxa_ITS_traits <- as.matrix(taxa_ITS_traits)
row.names(taxa_ITS_traits) <- row.names(taxa_ITS)
```

#create new tax_table with functions:
```{r}
taxa_ITS_traits <- tax_table(taxa_ITS_traits)
```

#replace the taxonomy table in the original ps:
```{r}
tax_table(psITS) <- taxa_ITS_traits
```

#save new phyloseq object:
```{r}
saveRDS(psITS, "Results/psITS_remove_8_reads_fungaltraits_mip.rds")
```


