---
title: "Factor Analysis"
author: "Anna Goodman"
date: "2025-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#load needed packages:
```{r}
library('tidyverse')
packageVersion('tidyverse')

library('psych')
packageVersion('psych')

library('ggplot2')
packageVersion('ggplot2')

library(grid)

library(gridGraphics)

```

#load per-sample metadata:
```{r}
metadata <- read_csv("../../../TidyData/Metadata/Results/per_sample_metadata_mip.csv")
```

#Select (numeric only) predictive data
```{r}
sample_data_for_FA <- metadata %>% 
    select(where(is.numeric)) #get numeric

write_csv(sample_data_for_FA, "Results/metadata_temp_mip.csv")

sample_data_for_FA <- select(sample_data_for_FA,
                                              core_bottom_depth_cm,
                                              latitude,
                                              longitude,
                                              PerCMoisture,
                                              bulk_density_g_cm3,
                                              Value_dry,
                                              Value_wet,
                                              Chroma_dry,
                                              Chroma_wet,
                                              clay_percent,
                                              sand_percent,
                                              pH_FM_CaCl2,
                                              Total_Exchange_Capacity_meq_per_100_g,
                                              organic_matter_percent,
                                              S_ppm,
                                              P_mg_per_kg,
                                              Ca_mg_per_kg,
                                              Mg_mg_per_kg,
                                              K_mg_per_kg,
                                              Na_mg_per_kg,
                                              #Ca_percent,
                                              #Mg_percent,
                                              #K_percent,
                                              #Na_percent,
                                              B_mg_per_kg,
                                              Fe_mg_per_kg,
                                              Mn_mg_per_kg,
                                              Cu_mg_per_kg,
                                              Zn_mg_per_kg,
                                              Al_mg_per_kg,
                                              "NO3-N_ppm",
                                              "NH4-N_ppm",
                                              Bray_I_P_ppm,
                                              carbon_percent,
                                              nitrogen_percent,
                                              C_per_N_Ratio,
                                              # DNA_qubit_ng_g_soil,
                                              # RNA_qubit_ng_g_soil,
                                              PLFA_Total_Living_Microbial_Biomass_ng_g,
                                              PLFA_Total_Bacteria_ng_g,
                                              PLFA_Total_Fungi_ng_g,
                                              PLFA_Fungi_Bacteria_Ratio_ng_g)

write_csv(sample_data_for_FA, "Results/sample_data_for_FA_mip.csv")
```

#factor analysis will not run with NAs. So, do we drop samples with NA values, or variables with NA values?

#Given the relatively small number of samples, and the need to accuralte represent deep samples, drop NA variables:
```{r}
sample_data_for_FA <- sample_data_for_FA[ , colSums(is.na(sample_data_for_FA)) == 0]
```

### BEGIN ANALYSIS ###

#view dataset:
```{r}
describe(sample_data_for_FA)
```

#retrieve dataset dimensions:
```{r}
dim(sample_data_for_FA)
```

#"KMO"
#"The Kaiser-Meyer-Olkin (KMO) used to measure sampling adequacy is a better measure of factorability."
```{r}
KMO(sample_data_for_FA)
```

#"According to Kaiser’s (1974) guidelines, a suggested cutoff for determining the factorability of the sample data is KMO ≥ 60.

#increase the KMO by omitting 'low-contribution' variables:
```{r}
sample_data_for_FA_trim <- sample_data_for_FA[, KMO(sample_data_for_FA)$MSAi > 0.12] # Get rid of all variables with MSA < 0.12

KMO(sample_data_for_FA_trim)
```

#"Bartlett’s Test of Sphericity"
```{r}
cortest.bartlett(sample_data_for_FA_trim)
cortest.bartlett(cor(sample_data_for_FA_trim))
```

#explanation: "Small values (p < 0.05) of the significance level indicate that a factor analysis may be useful with our data.

#test determinant:
```{r}
det(cor(sample_data_for_FA_trim))
```

#Explanation: "We have a positive determinant, which means the factor analysis will probably run."
#So it's miniscule but technically positive; will it go?

#The number of factors to extract

#Scree plot:
```{r}
library(ggplot2)

fafitfree <- fa(sample_data_for_FA_trim,
                nfactors = ncol(sample_data_for_FA_trim),
                rotate = "none")

n_factors <- length(fafitfree$e.values)

scree <- data.frame(
  Factor_n =  as.factor(1:n_factors), 
  Eigenvalue = fafitfree$e.values)

ggplot(scree, aes(x = Factor_n, y = Eigenvalue, group = 1)) + 
  geom_point() + geom_line() +
  xlab("Number of factors") +
  ylab("Initial eigenvalue") +
  labs( title = "Scree Plot", 
        subtitle = "(Based on the unreduced correlation matrix)")
```


#Parallel Analysis

#"We can use the parallel() function from the nFactors package (Raiche & Magis, 2020) to perform a parallel analysis"
```{r}
parallel <- fa.parallel(sample_data_for_FA_trim)
```


#Conducting the Factor Analysis

#Factor analysis using fa method, 3 factors:
```{r}
fa.none <- fa(r = sample_data_for_FA_trim, 
              nfactors = 3,
              # covar = FALSE,
              #SMC = TRUE,
              fm = "pa", # type of factor analysis we want to use (“pa” is principal axis factoring)
              max.iter = 100, # (50 is the default, but we have changed it to 100
              rotate = "varimax") # none rotation
print(fa.none)
```

#Factor Analysis using the factanal method:
```{r}
factanal.none <- factanal(sample_data_for_FA_trim,
                          factors = 3,
                          scores = c("regression"),
                          rotation = "varimax")
print(factanal.none)
```


#Regression analysis:
```{r}
head(fa.none$scores)
```

#Labeling the data
```{r}
fa_data <- cbind(metadata["DO_ID"], fa.none$scores)

#Labeling the data
names(fa_data) <- c("DO_ID", "F1", "F2", "F3")
head(fa_data)

write_csv(fa_data, "Results/factors_3.csv")
```


#Factor analysis using fa method, 2 factors:
```{r}
fa.none <- fa(r = sample_data_for_FA_trim, 
              nfactors = 2,
              # covar = FALSE,
              #SMC = TRUE,
              fm = "pa", # type of factor analysis we want to use (“pa” is principal axis factoring)
              max.iter = 100, # (50 is the default, but we have changed it to 100
              rotate = "varimax") # none rotation
print(fa.none)
```

#Factor Analysis using the factanal method:
```{r}
factanal.none <- factanal(sample_data_for_FA_trim,
                          factors = 2,
                          scores = c("regression"),
                          rotation = "varimax")
print(factanal.none)
```


#Graph factor Loading Matrices, save as a pdf in various places
## this results folder
```{r}
pdf(file = "Results/Factor_Analysis_Graph_2factors_mip.pdf",   # The directory you want to save the file in
    width = 30, # The width of the plot in inches
    height = 8) # The height of the plot in inches

 fa.diagram(fa.none)
 
dev.off()
```
## figures folder
```{r}
pdf(file = "../../../Figures/Supplementary Figures/Results/sup_fig_4_Factor_Analysis_Graph_2factors_mip.pdf",   # The directory you want to save the file in
    width = 30, # The width of the plot in inches. stupid wide because that's the only way to not get the graph cut off
    height = 8) # The height of the plot in inches

 fa.diagram(fa.none)
 
dev.off()
```
## reports folder
```{r}
pdf(file = "../../../Reports/Supplementary Material/Figures/sup_fig_4_Factor_Analysis_Graph_2factors_mip.pdf",   # The directory you want to save the file in
    width = 30, # The width of the plot in inches
    height = 8) # The height of the plot in inches

 fa.diagram(fa.none)
 
dev.off()
```


#Regression analysis:
```{r}
head(fa.none$scores)
```

#Labeling the data
```{r}
fa_data <- cbind(metadata["DO_ID"], fa.none$scores)

#Labeling the data
names(fa_data) <- c("DO_ID", "F1", "F2")
head(fa_data)

write_csv(fa_data, "Results/factors_2.csv")
write_csv(fa_data, "../../../Reports/Supplementary Material/sup_table_factors2.csv")

```
