---
title: "Figure 3"
author: "Anna Goodman"
date: "2024-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load packages
```{r}
library(tidyverse)
library(cowplot)
```

# load data
```{r}
imputed_biomass <- read.csv("../../Analysis/Microbial Communities/Imputed Biomass/Results/RNA_based_fungal_trait_biomass.csv")
```

# aglomorate biomasses of plant pathogens and saprotrophs
```{r}
imputed_biomass <- mutate(imputed_biomass, mass_saprotrophs_plant_pathogens = dung_saprotroph +
                               litter_saprotroph +
                               nectar.tap_saprotroph +
                               plant_pathogen +
                               pollen_saprotroph +
                               soil_saprotroph +
                               sooty_mold +
                               unspecified_saprotroph +
                               wood_saprotroph) # add column to imputed_biomass with each sample's mass sum of saprotrophs and plant pathogens
```

# aglomorate biomasses of fungi from all other assigned lifestyles except for ectomycorrhizal fungi
```{r}
imputed_biomass <- mutate(imputed_biomass, other = algal_parasite + 
                        animal_parasite +
                        epiphyte + 
                        lichen_parasite + 
                        protistan_parasite +
                        unspecified +
                        arbuscular_mycorrhizal +
                        foliar_endophyte +
                        lichenized +
                        moss_symbiont +
                        root_endophyte +
                        mycoparasite) # add column to imputed_biomass with each sample's mass sum of other named lifestyles, excluding ectomycorrhizae
```

### double-check that these aglomerated mass sums still add up to the total fungal biomass in each sample
```{r}
aglom_totals = mutate(imputed_biomass, sumtotal = ectomycorrhizal +
                        mass_saprotrophs_plant_pathogens +
                        NA. +
                        other)
aglom_totals = select(aglom_totals, c(sumtotal, PLFA_Total_Fungi_ng_g))
#View(aglom_totals)
```


# tidy and reformat the dataframe for ggplot
```{r} 
imputed_biomass = select(imputed_biomass, c(ectomycorrhizal, mass_saprotrophs_plant_pathogens, NA., other, PLFA_Total_Fungi_ng_g, core_bottom_depth_cm, Tree_Health, DO_ID)) # select only variables used in the plot

imputed_biomass = dplyr::rename(imputed_biomass, "Ectomycorrhizal fungi" = ectomycorrhizal, "Plant pathogens and saprotrophs" = mass_saprotrophs_plant_pathogens, "All other lifestyles" = other, "Unknown lifestyle" = NA., "TOTAL fungal biomass" = PLFA_Total_Fungi_ng_g) # rename response variables for clarity in the display

imputed_biomass = pivot_longer(imputed_biomass, 1:5, names_to = "Primary Lifestyle", values_to = "biomass") # make primary lifestyle a single categorical variable

imputed_biomass$'Primary Lifestyle' <- factor(imputed_biomass$'Primary Lifestyle', levels=c('TOTAL fungal biomass', 'Ectomycorrhizal fungi', 'Plant pathogens and saprotrophs', 'Unknown lifestyle', 'All other lifestyles')) # order primary lifestyles to appear logically in the legend
```

# make tibbles with only some lifestyles in them for slow reveal
```{r}
just_total = imputed_biomass
for (i in 1:length(just_total$`Primary Lifestyle`)) {
  if (just_total$`Primary Lifestyle`[i] != 'TOTAL fungal biomass') {
    just_total$biomass[i] = NA
  }
} # makes just_total as a version of imputed_biomass that only has total fungal biomass data, with NAs in every other lifestye category (this will make a figure with the full legend but only totals data)

ecm_total = imputed_biomass
for (i in 1:length(ecm_total$`Primary Lifestyle`)) {
  if (ecm_total$`Primary Lifestyle`[i] != 'TOTAL fungal biomass' & just_total$`Primary Lifestyle`[i] != 'Ectomycorrhizal fungi') {
    ecm_total$biomass[i] = NA
  }
} # does the same thing but also displaying ecm biomass data



```


# make figure 3
```{r}
manual_colors = c('TOTAL fungal biomass' = "black", 
                  'Ectomycorrhizal fungi' = "#C87A8A",
                  'Plant pathogens and saprotrophs' = "#A29048",
                  'Unknown lifestyle' = "#00A396",
                  'All other lifestyles' = "#9189C7") # assign colors to primary lifestyles

fig3_just_total = ggplot(just_total, aes(x = core_bottom_depth_cm, y = biomass, color = `Primary Lifestyle`)) + 
    geom_smooth(method='loess', se = FALSE) + 
   geom_point(size = 2) +
   facet_wrap(~Tree_Health) + 
    ylab("Imputed biomass (ng PLFA/g soil)") +
    xlab("Core bottom depth (cm)") +
    xlim(7, NA) +
    ylim(-10, 1200) + theme_classic(base_size = 7) +
    scale_color_manual(values = manual_colors) +  
    theme(legend.title=element_blank(), 
        legend.box.background = element_rect(color="black"),
        legend.box.margin = margin(3, 3, 3, 3))

fig3_ecm_total = ggplot(ecm_total, aes(x = core_bottom_depth_cm, y = biomass, color = `Primary Lifestyle`)) + 
    geom_smooth(method='loess', se = FALSE) + 
   geom_point(size = 2) +
   facet_wrap(~Tree_Health) + 
    ylab("Imputed biomass (ng PLFA/g soil)") +
    xlab("Core bottom depth (cm)") +
    xlim(7, NA) +
    ylim(-10, 1200) + theme_classic(base_size = 7) +
    scale_color_manual(values = manual_colors) +  
    theme(legend.title=element_blank(), 
        legend.box.background = element_rect(color="black"),
        legend.box.margin = margin(3, 3, 3, 3))

fig3_all = ggplot(imputed_biomass, aes(x = core_bottom_depth_cm, y = biomass, color = `Primary Lifestyle`)) + 
    geom_smooth(method='loess', se = FALSE) + 
   geom_point(size = 2) +
   facet_wrap(~Tree_Health) + 
    ylab("Imputed biomass (ng PLFA/g soil)") +
    xlab("Core bottom depth (cm)") +
    xlim(7, NA) +
    ylim(-10, 1200) + theme_classic(base_size = 7) +
    scale_color_manual(values = manual_colors) +  
    theme(legend.title=element_blank(), 
        legend.box.background = element_rect(color="black"),
        legend.box.margin = margin(3, 3, 3, 3))


fig3_just_total
fig3_ecm_total
fig3_all
```





```{r}
# create the jpg figure file in this results folder and in the reports folder

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 190
hw_ratio = 5/11 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

jpeg("Results/Figure3_just_total_for_slides.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
jpeg("../../Reports/Figure3_just_total_for_slides.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)

fig3_just_total
dev.off()

jpeg("Results/Figure3_ecm_total_for_slides.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
jpeg("../../Reports/Figure3_ecm_total_for_slides.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)

fig3_ecm_total
dev.off()

jpeg("Results/Figure3_all_for_slides.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)
jpeg("../../Reports/Figure3_all_for_slides.jpg", width = w_i * ppi, height = h_i * ppi, res = ppi)

fig3_all
dev.off()
```

```{r}
# create the png figure file in this results folder and in the reports folder

# requirements: 1000 pp1 (pixels per inch), 140 mm wide
# 1 inch = 0.0393701 mm, 1 mm = 25.4 inches

ppi = 1000

w_mm = 190
hw_ratio = 5/11 # arbitrary

h_mm = w_mm*hw_ratio
w_i = w_mm/25.4
h_i = h_mm/25.4

png("Results/Figure3.png", width = w_i * ppi, height = h_i * ppi, res = ppi)
png("../../Reports/Figure3.png", width = w_i * ppi, height = h_i * ppi, res = ppi)


fig3_cow
dev.off()
```

