---
title: "Linear Mixed-Effect Models Formalized"
author: "Anna Goodman"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Mixed-Effect Models

## 10 analyses of interest:

### Microbial population sizes

#### 1) Total living microbial biomass; PLFA ~ depth * tree mortality, block = tree identity
#### 2) Total living bacterial biomass; PLFA ~ depth * tree mortality, block = tree identity
#### 3) Total living fungal biomass; log(PLFA) ~ depth * tree mortality, block = tree identity
#### 4) Fungi:bacteria ratio; log(total fungal PLFA/total bacterial PLFA) ~ depth * tree mortality, block = tree identity

### Prevalence of ectomycorrhizal fungi

#### 5) Relative abundance; log(proportion of RNA-based ITS ASVs belonging to ectomycorrhizal fungi) ~ depth * tree mortality, block = tree identity
#### 6) Imputed biomass; log((Proportion of RNA-based ITS ASVs belonging to ectomycorrhizal fungi) * (total fungal PLFA)) ~ depth * tree mortality, block = tree identity

### Carbon by depth and tree mortality

#### 7) Total carbon; (MgC/ha) ~ depth * tree mortality, block = tree identity
#### 8) C:N ratio; (C:N) ~ depth * tree mortality, block = tree identity

### Carbon by ectomycorrhizal-associated biomass

#### 9) Total carbon; MgC/ha ~ (imputed biomass of ectomycorrhizal fungi + imputed biomass of ectomycorrhizal-assoicated bacteria), block = tree identity
#### 10) C:N ratio; (C:N) ~ (imputed biomass of ectomycorrhizal fungi + imputed biomass of ectomycorrhizal-assoicated bacteria), block = tree identity

## load required packages:

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
```

## Models 1, 2, 7, and 8: untransformed, nucleic acid-independent LMEs

### load data
```{r}
metadata36 <- read.csv("../Carbon/Density_Corrected_C/Results/per_sample_metadata_Ccorrected_mip.csv") # load metadata containing all 36 samples (other versions which include nucleic acid data contain fewer samples, reflecting sample loss during nucleic acid extraction and amplification processes)
```


### make a list of response variables to be modeled with untransformed, nucleic acid-independent LMEs
```{r}
nonlog_36_over_depth = c("PLFA_Total_Living_Microbial_Biomass_ng_g", "PLFA_Total_Bacteria_ng_g", "C_per_N_Ratio", "mass_C_Mg_per_ha") # these variables do not require transformation, as their untransformed LMEs lmes meet statistical assumptions of normality
```

### generate LMEs
```{r}
lmes.depth.36 = list() # initialize empty list to store these models
lmes.depth.36_summaries = list() # initialize empty list to store model summaries
lmes.depth.36_summaries_df = list() # initialize empty list to store model summaries

for (var in nonlog_36_over_depth) { # begin a loop which iterates through the list of variables modeled with this common formula, creating the model for each
  formula = as.formula(paste(var, "~core_bottom_depth_cm * Tree_Health + (1|Tree)")) # define the common formula
  
  lme1 <- lmer(formula, data = metadata36) # create the model
  
  lmes.depth.36[[var]] = lme1 # store the model in the list initialized above
  lmes.depth.36_summaries[[var]] = summary(lme1) # store the model summary in the list initialized above
  lmes.depth.36_summaries_df[[var]] = as.data.frame(summary(lme1)$coefficients) # store dataframe of the model summary in the list initialized above
}

lmes.depth.36_summaries # display model summaries


```

### view diagnostic plots comparing each model's residuals to fitted values
```{r}
modelcheckplots = list() # initialize empty list to store diagnostic plots of these models
for (var in nonlog_36_over_depth) { # begin a loop which iterates through the list of variables modeled with this common formula, generating diagnostic plots for each
  p = plot(lmes.depth.36[[var]]) # generate diagnostic plots
  modelcheckplots[[var]] = p # store plots in the list initialized above
}

modelcheckplots # display diagnostic plots of each model
```

### check normality of residuals with Shapiro-Wilks Test, residuals are normal if p > 0.05
```{r}
shap.tests = list() # initialize empty list to store results of the Shapiro-Wilks test
for (var in nonlog_36_over_depth) { # begin a loop which iterates through the list of variables modeled with this common formula, performing the Shapiro-Wilks test for each
  shap.test1 = shapiro.test(residuals(lmes.depth.36[[var]])) # perform Shapiro-Wilks test
  shap.tests[[var]] = shap.test1 # store results in the list initialized above
}

shap.tests # display test results
```

### Model 2 extended: Address the tree health:depth interaction in total bacteria
```{r}
# separate data from living and dead trees

living = filter(metadata36, Tree_Health == "living") # make new dataframe containing only data from soils under living trees
dead = filter(metadata36, Tree_Health == "dead") # make new dataframe containing only data from soils under living dead

living.bac = lmer(PLFA_Total_Bacteria_ng_g ~ core_bottom_depth_cm + (1|Tree), data = living) # recreate the above LME on living tree samples only, omitting Tree_Health

summary(living.bac)  # display model summary
model2_living_bac_summary = as.data.frame(summary(living.bac)$coefficients) # store summary as dataframe


plot(living.bac) # display diagnostic plot
shapiro.test(residuals(living.bac)) # display Shapiro-Wilks test results

dead.bac = lmer(PLFA_Total_Bacteria_ng_g ~ core_bottom_depth_cm + (1|Tree), data = dead) # recreate the above LME on dead tree samples only, omitting Tree_Health

model2_dead_bac_summary = as.data.frame(summary(dead.bac)$coefficients) # store summary as dataframe) 
summary(dead.bac) # display model summary

plot(dead.bac) # display diagnostic plots
shapiro.test(residuals(dead.bac)) # display Shapiro-Wilks test results

```


## Models 3 and 4: log-transformed, nucleic acid-independent LMEs

### 3) Total living fungal biomass; log(PLFA) ~ depth * tree mortality, block = tree identity

#### generate model
```{r}
PLFA_fungi = lmer(log(PLFA_Total_Fungi_ng_g) ~ core_bottom_depth_cm * Tree_Health + (1|Tree), data = metadata36) # generate model

model3_summary = as.data.frame(summary(PLFA_fungi)$coefficients) # store summary as dataframe
summary(PLFA_fungi) # display model summary
```

#### check statistical assumptions
```{r}
plot(PLFA_fungi) # generating diagnostic plots

shapiro.test(residuals(PLFA_fungi)) # perform Shapiro-Wilks test for normality of residuals, residuals are normal if p > 0.05
```

### Model 3 extended: Address the tree health:depth interaction in total fungi

```{r}
living.fun = lmer(log(PLFA_Total_Fungi_ng_g) ~ core_bottom_depth_cm + (1|Tree), data = living) # recreate the above LME on living tree samples only, omitting Tree_Health
model3_living_fun_summary = as.data.frame(summary(living.fun)$coefficients) # store summary as dataframe 
summary(living.fun) # display model summary

plot(living.fun) # display diagnostic plots
shapiro.test(residuals(living.fun)) # display Shapiro-Wilks test results




dead.fun = lmer(log(PLFA_Total_Fungi_ng_g) ~ core_bottom_depth_cm + (1|Tree), data = dead) # recreate the above LME on dead tree samples only, omitting Tree_Health
model3_dead_fun_summary = as.data.frame(summary(dead.fun)$coefficients) # store summary as dataframe ) 
summary(dead.fun) # display model summary

plot(dead.fun) # display diagnostic plots
shapiro.test(residuals(dead.fun)) # display Shapiro-Wilks test results
```


### 4) Fungi:bacteria ratio; log(total fungal PLFA/total bacterial PLFA) ~ depth * tree mortality, block = tree identity

#### generate model
```{r}
PLFA_fungi_bacteria_ratio = lmer(log(PLFA_Fungi_Bacteria_Ratio_ng_g) ~ core_bottom_depth_cm * Tree_Health + (1|Tree), data = metadata36) # generate model

model4_summary = as.data.frame(summary(PLFA_fungi_bacteria_ratio)$coefficients) # store summary as dataframe  
summary(PLFA_fungi_bacteria_ratio) # display model summary
```

#### check statistical assumptions
```{r}
plot(PLFA_fungi_bacteria_ratio) # generating diagnostic plots

shapiro.test(residuals(PLFA_fungi_bacteria_ratio)) # perform Shapiro-Wilks test for normality of residuals, residuals are normal if p > 0.05
```


## Models 5 and 6: log-transformed, nucleic acid-dependent LMEs

### 5) Relative abundance; log(proportion of RNA-based ITS ASVs belonging to ectomycorrhizal fungi) ~ depth * tree mortality, block = tree identity

#### load data

```{r}
metadata21.relabund <- read.csv("../Microbial Communities/Imputed Biomass/Results/RNA_based_fungal_trait_relabund_mip.csv") # load metadata containing the relative abundance of all fungal primary lifestyles, as assigned to RNA-based ITS ASVs by the Fungal Traits database. This dataframe has no NAs, and contains data from each sample for which we successfully amplified ITS cDNA. n = 21.

metadata21.biomass <- read.csv("../Microbial Communities/Imputed Biomass/Results/RNA_based_fungal_trait_biomass_mip.csv") # load metadata containing the imputed biomass of all fungal primary lifestyles, as assigned to RNA-based ITS ASVs by the Fungal Traits database. This dataframe has no NAs, and contains data from each sample for which we successfully amplified ITS cDNA. n = 21. 

metadata27 <- read.csv("../Microbial Communities/Fungal-Bacterial Interactions/DeSeq2 Indicator Species Analysis/Results/metadata_with_ECMasssociated3trees_mip.csv") # load metadata containing imputed biomass of ectomycorrhizal fungi and ectomycorrhizal-associated bacteria. This dataframe has NAs, as it contains data from each sample for which we successfully amplified ITS cDNA OR 16S cDNA (both RNA-derived). Only 20 samples had both.
metadata27 <- mutate(metadata27, ectomycorrhizal_relabund = ectomycorrhizal/PLFA_Total_Fungi_ng_g) # add ectomycorrhizal relative abundance
metadata27 = dplyr::rename(metadata27, ectomycorrhizal_biomass = ectomycorrhizal) # rename imputed biomass of ectomycorrhizal fungi to distinguish from relatibe abundance

metadata27_tidy = filter(metadata27, !(is.na(metadata27$ectomycorrhizal_relabund))) # take the rows with NAs out of metadata27. n = 20.
```

#### generate model from metadata21
```{r}
ecm_relabund = lmer(log(ectomycorrhizal) ~ core_bottom_depth_cm * Tree_Health + (1|Tree), data = metadata21.relabund) # generate model

model5_ecm_relabund_summary = as.data.frame(summary(ecm_relabund)$coefficients) # store summary as dataframe
summary(ecm_relabund) # display model summary
```

##### check statistical assumptions
```{r}
plot(ecm_relabund) # generating diagnostic plots

shapiro.test(residuals(ecm_relabund)) # perform Shapiro-Wilks test for normality of residuals, residuals are normal if p > 0.05
```


### 6) Imputed biomass; log((Proportion of RNA-based ITS ASVs belonging to ectomycorrhizal fungi) * (total fungal PLFA)) ~ depth * tree mortality, block = tree identity

#### generate model from metadata21.biomass
```{r}
ecm_biomass.21 = lmer(log(ectomycorrhizal) ~ core_bottom_depth_cm * Tree_Health + (1|Tree), data = metadata21.biomass) # generate model

model6_ecm_biomass_summary = as.data.frame(summary(ecm_biomass.21)$coefficients) # store summary as dataframe
summary(ecm_biomass.21) # display model summary
```

#### check statistical assumptions
```{r}
plot(ecm_biomass.21) # generating diagnostic plots

shapiro.test(residuals(ecm_biomass.21)) # perform Shapiro-Wilks test for normality of residuals, residuals are normal if p > 0.05
```

## Models 9 and 10: untransformed, nucleic acid dependent LMEs

### 9) Total carbon; MgC/ha ~ (imputed biomass of ectomycorrhizal fungi + imputed biomass of ectomycorrhizal-assoicated bacteria), block = tree identity

#### generate model from metadata27
```{r}
totalC = lmer(mass_C_Mg_per_ha ~ ECM_associated_biomass + (1|Tree), data = metadata27_tidy) # generate model

model9_totalC_summary = as.data.frame(summary(totalC)$coefficients) # store summary as dataframe
summary(totalC) # display model summary
```

#### check statistical assumptions
```{r}
plot(totalC) # generating diagnostic plots

shapiro.test(residuals(totalC)) # perform Shapiro-Wilks test for normality of residuals, residuals are normal if p > 0.05
```



### 10) C:N ratio; (C:N) ~ (imputed biomass of ectomycorrhizal fungi + imputed biomass of ectomycorrhizal-assoicated bacteria), block = tree identity

```{r}
C_per_N = lmer(C_per_N_Ratio ~ ECM_associated_biomass + (1|Tree), data = metadata27_tidy) # generate model

summary(C_per_N) # display model summary
model10_C_per_N_summary= as.data.frame(summary(C_per_N)$coefficients) # store summary as dataframe
```

#### check statistical assumptions
```{r}
plot(C_per_N) # generating diagnostic plots

shapiro.test(residuals(C_per_N)) # perform Shapiro-Wilks test for normality of residuals, residuals are normal if p > 0.05
```


# compile model summaries for export
```{r}
# Models 1, 2, 7, and 8: untransformed, nucleic acid-independent LMEs
# lmes.depth.36_summaries_df is already a list of summaries
names(lmes.depth.36_summaries_df) = c("Model 1", "Model 2", "Model 7", "Model 8")



# start our big list of all the model summaries with this existing list 
all_models = lmes.depth.36_summaries_df

# Model 2 extended
all_models[["Model 2.1 - living trees only"]] <- model2_living_bac_summary
all_models[["Model 2.2 - dead trees only"]] <- model2_dead_bac_summary

# Models 3 and 4: log-transformed, nucleic acid-independent LMEs
all_models[["Model 3"]] <- model3_summary
all_models[["Model 3.1 - living trees only"]] <- model3_living_fun_summary
all_models[["Model 3.2 - dead trees only"]] <- model3_dead_fun_summary
all_models[["Model 4"]] <- model4_summary

# Models 5 and 6: log-transformed, nucleic acid-dependent LMEs
all_models[["Model 5"]] <- model5_ecm_relabund_summary
all_models[["Model 6"]] <- model6_ecm_biomass_summary

# Models 9 and 10: untransformed, nucleic acid dependent LMEs
all_models[["Model 9"]] <- model9_totalC_summary
all_models[["Model 10"]] <- model10_C_per_N_summary


```

```{r}
all_models <- Map(function(df, name) { # Add the name of the model as a column, and put it on the far left side of the dataframe
  df$Model <- name
  df
}, all_models, names(all_models))


```

```{r}
all_models <- lapply(all_models, function(df) {
  df <- cbind(Factor = rownames(df), df) # copy rownames into a column, put it on the left side of the dataframe
  rownames(df) <- NULL # remove the rownames now that factors are represented in a column
  
  df <- select(df, c(Model, Factor, Estimate, 'Std. Error', df, 't value', 'Pr(>|t|)')) # reorder the columns to Model comes first
  
  df
})

```


## bind tidy results into a single dataframe
```{r}
table <- do.call(rbind, all_models)

rownames(table) <- NULL # remove the rownames now that factors are represented in a column

colnames(table)

table <- table %>%
      mutate(Factor = case_when(
        Factor == "core_bottom_depth_cm" ~ "Depth",
        Factor == "Tree_Healthliving" ~ "Tree mortality",
        Factor == "core_bottom_depth_cm:Tree_Healthliving" ~ "Depth:Tree mortality",
        Factor == "ECM_associated_biomass" ~ "Biomass sum of ectomycorrhizal fungi + associated bacteria",
        TRUE ~ Factor # Keep other values unchanged
      ))

table
```

## export
``` {r}
write_csv(table, "Results/LME_table.csv")
write_csv(table, "../../Reports/Supplementary Material/LME_table.csv")

```

