---
title: "DO_Tidy_Extractions_Used_For_Sequencing"
author: "Glade Bogar"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Load required packages:
```{r}
library(tidyverse)
```

#Load required data:
```{r}
extractions <- read_csv("../../Data/Transcribed data/RNA-DNA Extractions and Lib prep planning/Extractions Aliquoted For Library Prep.csv")
```

#Make a soil sample ID constent with other data:
```{r}
extractions <- extractions %>%
  mutate(new_column = "DO")

extractions <- extractions %>%
  unite('DO_ID', 'new_column', 'Tree', 'Depth cm', sep="_")
```

#Rename variable names:
```{r}
extractions <- extractions %>%
  rename(extraction_date = `Extraction Date`)
extractions <- extractions %>%
  rename(date_collected = `Date Collected`)
extractions <- extractions %>%
  rename(starting_depth_cm = `Starting depth (cm)`)
extractions <- extractions %>%
  rename(RNA_ng_per_uL = `RNA ng/uL`)
extractions <- extractions %>%
  rename(DNA_ng_per_uL = `DNA ng/uL`)
extractions <- extractions %>%
  rename(twenty_uL_RNA_dilution_2_ng_per_uL = `20uL RNA DILUTION 2ng/uL`)
extractions <- extractions %>%
  rename(twenty_uL_DNA_dilution_2_ng_per_uL = `20uL DNA Dilution 2ng/uL`)
extractions <- extractions %>%
  rename(soil_mass_g_extraction = `Soil Mass (g)`)
extractions <- extractions %>%
  rename(batch_letter = `Letter in Batch`)
extractions <- extractions %>%
  rename(DO_RNA_Box_ID = `DO RNA Box ID`)
extractions <- extractions %>%
  rename(DO_DNA_Box_ID = `DO DNA Box ID`)
extractions <- extractions %>%
  rename(RNA_PCR_water = `RNA PCR Water to be Added?`)
extractions <- extractions %>%
  rename(DNA_PCR_water = `DNA PCR Water to be added?`)
extractions <- extractions %>%
  rename(cDNA_conc_ng_per_uL = `cDNA conc. ng/uL`)
extractions <- extractions %>%
  rename(uL_cDNA_input = `uL of cDNA to input`)
extractions <- extractions %>%
  rename(cDNA_final_input_ng_per_uL = `Final Input ng`)
```

#Complex dataset. Sure there must be a smart way to deal with this, but instead I'm going to split and re-join to legthen data.

#write .csv jut for copy/paste:
```{r}
write_csv(extractions, "extractions_temp.csv")
```

#Select general extraction variables:
```{r}
extractions_general <- select(extractions,
                              DO_ID,
                              extraction_date,
                              date_collected,
                              starting_depth_cm,
                              soil_mass_g_extraction,
                              batch_letter,
                              notes_RNA_DNA,
                              )
```


#Select DNA extraction variables:
```{r}
extractions_DNA <- select(extractions,
                          DO_ID,
                          DNA_ng_per_uL,
                          twenty_uL_DNA_dilution_2_ng_per_uL,
                          DO_DNA_Box_ID,
                          DNA_PCR_water)
```

#Select RNA extraction variables:
```{r}
extractions_RNA <- select(extractions,
                          DO_ID,
                          RNA_ng_per_uL,
                          twenty_uL_RNA_dilution_2_ng_per_uL,
                          DO_RNA_Box_ID,
                          RNA_PCR_water,
                          cDNA_conc_ng_per_uL,
                          uL_cDNA_input,
                          cDNA_final_input_ng_per_uL)
```

#Rename variables for join:
```{r}
extractions_general <- rename(extractions_general, notes_extraction = "notes_RNA_DNA")

extractions_DNA <- rename(extractions_DNA, Extraction_ID = "DO_DNA_Box_ID")

extractions_RNA <- rename(extractions_RNA, Extraction_ID = "DO_RNA_Box_ID")
```

#Left join general extraction variables to each DNA, RNA sheets:
```{r}
extractions_DNA <- left_join(extractions_general, extractions_DNA)
extractions_RNA <- left_join(extractions_general, extractions_RNA)
```


#Bind rows to make joined datasheet by individual DNA/RNA extraction
```{r}
extractions_tidy <- bind_rows(extractions_DNA, extractions_RNA)
```

#export tidy data:
```{r}
write_csv(extractions_tidy, "Results/nucleic_acid_extractions_for_library_prep_tidy_mip.csv")
```






















