---
title: "microeco"
author: "Glade Bogar"
date: "2024-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Implement FAPROTAX using package microeco. Follow along with this tutorial: https://chiliubio.github.io/microeco_tutorial/

#install all needed packages specific to microeco (overkill, but recommended for tutorial):
```{r install, eval = FALSE}
# allow more waiting time to download each package
options(timeout = 1000)
# If a package is not installed, it will be installed from CRAN
# First select the packages of interest
tmp <- c("microeco", "mecoturn", "MASS", "GUniFrac", "ggpubr", "randomForest", "ggdendro", "ggrepel", "agricolae", "igraph", "picante", "pheatmap", "rgexf", 
    "ggalluvial", "ggh4x", "rcompanion", "FSA", "gridExtra", "aplot", "NST", "GGally", "ggraph", "networkD3", "poweRlaw", "ggtern", "SRS", "performance")
# Now check or install
for(x in tmp){
    if(!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
    }
}

install.packages("BiocManager")
install.packages("file2meco", repos = BiocManager::repositories())
install.packages("MicrobiomeStat", repos = BiocManager::repositories())
install.packages("WGCNA", repos = BiocManager::repositories())
BiocManager::install("ggtree")
BiocManager::install("metagenomeSeq")
BiocManager::install("ALDEx2")
BiocManager::install("ANCOMBC")

# download link of the compressed packages archive
```

# Alternative package installation from Gitee "https://gitee.com/chiliubio/microeco_dependence/releases/download/v0.20.0/microeco_dependence.zip"
```{r}
url <- "https://github.com/ChiLiubio/microeco_dependence/releases/download/v0.20.0/microeco_dependence.zip"
# allow more time to download the zip file in R
options(timeout = 2000)
# Another way is to open the upper url in browser to download the zip file and move it to the current R working directory
download.file(url = url, destfile = "microeco_dependence.zip")
# uncompress the file in R
tmp <- "microeco_dependence"
unzip(paste0(tmp, ".zip"))
# install devtools
if(!require("devtools", character.only = TRUE)){install.packages("devtools", dependencies = TRUE)}
# run these one by one
devtools::install_local(paste0(tmp, "/", "SpiecEasi-master.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "mixedCCA-master.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "SPRING-master.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "NetCoMi-main.zip"), repos = BiocManager::repositories())
devtools::install_local(paste0(tmp, "/", "beem-static-master.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "chorddiag-master.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "ggradar-master.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "ggnested-main.zip"), dependencies = TRUE)
devtools::install_local(paste0(tmp, "/", "ggcor-1-master.zip"), dependencies = TRUE)
```


#load required packages:
```{r}
library(tidyverse)
library(phyloseq)
library(microeco)
```

#begin tutorial:

#load phyloseq object:
```{r}
ps16S <- read_rds("../Phyloseq_16S/Results/ps16S_remove_3_reads.rds")
ps16S
```

#read depths:
```{r}
sort(sample_sums(ps16S))
```


#get microeco object components from the 16S phyloseq:
```{r}
ps16S_sample_data <- data.frame(sample_data(ps16S))
ps16S_asv_table <- data.frame(otu_table(ps16S))
ps16S_taxa_table <- data.frame(tax_table(ps16S))
```


#check data classes for compatibility:
```{r}
class(ps16S_sample_data)
class(ps16S_asv_table)
class(ps16S_taxa_table)
```

#check format for compatibility:
```{r}
ps16S_sample_data[1:5, 1:5]
ps16S_asv_table[1:5, 1:5]
ps16S_taxa_table[1:5, 1:5]
```

#looks like it should work

# make the taxonomic information unified, very important
```{r}
ps16S_taxa_table <- tidy_taxonomy(ps16S_taxa_table)
```

#are the rownames still there?
```{r}
ps16S_taxa_table[1:5, 1:5]
```

#yes

#similar to phyloseq, now create a 'microtable' object:
```{r}
mt16S <- microtable$new(sample_table = ps16S_sample_data,
                                 otu_table = ps16S_asv_table,
                                 tax_table = ps16S_taxa_table)
mt16S
```

#that seems to have worked

#skip ahead to 7.2 trans_func class for FAPROTAX assignment

#could it be this easy?

#so what does it think the sample names are?

#see if some other basic functions will work
```{r}
mt16S$sample_sums() %>% range
```

#ok, what does it think the sample names are?
```{r}
mt16S$sample_names()
```

#Well, that's correct.

#check individual things:
```{r}
rownames(mt16S$sample_table)[1:5]
```

#check individual things:
```{r}
colnames(mt16S$otu_table)[1:5]
```

#ok so the otu table needs to be inverted, maybe?

#try:
```{r}
ps16S_asv_table_t <- data.frame(t(ps16S_asv_table))

colnames(ps16S_asv_table_t)[1:5]
rownames(ps16S_asv_table_t)[1:5]
```

#similar to phyloseq, now create a 'microtable' object, with transposed ASV table:
```{r}
mt16S <- microtable$new(sample_table = ps16S_sample_data,
                                 otu_table = ps16S_asv_table_t,
                                 tax_table = ps16S_taxa_table)
mt16S
```

#re-try:default agglomeration function for every level, how long will it run?
```{r}
mt16S$cal_abund()
```

#worked, and pretty fast.

#re-try functional trait assignment:
```{r}
mt16S_func <- trans_func$new(mt16S)
# mapping the taxonomy to the database
# this can recognize prokaryotes or fungi automatically if the names of taxonomic levels are standard.
# for fungi example, see https://chiliubio.github.io/microeco_tutorial/other-dataset.html#fungi-data
# default database for prokaryotes is FAPROTAX database
mt16S_func$cal_spe_func(prok_database = "FAPROTAX")
```

## return t2$res_spe_func, 1 represent trait exists, 0 represent no or cannot confirmed.
```{r}
mt16S_func$res_spe_func[1:5, 1:2]
```

#get all:
```{r}
mt16S_func_traitsbinary <- mt16S_func$res_spe_func
mt16S_func_traitsbinary <- rownames_to_column(mt16S_func_traitsbinary, var = "ASV")

write_csv(mt16S_func_traitsbinary, "Results/16S_traits_mip.csv")
```


# calculate the percentages for communities:
```{r}
mt16S_func$cal_spe_func_perc(abundance_weighted = FALSE)
```

#observe:
```{r}
mt16S_func$res_spe_func_perc[1:5, 1:2]
```

#hoo ho hoo:
```{r}
mt16S_func_traits_persample <- mt16S_func$res_spe_func_perc
mt16S_func_traits_persample <- mt16S_func_traits_persample %>%
  mutate_all(~ . / 100) #convert from % to relative abundance
mt16S_func_traits_persample <- rownames_to_column(mt16S_func_traits_persample, var = "Seq_ID")

write_csv(mt16S_func_traits_persample, "Results/16S_traits_per_sample_mip.csv")
```

#attempt to de-nest functional traits:
```{r}
mt16S_func$trans_spe_func_perc()
```

#retrieve:
```{r}
mt16S_func_long <- mt16S_func$res_spe_func_perc_trans
```

#don't think that did what I need it to do.

##add 16S functional assignments to 16S ps, to repeat fungal functional trait analyses with 16S.

#load the fungal traits phyloseq, to see what we need to do:
```{r}
psITS <- read_rds("../Phyloseq_ITS/Results/psITS_remove_8_reads_fungaltraits.rds")
psITS
```

#list available functional traits:
```{r}
rank_names(psITS)
```

#view taxa table:
```{r}
psITS_taxa_table <- data.frame(tax_table(psITS))
```

#so we need one column, 'FAPROTAX', with one function name per ASV for the taxa table.

#issue; there are more than one function per OTU...

#combine them all, in alphabetic order, to create distinct functional trait combinations. Not elegant, but if there are limited combinations, then will be usable. (i.e., if traits cluster and there are few permutations).

#first assign column titles where there are currently '1' values in the binary datasheet:
```{r}
class(mt16S_func_traitsbinary)

mt16S_func_traitsbinary_char <- mt16S_func_traitsbinary %>% 
  lapply(as.character) %>%
  as.data.frame()

class(mt16S_func_traitsbinary_char)

mt16S_func_traitsnamed <- mt16S_func_traitsbinary_char %>%
  transmute(across(everything(), ~str_replace_all(., "1", cur_column()))) #replace "1" with column title

mt16S_func_traitsnamed <- column_to_rownames(mt16S_func_traitsnamed, var = "ASV")

mt16S_func_traitsnamed_no0 <- mt16S_func_traitsnamed

mt16S_func_traitsnamed_no0[] <- lapply(mt16S_func_traitsnamed, function(x) str_replace_all(x, "0", ""))

mt16S_func_traitsnamed_no0 <- select(mt16S_func_traitsnamed_no0, -chemoheterotrophy) #no ASVs only 'chemoheterotrophy', redundant

mt16S_func_traitsnamed_no0_test <- head(mt16S_func_traitsnamed_no0)

mt16S_func_traitsnamed_no0_test_cat <- mt16S_func_traitsnamed_no0_test %>%
  mutate(combined_traits = apply(mt16S_func_traitsnamed_no0_test, 1, function(x) paste(x, collapse = "")))

mt16S_func_traitsnamed_no0_cat <- mt16S_func_traitsnamed_no0 %>%
  mutate(combined_traits = apply(mt16S_func_traitsnamed_no0, 1, function(x) paste(x, collapse = "")))

mt16S_func_traitsnamed_no0_cat <- rownames_to_column(mt16S_func_traitsnamed_no0_cat, var = "ASV")

write_csv(mt16S_func_traitsnamed_no0_cat, "Results/concatenatedtraitstest_mip.csv")
```

#select and name "unassigned":
```{r}
mt16S_func_traits_for_join <- select(mt16S_func_traitsnamed_no0_cat, ASV, combined_traits)

mt16S_func_traits_for_join_NA <- mt16S_func_traits_for_join %>%
  mutate(across(everything(), ~na_if(str_trim(.), "")))

mt16S_func_traits_for_join_unassigned <- mt16S_func_traits_for_join_NA %>%
  mutate(across(everything(), ~replace_na(., "unassigned")))
```

#join to 16S taxa table as a new taxonomic rank, as fungaltraits:
```{r}
ps16S_taxa_table_for_join <- rownames_to_column(ps16S_taxa_table, var = "ASV")

ps16S_taxa_table_for_join <- left_join(ps16S_taxa_table_for_join, mt16S_func_traits_for_join_unassigned)

ps16S_taxa_table_for_join <- column_to_rownames(ps16S_taxa_table_for_join, var = "ASV")
```

#re-make original phyloseq object with functions:
```{r}
ps16S_taxa_table_traits <- as.matrix(ps16S_taxa_table_for_join)
row.names(ps16S_taxa_table_traits) <- row.names(ps16S_taxa_table_for_join)

taxa_16S_traits <- tax_table(ps16S_taxa_table_traits)

tax_table(ps16S) <- taxa_16S_traits

ps16S
```

#think that worked.

#export:
```{r}
write_rds(ps16S, "Results/ps16S_faprotax_mip.rds")
```


# anna check
```{r}
dd <- read_rds("Results/ps16S_faprotax.rds")
identical(dd, ps16S)
```





